### 2014-11-21 scifile.tex
tex Datei erstellt

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}

% Use times if you have the font installed; otherwise, comment out the
% following line.

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{A simple {\it Science\/} Template}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{John Smith,$^{1\ast}$ Jane Doe,$^{1}$ Joe Scientist$^{2}$\\
\\
\normalsize{$^{1}$Department of Chemistry, University of Wherever,}\\
\normalsize{An Unknown Address, Wherever, ST 00000, USA}\\
\normalsize{$^{2}$Another Unknown Address, Palookaville, ST 99999, USA}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  jsmith@wherever.edu.}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
  This document presents a number of hints about how to set up your
  {\it Science\/} paper in \LaTeX\ .  We provide a template file,
  \texttt{scifile.tex}, that you can use to set up the \LaTeX\ source
  for your article.  An example of the style is the special
  \texttt{\{sciabstract\}} environment used to set up the abstract you
  see here.
\end{sciabstract}

% In setting up this template for *Science* papers, we've used both
% the \section* command and the \paragraph* command for topical
% divisions.  Which you use will of course depend on the type of paper
% you're writing.  Review Articles tend to have displayed headings, for
% which \section* is more appropriate; Research Articles, when they have
% formal topical divisions at all, tend to signal them with bold text
% that runs into the paragraph, for which \paragraph* is the right
% choice.  Either way, use the asterisk (*) modifier, as shown, to
% suppress numbering.

\section*{Introduction}

In this file, we present some tips and sample mark-up to assure your
\LaTeX\ file of the smoothest possible journey from review manuscript
to published {\it Science\/} paper.  We focus here particularly on
issues related to style files, citation, and math, tables, and
figures, as those tend to be the biggest sticking points.  Please use
the source file for this document, \texttt{scifile.tex}, as a template
for your manuscript, cutting and pasting your content into the file at
the appropriate places.

{\it Science\/}'s publication workflow relies on Microsoft Word.  To
translate \LaTeX\ files into Word, we use an intermediate MS-DOS
routine \cite{tth} that converts the \TeX\ source into HTML\@.  The
routine is generally robust, but it works best if the source document
is clean \LaTeX\ without a significant freight of local macros or
\texttt{.sty} files.  Use of the source file \texttt{scifile.tex} as a
template, and calling {\it only\/} the \texttt{.sty} and \texttt{.bst}
files specifically mentioned here, will generate a manuscript that
should be eminently reviewable, and yet will allow your paper to
proceed quickly into our production flow upon acceptance \cite{use2e}.

\section*{Formatting Citations}

Citations can be handled in one of three ways.  The most
straightforward (albeit labor-intensive) would be to hardwire your
citations into your \LaTeX\ source, as you would if you were using an
ordinary word processor.  Thus, your code might look something like
this:

\begin{quote}
\begin{verbatim}
However, this record of the solar nebula may have been
partly erased by the complex history of the meteorite
parent bodies, which includes collision-induced shock,
thermal metamorphism, and aqueous alteration
({\it 1, 2, 5--7\/}).
\end{verbatim}
\end{quote}

\noindent Compiled, the last two lines of the code above, of course, would give notecalls in {\it Science\/} style:

\begin{quote}
\ldots thermal metamorphism, and aqueous alteration ({\it 1, 2, 5--7\/}).
\end{quote}

Under the same logic, the author could set up his or her reference list as a simple enumeration,

\begin{quote}
\begin{verbatim}
{\bf References and Notes}

\begin{enumerate}
\item G. Gamow, {\it The Constitution of Atomic Nuclei
and Radioactivity\/} (Oxford Univ. Press, New York, 1931).
\item W. Heisenberg and W. Pauli, {\it Zeitschr.\ f.\ 
Physik\/} {\bf 56}, 1 (1929).
\end{enumerate}
\end{verbatim}
\end{quote}

\noindent yielding

\begin{quote}
{\bf References and Notes}

\begin{enumerate}
\item G. Gamow, {\it The Constitution of Atomic Nuclei and
Radioactivity\/} (Oxford Univ. Press, New York, 1931).
\item W. Heisenberg and W. Pauli, {\it Zeitschr.\ f.\ Physik} {\bf 56},
1 (1929).
\end{enumerate}
\end{quote}

That's not a solution that's likely to appeal to everyone, however ---
especially not to users of B{\small{IB}}\TeX\ \cite{inclme}.  If you
are a B{\small{IB}}\TeX\ user, we suggest that you use the
\texttt{Science.bst} bibliography style file and the
\texttt{scicite.sty} package, both of which we are downloadable from our author help site
(http://www.sciencemag.org/about/authors/prep/TeX\_help/).  You can also
generate your reference lists by using the list environment
\texttt{\{thebibliography\}} at the end of your source document; here
again, you may find the \texttt{scicite.sty} file useful.

Whether you use B{\small{IB}}\TeX\ or \texttt{\{thebibliography\}}, be
very careful about how you set up your in-text reference calls and
notecalls.  In particular, observe the following requirements:

\begin{enumerate}
\item Please follow the style for references outlined at our author
  help site and embodied in recent issues of {\it Science}.  Each
  citation number should refer to a single reference; please do not
  concatenate several references under a single number.
\item Please cite your references and notes in text {\it only\/} using
  the standard \LaTeX\ \verb+\cite+ command, not another command
  driven by outside macros.
\item Please separate multiple citations within a single \verb+\cite+
  command using commas only; there should be {\it no space\/}
  between reference keynames.  That is, if you are citing two
  papers whose bibliography keys are \texttt{keyname1} and
  \texttt{keyname2}, the in-text cite should read
  \verb+\cite{keyname1,keyname2}+, {\it not\/}
  \verb+\cite{keyname1, keyname2}+.
\end{enumerate}

\noindent Failure to follow these guidelines could lead
to the omission of the references in an accepted paper when the source
file is translated to Word via HTML.

\section*{Handling Math, Tables, and Figures}

Following are a few things to keep in mind in coding equations,
tables, and figures for submission to {\it Science}.

\paragraph*{In-line math.}  The utility that we use for converting
from \LaTeX\ to HTML handles in-line math relatively well.  It is best
to avoid using built-up fractions in in-line equations, and going for
the more boring ``slash'' presentation whenever possible --- that is,
for \verb+$a/b$+ (which comes out as $a/b$) rather than
\verb+$\frac{a}{b}$+ (which compiles as $\frac{a}{b}$).  Likewise,
HTML isn't tooled to handle certain overaccented special characters
in-line; for $\hat{\alpha}$ (coded \verb+$\hat{\alpha}$+), for
example, the HTML translation code will return [\^{}$(\alpha)$].
Don't drive yourself crazy --- but if it's possible to avoid such
constructs, please do so.  Please do not code arrays or matrices as
in-line math; display them instead.  And please keep your coding as
\TeX-y as possible --- avoid using specialized math macro packages
like \texttt{amstex.sty}.

\paragraph*{Displayed math.} Our HTML converter sets up \TeX\
displayed equations using nested HTML tables.  That works well for an
HTML presentation, but Word chokes when it comes across a nested
table in an HTML file.  We surmount that problem by simply cutting the
displayed equations out of the HTML before it's imported into Word,
and then replacing them in the Word document using either images or
equations generated by a Word equation editor.  Strictly speaking,
this procedure doesn't bear on how you should prepare your manuscript
--- although, for reasons best consigned to a note \cite{nattex}, we'd
prefer that you use native \TeX\ commands within displayed-math
environments, rather than \LaTeX\ sub-environments.

\paragraph*{Tables.}  The HTML converter that we use seems to handle
reasonably well simple tables generated using the \LaTeX\
\texttt{\{tabular\}} environment.  For very complicated tables, you
may want to consider generating them in a word processing program and
including them as a separate file.

\paragraph*{Figures.}  Figure callouts within the text should not be
in the form of \LaTeX\ references, but should simply be typed in ---
that is, \verb+(Fig. 1)+ rather than \verb+\ref{fig1}+.  For the
figures themselves, treatment can differ depending on whether the
manuscript is an initial submission or a final revision for acceptance
and publication.  For an initial submission and review copy, you can
use the \LaTeX\ \verb+{figure}+ environment and the
\verb+\includegraphics+ command to include your PostScript figures at
the end of the compiled PostScript file.  For the final revision,
however, the \verb+{figure}+ environment should {\it not\/} be used;
instead, the figure captions themselves should be typed in as regular
text at the end of the source file (an example is included here), and
the figures should be uploaded separately according to the Art
Department's instructions.

\textit{•}\section*{What to Send In}

What you should send to {\it Science\/} will depend on the stage your manuscript is in:

\begin{itemize}
\item {\bf Important:} If you're sending in the initial submission of
  your manuscript (that is, the copy for evaluation and peer review),
  please send in {\it only\/} a PostScript or PDF version of the
  compiled file (including figures).  Please do not send in the \TeX\ 
  source, \texttt{.sty}, \texttt{.bbl}, or other associated files with
  your initial submission.  (For more information, please see the
  instructions at our Web submission site,
  http://www.submit2science.org/ .)
\item When the time comes for you to send in your revised final
  manuscript (i.e., after peer review), we require that you include
  all source files and generated files in your upload.  Thus, if the
  name of your main source document is \texttt{ltxfile.tex}, you
  need to include:
\begin{itemize}
\item \texttt{ltxfile.tex}.
\item \texttt{ltxfile.aux}, the auxilliary file generated by the
  compilation.
\item A PostScript file (compiled using \texttt{dvips} or some other
  driver) of the \texttt{.dvi} file generated from
  \texttt{ltxfile.tex}, or a PDF file distilled from that
  PostScript.  You do not need to include the actual \texttt{.dvi}
  file in your upload.
\item From B{\small{IB}}\TeX\ users, your bibliography (\texttt{.bib})
  file, {\it and\/} the generated file \texttt{ltxfile.bbl} created
  when you run B{\small{IB}}\TeX.
\item Any additional \texttt{.sty} and \texttt{.bst} files called by
  the source code (though, for reasons noted earlier, we {\it
    strongly\/} discourage the use of such files beyond those
  mentioned in this document).
\end{itemize}
\end{itemize}

% Your references go at the end of the main text, and before the
% figures.  For this document we've used BibTeX, the .bib file
% scibib.bib, and the .bst file Science.bst.  The package scicite.sty
% was included to format the reference numbers according to *Science*
% style.

\bibliography{scibib}

\bibliographystyle{Science}

% Following is a new environment, {scilastnote}, that's defined in the
% preamble and that allows authors to add a reference at the end of the
% list that's not signaled in the text; such references are used in
% *Science* for acknowledgments of funding, help, etc.

\begin{scilastnote}
\item We've included in the template file \texttt{scifile.tex} a new
environment, \texttt{\{scilastnote\}}, that generates a numbered final
citation without a corresponding signal in the text.  This environment
can be used to generate a final numbered reference containing
acknowledgments, sources of funding, and the like, per {\it Science\/}
style.
\end{scilastnote}

% For your review copy (i.e., the file you initially send in for
% evaluation), you can use the {figure} environment and the
% \includegraphics command to stream your figures into the text, placing
% all figures at the end.  For the final, revised manuscript for
% acceptance and production, however, PostScript or other graphics
% should not be streamed into your compliled file.  Instead, set
% captions as simple paragraphs (with a \noindent tag), setting them
% off from the rest of the text with a \clearpage as shown  below, and
% submit figures as separate files according to the Art Department's
% instructions.

\clearpage

\noindent {\bf Fig. 1.} Please do not use figure environments to set
up your figures in the final (post-peer-review) draft, do not include graphics in your
source code, and do not cite figures in the text using \LaTeX\
\verb+\ref+ commands.  Instead, simply refer to the figure numbers in
the text per {\it Science\/} style, and include the list of captions at
the end of the document, coded as ordinary paragraphs as shown in the
\texttt{scifile.tex} template file.  Your actual figure files should
be submitted separately.

\end{document}

### 2015-01-15 Paper/PaperJahrSchlich.tex
Einleitung Abschlusssatz gekürzt, FFT angepasst

\documentclass[12pt]{scrartcl}
\usepackage[ngerman, english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[color=yellow!15]{todonotes}
\usepackage{graphicx}
\usepackage[numbers, square]{natbib}

% used for line numbering
\usepackage{lineno}

% The following parameters seem to provide a reasonable page setup.
\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.
\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{\textbf{ \Large{Decentralized, autonomous sensor fault detection using neural networks}} }

\author
{Katrin Jahr, Robert Schlich$^{1}$\\
\\
\normalsize{$^{1}$Degree program “Civil Engineering” (M.Sc.)}\\
\normalsize{Bauhaus-Universität Weimar, Germany}\\
\\
\normalsize{katrin.jahr@uni-weimar.de}\\
\normalsize{robert.schlich@uni-weimar.de}
}

% Include the date command, but leave its argument blank.
\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}

The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults. 
In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
For self-detecting sensor faults, each sensor node predicts its expected sensor data and compares it to the measured sensor data. 
The predictions are computed using neural networks \textbf{with measured sensor data} of adjacent sensor nodes.
In laboratory experiments, devised to validate the proposed approach, several simulated sensor faults are detected.
These results indicate that the use of neural networks increases the dependability and the accuracy of structural health monitoring systems.

%1st Draft:
%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults. Autonomous fault detection can be obtained by comparing sensor data with data predicted by neural networks. In this paper, the design and implementation of a wireless structural health monitoring system capable of autonomous sensor fault detection using a neural network is presented. The system was validated in laboratory experiments showing its efficency.

%2nd Draft
%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults. In this paper, the design and implementation of a wireless structural health monitoring system is presented. A neural network is used for autonomous fault detection by comparing the predicted data with measured sensor data. In laboratory experiments, sensor faults were detected reliably, proving that the use of neural networks increases the reliability of structural health monitoring systems.

%3th Draft 
%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults. In this paper, the design and implementation of a wireless structural health monitoring system, capable of autonomous fault detection, is presented. A neural network is used for self-monitoring of each sensor by predicting data values and comparing these values to measured sensor data. In laboratory experiments, sensor faults were detected reliably, proving that the use of neural networks increases the reliability of structural health monitoring systems.

%4th Draft
%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults. 
%In this paper, the design and implementation of a wireless structural health monitoring system, capable of autonomous fault detection, is presented. 
%A neural network is used for self-monitoring of each sensor node by predicting the measured data with data values submitted by adjacent sensor nodes. 
%These predictions are then compared to the measured data.
%In laboratory experiments, sensor faults were detected, proving that the use of neural networks increases the reliability of structural health monitoring systems.

\end{sciabstract}

%----------------------------------------------------------------------------------------

\linenumbers % Schaltet Zeilennummerierung ein
%\modulolinenumbers[5] % nur jede 5. Zeile
\section*{Dictionary}

\begin{tabular}{|l|l|}
\hline 
Sensorknoten (SunSPOT) & sensor node \\ 
\hline 
einzelner Messsensor (Thermometer) & sensor \\ 
\hline 
Knoten im neuronalen Netz & neuron \\ 
\hline 
eine abgeschlossene Messung & measurement \\ 
\hline 
gemessene Werte & sensor data \\ 
\hline 
vorhergesagte Werte & predicted data \\ 
\hline 
durch Vorhersage erwartete Werte & expected data \\ 
\hline 
einzelner Messwert & measurement value \\ 
\hline 
\end{tabular}

\todo[inline]{"tatsächliche" / reale; virtuelle Messung?}

%----------------------------------------------------------------------------------------

\section*{Introduction}

%\begin{itemize}
%\item Unfälle gefährden Menschenleben und sind teuer
%\item SHMS können Unfälle verhindern
%\item kabellose Sensoren sind besser
%\item Sensoren sind fehleranfällig
%\item Lösung: autonome Fehlererkennung 
%\item Lösungsansätze mit Literaturbelegen
%\item Ansatz: Neuronales Netz
%\item Überblick, Vorteile
%\item Literatur zu neuronalen Netzen
%\item Überblick über geleistete Arbeit
%\end{itemize}
%
%\newpage
%
%Civil engineering structures are designed to resist extensive external impact over several decades. 
%Over time, the structural stability can decrease to an hazardous level. 
%To prevent structural failure, demolition and replacement of the building or rehabilitation and repair works have to be considered.
%From the outside, it is hard to determine whether building structures are damaged.
%To decide whether cost- and scheduling-intensive strengthening actions are required and to predict failure in advance, structural health monitoring (SHM) is used.
%
%\cite[4]{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure". 
%The obtained sensor data is "collected, analyzed and stored".
%The structure is observed over long periods of time - up to several years - using data obtained by sensor nodes attached to the \textbf{construction}.
%Via analyzing the measured data, such as acceleration or strain, changes in material and geometric behaviour can be recognized early on.
%
%
%((((
%
%Traditionally, the used sensor nodes are wired with optical fiber cables. 
%However, using wired SHM systems has several disadvantages: 
%For monitoring large constructions like bridges or high-rise buildings, many sensor nodes and therefore a big amount of wire is needed, raising material and personnel expenses.
%Optimal positions for sensor nodes can be impossible to wire and clamps can damage the facade. 
%Visible wires compromise the aesthetic view of a construction and draw interest to vandalism.
%
%))))
%
%Wireless SHM systems are composed of a several sensor nodes, a base station and a personal computer, that communicate by radio. 
%Therefore, the named disadvantages of cables do not apply.
%Wireless SHM, however, bear some challenges that have to be considered. 
%Electricity supply is constrained to battery life or expensive alternative power suppliers.
%The transmitter signal strength has to be optimized to span long distances without disturbing nearby electronic devices or humans.
%Data security may have to be ensured by encrypting the transmitted data and eliminating the possibility of data manipulation.In this case
%When using several different sensor nodes, the timestamps of single measurement values can differ. 
%To assure the relatability of the sensor data, it is therefore necessary to synchronize all sensor timers. 
%
%Over time, sensors can break and become faulty. Faults can be due to either sudden failure or gradually loss of accuracy.
%To ensure the dependability and the accuracy of the structural health monitoring system, these sensor faults must be detected and eliminated in real time. 
%
%To autonomously detect sensor faults, several different approaches have been tested.
%
%...
%
%\todo[inline]{Literatur zu nicht neuronalen netzwerken und deren Probleme}
%
%\todo[inline]{neuronale Netze. später Abgleich mit Abschnitt.}
%
%This paper shows a wireless structural health monitoring system with autonomous fault detection. 
%At first, the SHM system is designed and implemented. 
%Next, a neural network is implemented in each sensor and trained to detect faulty behaviour. 
%The system is then tested in laboratory experiments. 
%Finally, the experimental results are analyzed and future research directions proposed.

Civil engineering structures are exposed to external impacts over several decades. 
To ensure the structural stability and evaluate the conditions of civil engineering structures, structural health monitoring (SHM) systems can be installed.
\cite[4]{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure". 
The obtained sensor data is "collected, analyzed and stored" over long periods of time.
Analysis of the sensor data can reveal abnormal changes in material and geometric behaviour early on.

Traditionally, the sensor nodes are wired with cables and connected to computer systems.
However, using wired SHM systems has several disadvantages, including expensive wiring, high personnal costs and inaccessibility of optimal positions with wiresIn this case.
In wireless SHM systems, the sensor nodes communicate with a base station via radio, ruling out wiring-specific problems.
When using wireless SHM systems, one has to consider battery life and sensor node \textit{sychronization}.
\todo{wird später darauf eingegangen? Sonst raus.}

Over time, sensors can break and become faulty. Failures can either occur suddenly, or be due to gradual loss of accuracy.
To ensure the dependability and the accuracy of the SHM system, these sensor faults must be detected and isolated in real time. 
A simple approach to fault detection is the installation of physically redundant sensors.
Faulty sensors can easily be identified through the deviation of their measurements from the measurements of other sensors.
Physical redundancy, although efficient at sensor fault detection, has the disadvantage of increased installation and maintenance costs. \cite{SmarslyDA}
A technologically more advanced approach, called analytical redundancy, applies knowledge of the structure and the SHM system itself to compute virtual sensor measurements for each sensor. 
For example, a FEM model could be used in combination with data from adjacent sensor nodes to calculate a virtual measurement of one sensor.
If analytical redundancy is implemented correctly, virtual measurements always simulate a correct working sensor; therefore actual measurements deviating from their virtual counterparts indicate faulty sensors.\cite{SmarslyDA} \todo{andere Quelle}

In this study, analytical redundancy is realized by artificial neural networks.
Artificial neural networks consist of interconnected data processing elements called neurons. 
Neurons are grouped in different layers; they consist of one input layer, a number of hidden layers and one output layer.
Artificial neural networks have the distinct ability of learning by adjusting the weights of the inter-neuron connections, until a set of given input variables results in the desired output variables.
If implemented correctly, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy. \todo{Zitat Li}

This paper is organized as follows:
First, a wireless structural health monitoring system is designed and implemented. 
Next, a neural network is implemented in each sensor node and trained to predict the measurements of said node in order to detect faults. 
The system is then tested in laboratory experiments. 
Finally, the experimental results are analyzed and future research directions proposed.

%----------------------------------------------------------------------------------------

\newpage

\section*{Design and implementation of the wireless structural health monitoring system}

\textbf{1. paragraph: SHM general}

\begin{itemize}
\item wireless SHM system consits of nodes and base station
\item nodes collect data, and process data
\item processed data is sent to base station and personal computer
\item data is stored in database
\end{itemize}

According to \cite[5]{BisbySHM}, a structural health monitoring system commonly consists of the following components:
1. data acquisition,
2. data transmission, 
3. data processing,
4. data storage,
5. diagnostics and 
6. information retrieval.

Task 1. to 3. are implemented on the building site, tasks 4. to 6. at a workstation. 
A wireless SHM system installed on the building site consists of sensor nodes attached to the construction, a base station and a local computer. 
The base station, which can be an external device or a computer program, links the individual sensor nodes with each other and the local computer.
To each sensor node, several sensors measuring different physical quantities may be connected. 
During a measurement, the sensor nodes acquire sensor data and may perform initial processing.
The processed data is then transmitted via radio to the base station and, from there, to the local computer. 
At the computer, the data gets stored accessible and save, e. g. in a database for analysis and diagnosis.

\todo[inline]{vielleicht eine Abbildung?}

\textbf{2. paragraph: SunSPOTs}

\begin{itemize}
\item SunSPOTs...
\item several sensors: acc, temp, light...
\item wireless communication
\item base station
\end{itemize}

\textbf{3. paragraph: UML / Java implementaion}

%According to [3], a structural health monitoring system basically consists of the following components:
%1. data acquisition, 2. data communication, 3. data processing, 4. data storage, 5. diagnostics and 6. data retrieval.
%
%
%
%For autonomous fault detection in wireless SHM systems, an integrated fault detection and synchronization strategy is implemented into a wireless SHM system. The SHM system consists of wireless sensor nodes and a base station. The sensor nodes are installed on the monitored structure and the base station links the wireless SHM system and a local computer.
%
%
%In the present study, wireless sensor nodes of the type “Oracle Sun SPOTs” are used. The sensor nodes are Java-programmable enabling embedded analyses and the sensor nodes include a 3-dimensional accelerometer, temperature and other sensors (Oracle Lab, 2014). In this study, acceleration measurements taken from the structure are used for analyzing the structural condition. Furthermore, the radios of the sensor nodes ensure wireless communications between the sensor nodes as well as between the sensor nodes and the base station in order to transmit monitoring data or control commands. The base station is able to receive monitoring data measured by the sensor nodes, and the base station can store the data in a database installed on the local computer. The main Java classes are shown in an Unified Modelling Language diagram (UML diagram), which is a standard way to visualize the implementation of a application. The following UML diagrams (cp. Figure 1 and Figure 2) show just the main classes embedded into the base station and the sensor nodes. In the following subsection, the implementation of the main components of the fault detection strategy, including synchronization and fault detection, are shown.

\newpage

%----------------------------------------------------------------------------------------

\section*{Implementation and training of neural networks}

\textbf{1. paragraph: neural networks general}

\todo[inline]{Die Vorteile von neuronalen Netzwerken müssen m.E. noch in der Einleitung heruasgearbeitet werden}

Artificial neural networks were developed, driven by the desire to enable computer systems to learn autonomously.
This development is based on the attempt to emulate the data processing mechanics of the human brain. %(and is therefore inherently interwoven with advances in neuroscience)
In an artificial neural network \emph{neurons} are grouped in different \emph{layers}, as depicted in figure~\ref{fig:neuralnetwork}. 
Each neuron sends data exclusively to neurons from neighbouring layers through weighted \emph{connections}.
For each neuron a \emph{propagation function} processes incoming data, in consideration of the connection weights, to the net input.
An \emph{activation function} then processes the net input to calculate the activation of the neuron, which subsequently determines its output. \cite{KrieselNN}

A neural network learns, most commonly, by adjusting its connection weights until it recreates the results of a desired transfer function with sufficient accuracy.
In our case...
\todo[inline]{Weiteres zur Gewichtsveränderung in der Lernphase später im Abschnitt}

\begin{figure}[h]
    \centering
    \includegraphics{figures/neuralnetworks.pdf}
    \caption{Schematic drawing of an artificial neural network with three layers}
    \label{fig:neuralnetwork}
\end{figure}
\todo[inline]{Grafik kann nacher noch an unser neuronales Netz angepasst werden}

%\begin{itemize}
%\item neurons
%\item layer
%\item weights
%\item activation function / identity function
%\end{itemize}

\textbf{2. paragraph: proposed nn}

\begin{itemize}
\item SNIPE? implementation in java?
\item activation function / identity function
\end{itemize}

\textbf{3. paragraph: learning general}

\begin{itemize}
\item set with test values
\item iterative weight adjustment
\end{itemize}

\textbf{4. paragraph?: learning, proposed}

%----------------------------------------------------------------------------------------

\newpage
\section*{Laboratory experiments}

\textbf{1. paragraph: test setting}

\begin{itemize}
\item test structure, picture
\item sensor installation
\item type of excitation, acceleration is measured

\item learning phase
\end{itemize}

\textbf{2. paragraph: data measurement and processing}

\begin{itemize}
\item excitation
\item sensors start measuring
\item sensor nodes \textbf{perform fft} - what is fft
\item sensor nodes exchange frequencies
\item neural networks check integrity
\item if no faults detected, sensor nodes send data to base station
\end{itemize}

\textbf{3. paragraph: sensor fault}

\begin{itemize}
\item simulation of sensor fault
\item neural networks check integrity -> error!
\item sensor node sends alert to base station
\end{itemize}

%----------------------------------------------------------------------------------------

\section*{Discussion of the experimental results}

%----------------------------------------------------------------------------------------

\section*{Summary and conclusions}

%----------------------------------------------------------------------------------------

\bibliographystyle{unsrtnat}
\bibliography{literature}

\end{document}

### 2015-01-16 Paper/PaperJahrSchlich.tex

@@
 Faulty sensors can easily be identified through the deviation of their measurements from the measurements of other sensors.
-Physical redundancy, although efficient at sensor fault detection, has the disadvantage of increased installation and maintenance costs. \cite{SmarslyDA}
+Physical redundancy, although efficient at sensor fault detection, has the disadvantage of increased installation and maintenance costs. 
 A technologically more advanced approach, called analytical redundancy, applies knowledge of the structure and the SHM system itself to compute virtual sensor measurements for each sensor. 
 For example, a FEM model could be used in combination with data from adjacent sensor nodes to calculate a virtual measurement of one sensor.
-If analytical redundancy is implemented correctly, virtual measurements always simulate a correct working sensor; therefore actual measurements deviating from their virtual counterparts indicate faulty sensors.\cite{SmarslyDA} \todo{andere Quelle}
+If analytical redundancy is implemented correctly, virtual measurements always simulate a correct working sensor; therefore actual measurements deviating from their virtual counterparts indicate faulty sensors.\citep{Smarsly2014}
 
@@
 Artificial neural networks have the distinct ability of learning by adjusting the weights of the inter-neuron connections, until a set of given input variables results in the desired output variables.
-If implemented correctly, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy. \todo{Zitat Li}
+If implemented correctly, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy \citep{Li2011}.
 

### 2015-01-17 Paper/PaperJahrSchlich.tex
cleaned up a little
Überarbeitung Abschnitt SHM, UML/Java fehlt

@@
 \newenvironment{sciabstract}{%
-\begin{quote} \bf}
+\begin{quote} \bfseries}
 {\end{quote}}
@@
 
-%1st Draft:
-%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults.
-Autonomous fault detection can be obtained by comparing sensor data with data predicted by neural networks.
-In this paper, the design and implementation of a wireless structural health monitoring system capable of autonomous sensor fault detection using a neural network is presented.
-The system was validated in laboratory experiments showing its efficency.
-
-%2nd Draft
-%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults.
-In this paper, the design and implementation of a wireless structural health monitoring system is presented.
-A neural network is used for autonomous fault detection by comparing the predicted data with measured sensor data.
-In laboratory experiments, sensor faults were detected reliably, proving that the use of neural networks increases the reliability of structural health monitoring systems.
-
-%3th Draft 
-%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults.
-In this paper, the design and implementation of a wireless structural health monitoring system, capable of autonomous fault detection, is presented.
-A neural network is used for self-monitoring of each sensor by predicting data values and comparing these values to measured sensor data.
-In laboratory experiments, sensor faults were detected reliably, proving that the use of neural networks increases the reliability of structural health monitoring systems.
-
-%4th Draft
-%The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults. 
-%In this paper, the design and implementation of a wireless structural health monitoring system, capable of autonomous fault detection, is presented. 
-%A neural network is used for self-monitoring of each sensor node by predicting the measured data with data values submitted by adjacent sensor nodes. 
-%These predictions are then compared to the measured data.
-%In laboratory experiments, sensor faults were detected, proving that the use of neural networks increases the reliability of structural health monitoring systems.
-
 \end{sciabstract}
@@
 \hline 
-eine abgeschlossene Messung & measurement \\ 
+eine abgeschlossene Messungreihe & test run \\ 
 \hline 
@@
 \hline 
-einzelner Messwert & measurement value \\ 
+einzelner Messwert & measurement \\ 
 \hline 
+Test & laboratory experiments\\ 
+\hline
+tatsächliche, nicht virtuelle Messung & actual measurement \\
+\hline
+Messaufbau & test setup \\
+\hline
+
 \end{tabular}
-
-\todo[inline]{"tatsächliche" / reale; virtuelle Messung?}
 
@@
 \section*{Introduction}
-
-%\begin{itemize}
-%\item Unfälle gefährden Menschenleben und sind teuer
-%\item SHMS können Unfälle verhindern
-%\item kabellose Sensoren sind besser
-%\item Sensoren sind fehleranfällig
-%\item Lösung: autonome Fehlererkennung 
-%\item Lösungsansätze mit Literaturbelegen
-%\item Ansatz: Neuronales Netz
-%\item Überblick, Vorteile
-%\item Literatur zu neuronalen Netzen
-%\item Überblick über geleistete Arbeit
-%\end{itemize}
-%
-%\newpage
-%
-%Civil engineering structures are designed to resist extensive external impact over several decades. 
-%Over time, the structural stability can decrease to an hazardous level. 
-%To prevent structural failure, demolition and replacement of the building or rehabilitation and repair works have to be considered.
-%From the outside, it is hard to determine whether building structures are damaged.
-%To decide whether cost- and scheduling-intensive strengthening actions are required and to predict failure in advance, structural health monitoring (SHM) is used.
-%
-%\cite[4]{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure". 
-%The obtained sensor data is "collected, analyzed and stored".
-%The structure is observed over long periods of time - up to several years - using data obtained by sensor nodes attached to the \textbf{construction}.
-%Via analyzing the measured data, such as acceleration or strain, changes in material and geometric behaviour can be recognized early on.
-%
-%
-%((((
-%
-%Traditionally, the used sensor nodes are wired with optical fiber cables. 
-%However, using wired SHM systems has several disadvantages: 
-%For monitoring large constructions like bridges or high-rise buildings, many sensor nodes and therefore a big amount of wire is needed, raising material and personnel expenses.
-%Optimal positions for sensor nodes can be impossible to wire and clamps can damage the facade. 
-%Visible wires compromise the aesthetic view of a construction and draw interest to vandalism.
-%
-%))))
-%
-%Wireless SHM systems are composed of a several sensor nodes, a base station and a personal computer, that communicate by radio. 
-%Therefore, the named disadvantages of cables do not apply.
-%Wireless SHM, however, bear some challenges that have to be considered. 
-%Electricity supply is constrained to battery life or expensive alternative power suppliers.
-%The transmitter signal strength has to be optimized to span long distances without disturbing nearby electronic devices or humans.
-%Data security may have to be ensured by encrypting the transmitted data and eliminating the possibility of data manipulation.In this case
-%When using several different sensor nodes, the timestamps of single measurement values can differ. 
-%To assure the relatability of the sensor data, it is therefore necessary to synchronize all sensor timers. 
-%
-%Over time, sensors can break and become faulty. Faults can be due to either sudden failure or gradually loss of accuracy.
-%To ensure the dependability and the accuracy of the structural health monitoring system, these sensor faults must be detected and eliminated in real time. 
-%
-%To autonomously detect sensor faults, several different approaches have been tested.
-%
-%...
-%
-%\todo[inline]{Literatur zu nicht neuronalen netzwerken und deren Probleme}
-%
-%\todo[inline]{neuronale Netze. später Abgleich mit Abschnitt.}
-%
-%This paper shows a wireless structural health monitoring system with autonomous fault detection. 
-%At first, the SHM system is designed and implemented. 
-%Next, a neural network is implemented in each sensor and trained to detect faulty behaviour. 
-%The system is then tested in laboratory experiments. 
-%Finally, the experimental results are analyzed and future research directions proposed.
 
@@
 \section*{Design and implementation of the wireless structural health monitoring system}
-
-\textbf{1. paragraph: SHM general}
-
-\begin{itemize}
-\item wireless SHM system consits of nodes and base station
-\item nodes collect data, and process data
-\item processed data is sent to base station and personal computer
-\item data is stored in database
-\end{itemize}
+In the following section, a brief descriptions of SHM systems general is given. Furthermore, the used wireless SHM system is introduced and the Java implementation is described.
 
@@
 
-Task 1. to 3. are implemented on the building site, tasks 4. to 6. at a workstation. 
-A wireless SHM system installed on the building site consists of sensor nodes attached to the construction, a base station and a local computer. 
+Tasks 1 to 3 are usually implemented on site, while tasks 4 to 6 are usually implemented on computer systems off site.
+A wireless SHM system installed on the building site consists of sensor nodes attached to the construction, a base station, and a local computer. 
 The base station, which can be an external device or a computer program, links the individual sensor nodes with each other and the local computer.
 To each sensor node, several sensors measuring different physical quantities may be connected. 
-During a measurement, the sensor nodes acquire sensor data and may perform initial processing.
-The processed data is then transmitted via radio to the base station and, from there, to the local computer. 
-At the computer, the data gets stored accessible and save, e. g. in a database for analysis and diagnosis.
+In this project, the acceleration is measured using wireless sensor nodes and a basestation of the type "Oracle Sun SPOT". 
+The 3-axis digital output accelerometer with sensitivity ranging between $ \pm $ 2G and $\pm$ 8G has a maximum sample rate of 125Hz. \cite[9]{eDemo2010}
 
-\todo[inline]{vielleicht eine Abbildung?}
+In operation, the sensor nodes acquire sensor data and may perform initial processing.
+The processed data is then transmitted via radio to the local computer, where it is stored for analysis and diagnosis.
+In this project, the data is transmitted from the sensor nodes to the basestation by radio, and from the basestation to the local computer via USB cable.
+On the computer, the data is stored in a MySQL database.
 
-\textbf{2. paragraph: SunSPOTs}
-
-\begin{itemize}
-\item SunSPOTs...
-\item several sensors: acc, temp, light...
-\item wireless communication
-\item base station
-\end{itemize}
+\todo[inline]{vielleicht eine Abbildung? SunSPOTs? Aufbau SHM?}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-\textbf{1. paragraph: test setting}
+\textbf{1. paragraph: test setup}
 
@@
 
-\item learning phase
 \end{itemize}
@@
 \begin{itemize}
+\item learning phase
 \item excitation
@@
 
-\textbf{3. paragraph: sensor fault}
+\textbf{3. paragraph: sensor fault detection}
 

### 2015-01-19 Paper/PaperJahrSchlich.tex
Kapitel Laboratory Experiments angefangen - teststruktur beschrieben.

@@
-\documentclass[12pt]{scrartcl}
+\documentclass[12pt,a4paper]{scrartcl}
 \usepackage[ngerman, english]{babel}
@@
 % used for line numbering
-\usepackage{lineno}
+%\usepackage{lineno}
 
@@
 
-\baselineskip24pt
+\baselineskip32pt
 
@@
 
-\linenumbers % Schaltet Zeilennummerierung ein
+%\linenumbers % Schaltet Zeilennummerierung ein
 %\modulolinenumbers[5] % nur jede 5. Zeile
@@
 
-\textbf{1. paragraph: neural networks general}
-
-\todo[inline]{Die Vorteile von neuronalen Netzwerken müssen m.E. noch in der Einleitung heruasgearbeitet werden}
-
-Artificial neural networks were developed, driven by the desire to enable computer systems to learn autonomously.
-This development is based on the attempt to emulate the data processing mechanics of the human brain. %(and is therefore inherently interwoven with advances in neuroscience)
-In an artificial neural network \emph{neurons} are grouped in different \emph{layers}, as depicted in figure~\ref{fig:neuralnetwork}. 
-Each neuron sends data exclusively to neurons from neighbouring layers through weighted \emph{connections}.
-For each neuron a \emph{propagation function} processes incoming data, in consideration of the connection weights, to the net input.
-An \emph{activation function} then processes the net input to calculate the activation of the neuron, which subsequently determines its output. \cite{KrieselNN}
-
-A neural network learns, most commonly, by adjusting its connection weights until it recreates the results of a desired transfer function with sufficient accuracy.
-In our case...
-\todo[inline]{Weiteres zur Gewichtsveränderung in der Lernphase später im Abschnitt}
-
+Neural networks are implemented into the sensor nodes using SNIPE\footnote{}, a open source Java Library.
 \begin{figure}[ht]
     \centering
-    \includegraphics{figures/neuralnetworks.pdf}
+    \includegraphics{figures/neuralnetwork.pdf}
     \caption{Schematic drawing of an artificial neural network with three layers}
@@
 
-\textbf{1. paragraph: test setup}
+%\begin{itemize}
+%\item test structure
+%\item sensor installation
+%\item type of excitation, acceleration is measured
+%\end{itemize}
 
-\begin{itemize}
-\item test structure, picture
-\item sensor installation
-\item type of excitation, acceleration is measured
+To validate the proposed approach, the SHM system was installed on a test structure in order to conduct laboratory experiments.
+The test structure consists of four metal sheets with an area of 20~cm by 60~cm and 4~mm of thickness.
+Those metal sheets are mounted on threaded rods with a vertical clearance of approximately 40~cm.
+At the bottom, the threaded rods are fixed into a block of solid wood with an area of 40~cm by 60~cm and a height of 50~cm.
+The SHM system is installed onto the test structure by strapping a SunSPOT to each of the metal sheets.
+The finished laboratory setup is shown in figure~\ref{fig:teststructure}.
+\todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
+The structure is excited by pulling one of the metal sheets a few centimeters to the side and releasing it.
+This method of excitation ensures, that the structure vibrates in its Eigenfrequency with little interferences.
 
-\end{itemize}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=0.3]{figures/teststructure.jpg}
+    \caption{Picture of the laboratory setup}
+    \label{fig:teststructure}
+\end{figure}
 
@@
 
-\section*{Discussion of the experimental results}
+%\section*{Discussion of the experimental results}
 
@@
 
+Everything went better than expected!
 %----------------------------------------------------------------------------------------

### 2015-01-19 Paper/PaperJahrSchlich.tex
Zitatdarstellung geändert
Änderungen Introduction eingefügt, SHM und Lab Ex gesichtet

@@
 \usepackage{graphicx}
-\usepackage[numbers, square]{natbib}
+\usepackage[authoryear, square]{natbib}
 
@@
 
-\title{\textbf{ \Large{Decentralized, autonomous sensor fault detection using neural networks}} }
+\title{\textbf{ \Large{Decentralized autonomous sensor fault detection using neural networks}} }
 
 \author
-{Katrin Jahr, Robert Schlich$^{1}$\\
+{Katrin Jahr, Robert Schlich\\
 \\
-\normalsize{$^{1}$Degree program “Civil Engineering” (M.Sc.)}\\
+\normalsize{Degree program “Civil Engineering” (M.Sc.)}\\
 \normalsize{Bauhaus-Universität Weimar, Germany}\\
@@
 
-\baselineskip32pt
+\baselineskip24pt
 
@@
 
-The dependability and the accuracy of structural health monitoring systems are significantly affected by sensor faults. 
+The dependability and the accuracy of structural health monitoring systems can be affected significantly by sensor faults. 
 In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
-For self-detecting sensor faults, each sensor node predicts its expected sensor data and compares it to the measured sensor data. 
-The predictions are computed using neural networks \textbf{with measured sensor data} of adjacent sensor nodes.
+For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
+The predictions are computed using neural networks based on measured sensor data of adjacent sensor nodes.
 In laboratory experiments, devised to validate the proposed approach, several simulated sensor faults are detected.
-These results indicate that the use of neural networks increases the dependability and the accuracy of structural health monitoring systems.
+These results indicate that the use of neural networks for fault detection increases the dependability and the accuracy of structural health monitoring systems.
 
@@
 
-Civil engineering structures are exposed to external impacts over several decades. 
-To ensure the structural stability and evaluate the conditions of civil engineering structures, structural health monitoring (SHM) systems can be installed.
+Civil engineering structures are exposed to various external impacts during their lifetime. 
+Structural health monitoring (SHM) systems can be deployed to evaluate the conditions and to ensure the structural stability of civil engineering structures.
 \cite[4]{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure".
-The obtained sensor data is "collected, analyzed and stored" over long periods of time.
+The obtained sensor data is collected by sensor nodes, and then analyzed and stored on a computer system over long periods of time. 
 Analysis of the sensor data can reveal abnormal changes in material and geometric behaviour early on.
@@
 Traditionally, the sensor nodes are wired with cables and connected to computer systems.
-However, using wired SHM systems has several disadvantages, including expensive wiring, high personnal costs and inaccessibility of optimal positions with wiresIn this case.
-In wireless SHM systems, the sensor nodes communicate with a base station via radio, ruling out wiring-specific problems.
-When using wireless SHM systems, one has to consider battery life and sensor node \textit{sychronization}.
-\todo{wird später darauf eingegangen? Sonst raus.}
+However, using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs and inaccessibility of optimal sensor location with wires.
+In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
 
-Over time, sensors can break and become faulty. Failures can either occur suddenly, or be due to gradual loss of accuracy.
-To ensure the dependability and the accuracy of the SHM system, these sensor faults must be detected and isolated in real time. 
-A simple approach to fault detection is the installation of physically redundant sensors.
-Faulty sensors can easily be identified through the deviation of their measurements from the measurements of other sensors.
-Physical redundancy, although efficient at sensor fault detection, has the disadvantage of increased installation and maintenance costs. 
-A technologically more advanced approach, called analytical redundancy, applies knowledge of the structure and the SHM system itself to compute virtual sensor measurements for each sensor. 
-For example, a FEM model could be used in combination with data from adjacent sensor nodes to calculate a virtual measurement of one sensor.
-If analytical redundancy is implemented correctly, virtual measurements always simulate a correct working sensor; therefore actual measurements deviating from their virtual counterparts indicate faulty sensors.\citep{Smarsly2014}
+Over their lifetime, sensors can become inaccurate, faulty, or may even break.
+To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
+The traditional approach to fault detection is the installation of physically redundant sensors.
+Faulty sensors can easily be identified through the deviation of their measurements from the measurements of correlated sensors.
+Physical redundancy, although efficient for sensor fault detection, shows the disadvantage of increased installation and maintenance costs. 
+Representing a more efficient approach, analytical redundancy applies knowledge of the structure and the SHM system itself to compute virtual sensor measurements for each sensor. 
+For example, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor.
+\citep{Smarsly2014}
 
-In this study, analytical redundancy is realized by artificial neural networks.
-Artificial neural networks consist of interconnected data processing elements called neurons. 
-Neurons are grouped in different layers; they consist of one input layer, a number of hidden layers and one output layer.
-Artificial neural networks have the distinct ability of learning by adjusting the weights of the inter-neuron connections, until a set of given input variables results in the desired output variables.
-If implemented correctly, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy \citep{Li2011}.
+In this study, analytical redundancy is implemented into wireless sensor nodes based on artificial neural networks.
+Artificial neural networks essentially consist of interconnected data processing units called neurons. 
+Neurons are grouped in different layers; usually one input layer, a number of hidden layers and one output layer.
+Artificial neural networks have the distinct ability of learning, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy \citep{Li2011}.
 
@@
 First, a wireless structural health monitoring system is designed and implemented. 
-Next, a neural network is implemented in each sensor node and trained to predict the measurements of said node in order to detect faults. 
-The system is then tested in laboratory experiments. 
-Finally, the experimental results are analyzed and future research directions proposed.
+Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of said node in order to detect sensor faults in a decentralized manner. 
+Then, the system is tested in laboratory experiments. 
+Finally, the experimental results are discussed and future research directions are proposed.
 
@@
 \section*{Design and implementation of the wireless structural health monitoring system}
-In the following section, a brief descriptions of SHM systems general is given. Furthermore, the used wireless SHM system is introduced and the Java implementation is described.
+In the following section, a brief description of SHM systems in general is given. Furthermore, the used wireless SHM system is introduced and the Java implementation is described.
 
@@
 
-To validate the proposed approach, the SHM system was installed on a test structure in order to conduct laboratory experiments.
-The test structure consists of four metal sheets with an area of 20~cm by 60~cm and 4~mm of thickness.
-Those metal sheets are mounted on threaded rods with a vertical clearance of approximately 40~cm.
-At the bottom, the threaded rods are fixed into a block of solid wood with an area of 40~cm by 60~cm and a height of 50~cm.
-The SHM system is installed onto the test structure by strapping a SunSPOT to each of the metal sheets.
+To validate the proposed approach, the SHM system is installed on a test structure in order to conduct laboratory experiments.
+The test structure consists of four metal sheets with an area of 20\,cm by 60\,cm and 4\,mm of thickness.
+Those metal sheets are mounted on threaded rods with a vertical clearance of approximately 40\,cm.
+At the bottom, the threaded rods are fixed into a block of solid wood with an area of 40\,cm by 60\,cm and a height of 50\,cm.
+The SHM system is installed onto the test structure by fastening a SunSPOT to each of the metal sheets.
 The finished laboratory setup is shown in figure~\ref{fig:teststructure}.
 \todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
-The structure is excited by pulling one of the metal sheets a few centimeters to the side and releasing it.
+The structure is excited by displacing one of the metal sheets to the side and releasing it.
 This method of excitation ensures, that the structure vibrates in its Eigenfrequency with little interferences.
@@
 
-\section*{Summary and conclusions}
+\section*{Summary}
 
-Everything went better than expected!
+This paper presents a decentralized autonomous sensor fault detection strategy for structural health monitoring systems using neural networks. 
+By implementing the neural networks into the sensor nodes, sensor faults can be detected by individual sensor nodes.
+
 %----------------------------------------------------------------------------------------

### 2015-01-20 Marcel/Code/NeuralNetworkProject/src/desktop/src/org/sunspotworld/demo/ProtocolReader.java
Projekt 2013

@@ -1,0 +1,169 @@
+package desktop.src.org.sunspotworld.demo;
+
+import java.io.BufferedWriter;
+import java.io.DataInputStream;
+import java.io.DataOutputStream;
+import java.io.File;
+import java.io.FileWriter;
+import java.io.IOException;
+import java.io.PrintWriter;
+import java.text.DecimalFormat;
+import java.text.DecimalFormatSymbols;
+import java.util.ArrayList;
+
+import javax.microedition.io.Connector;
+
+import com.sun.spot.io.j2me.radiostream.RadiostreamConnection;
+import com.sun.spot.util.Utils;
+
+public class ProtocolReader {
+
+    //Stream-Connections
+    private RadiostreamConnection listenOn;
+    private RadiostreamConnection sendOn;
+    private DataInputStream listenOnInputStream;
+    private DataOutputStream sendOnOutputStream;
+    
+    //parameter
+    private int length;
+    private ArrayList<String> addresses;
+    private int defaultPort;
+    private int sample_period;
+    
+    private String directory = System.getProperty("user.home") + "/Documents/SensorValues";
+        
+    
+    /**
+     * Sets all important parameters to create the protocol file.
+     * 
+     * @param length of the data array
+     * @param addresses of the different nodes
+     * @param defaultPort for the radiostream connection
+     * @param sample_period of the measurement
+     */
+    public ProtocolReader(int length, ArrayList<String> addresses, int defaultPort, int sample_period){
+    	
+    	//werte übergeben
+    	this.length = length;
+    	this.addresses = addresses;
+    	this.defaultPort = defaultPort;
+    	this.sample_period = sample_period;
+    	
+    }//Ende ProtocolReader()
+    
+    /**
+     * Opens radiostream connections for sending an listening to a defined spot.
+     * 
+     * @param index
+     * @throws IOException
+     */
+    private void openStreams(int index) throws IOException{
+
+    	listenOn = (RadiostreamConnection) Connector.open("radiostream://" + addresses.get(index) + ":" + defaultPort);
+    	listenOnInputStream = listenOn.openDataInputStream();
+    	
+    	sendOn = (RadiostreamConnection) Connector.open("radiostream://" + addresses.get(index) + ":" + (defaultPort + index + 1));
+    	sendOnOutputStream = sendOn.openDataOutputStream();
+    	
+    }//Ende openStreams()
+    
+    
+    /**
[…]

### 2015-01-20 Marcel/Code/NeuralNetworkProject/src/spot/src/org/sunspotworld/demo/neuralnetwork/HiddenUnitCalculator.java

@@ -1,0 +1,31 @@
+package spot.src.org.sunspotworld.demo.neuralnetwork;
+
+
+public class HiddenUnitCalculator {
+
+	
+	/**
+	 * Calculates the number of hidden units for the neural network.
+	 * Using the formula:
+	 * <p>{@code ( (2*(sensor_count+1)/3) + (Math.sqrt(sensor_count)) ) / 2}
+	 * <p>If the sensor count is less than two, the hidden units will be 2.
+	 * 
+	 * @param sensor_count count of all nodes in the network, including the performing one itself
+	 * @return number of hidden units
+	 */
+	public static int getHiddenUnits(int sensor_count){
+		
+		int hidden_units = sensor_count - 1;
+		
+		hidden_units = (int)( (2*(sensor_count+1)/3) + (Math.sqrt(sensor_count)) ) / 2;
+		
+		if (hidden_units<2){
+			hidden_units = 2;
+		}
+		
+		return hidden_units;
+		
+	}//Ende getHiddenUnits()
+	
+	
+}

### 2015-01-20 Marcel/Module Neuronales Netz/ArrayOfArray.java

@@ -1,0 +1,9 @@
+package spot.src.org.sunspotworld.demo;
+
+public class ArrayOfArray {
+
+	double[][] feld1, feld2;
+	
+	
+	
+}

### 2015-01-20 Marcel/Module Synchronisation/onSPot/SyncStart.java

@@ -1,0 +1,59 @@
+package spot.src.org.sunspotworld.demo;
+
+import javax.microedition.io.Datagram;
+
+import com.sun.spot.io.j2me.radiogram.RadiogramConnection;
+import com.sun.spot.resources.transducers.ITriColorLED;
+import com.sun.spot.resources.transducers.LEDColor;
+
+/* Modul SyncStart wartet bis ein Signal (boolean) vom host gegeben wird 
+ * zur Benutzung für Testphase des neuronalen Netzes 
+ */
+
+public class SyncStart {
+	
+	private RadiogramConnection rCon;
+	private Datagram dg;
+	private ITriColorLED led;
+	
+	//Standardkonstruktor
+	public SyncStart(RadiogramConnection rCon, Datagram dg, ITriColorLED led){
+		
+		//Übergabe der Verbindung (RadiogramConnection, Datagram, LED)
+		this.rCon = rCon;
+		this.dg = dg;
+		this.led = led;
+		
+	}//Ende SyncStart
+	
+	
+	//Auf Startschuss warten
+	public void waitForSignal(){
+		
+		//Signal wartephase
+		led.setColor(LEDColor.ORANGE);
+		led.setOn();
+		
+		//Kontrollvariable
+		boolean ready;
+		
+		try{
+			
+    		while(true){
+    			dg.reset();					//datagram leeren
+    			rCon.receive(dg);			//datagram empfangen
+		       	ready = dg.readBoolean();	//boolean lesen
+		       	if(ready==true)				//kontrolle (Synchronisation bei ready == true)
+		       		break;
+    		 }
+    		
+		} catch(Exception e){
+			e.printStackTrace();
+		}
+		
+		//Signal Synchronisation erfolgreich
+		led.setColor(LEDColor.GREEN);
+		
+	}//Ende waitForSignal()
+
+}

### 2015-01-20 Marcel/NNProject/src/spot/src/org/sunspotworld/demo/neuralnetwork/HiddenUnitCalculator.java

@@ -1,0 +1,20 @@
+package spot.src.org.sunspotworld.demo.neuralnetwork;
+
+
+public class HiddenUnitCalculator {
+
+	public static int getHiddenUnits(int sensor_count){
+		
+		int hidden_units = sensor_count - 1;
+		
+		hidden_units = (int)( (2*(sensor_count+1)/3) + (Math.sqrt(sensor_count)) ) / 2;
+		
+		if (hidden_units<2)
+			hidden_units=2;
+		
+		return hidden_units;
+		
+	}
+	
+	
+}

### 2015-01-20 Marcel/NNProject/src/spot/src/org/sunspotworld/demo/spotfinder/SpotIdentifier.java

@@ -1,0 +1,291 @@
+package spot.src.org.sunspotworld.demo.spotfinder;
+
+import java.io.DataInputStream;
+import java.io.DataOutputStream;
+import java.io.IOException;
+import java.util.Random;
+
+import javax.microedition.io.Connector;
+import javax.microedition.io.Datagram;
+import javax.microedition.io.DatagramConnection;
+
+import spot.src.org.sunspotworld.demo.communication.LEDSignals;
+
+import com.sun.spot.io.j2me.radiogram.RadiogramConnection;
+import com.sun.spot.io.j2me.radiostream.RadiostreamConnection;
+import com.sun.spot.peripheral.radio.NoMeshLayerAckException;
+import com.sun.spot.resources.transducers.LEDColor;
+import com.sun.spot.util.Utils;
+
+public class SpotIdentifier {
+
+	//Spot-Adresse für Host
+	String ourAddress;
+	String otherAddress;
+	String[] otherAdresses;
+	double learning_rate;
+	int sample_period;
+	int amount_of_data;
+	double LEARN_LIM;
+	int LEARN_COUNT;
+	int Faktor;
+	
+	private RadiostreamConnection listenOn;
+    private RadiostreamConnection sendOn;
+ 
+    private DataInputStream listenOnInputStream;
+    private DataOutputStream sendOnOutputStream;
+ 
+    private int portOut;
+    
+	private int portIn;
+    
+    private int spotID;
+    private int spotCount;
+    private String[] adresses;
+	
+    private int initialBroadCastPort;
+    
+    private boolean foundByHost = false;
+    private long timeout = 2000;
+ 
+    private RadiogramConnection broadcastConnectionIn;
+    private DatagramConnection broadcastConnectionOut;
+
+    private LEDSignals ledsignals = new LEDSignals();
+    
+    private static int MAX_BROADCAST_SIZE;
+    
+	/**
+	 * Sets the own address parameter and opens the broadcast connections on the defined port.
+	 * LED Signal set red light on led[0]. 
+	 * 
+	 * @param port port for the default broadcast connection.
+	 */
+	public SpotIdentifier(int port){
+		
+		//Lesen der eigenen Adresse
+		ourAddress = System.getProperty("IEEE_ADDRESS");
+		
+		//Parameter übergeben
+		//port-nummer
+		initialBroadCastPort = port;
+	
+		//BroadcastConnection öffnen (port für eingang = ausgang+1 -> für host umgekehrt)
+		try{
[…]

### 2015-01-22 Paper/PaperJahrSchlich.tex
Zeilenabstand, Fußnote entfernt

@@
 % The following parameters seem to provide a reasonable page setup.
-\topmargin 0.0cm
+\topmargin -1.0cm
 \oddsidemargin 0.2cm
-\textwidth 16cm 
-\textheight 21cm
+\textwidth 15cm 
+\textheight 23cm
 \footskip 1.0cm
@@
 
-\baselineskip24pt
+\baselineskip36pt
 
@@
 Over their lifetime, sensors can become inaccurate, faulty, or may even break.
+A fault can be defined as a defect of a sensor, that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
 To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
@@
 Physical redundancy, although efficient for sensor fault detection, shows the disadvantage of increased installation and maintenance costs. 
-Representing a more efficient approach, analytical redundancy applies knowledge of the structure and the SHM system itself to compute virtual sensor measurements for each sensor. 
+Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors.
+Virtual sensor measurements are computed for each sensor and then compared to the actual measurements.
 For example, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor.
@@
 \end{figure}
-\todo[inline]{Grafik kann nacher noch an unser neuronales Netz angepasst werden}
+\todo[inline]{Grafik wird an unser neuronales Netz angepasst}
 
@@
 This paper presents a decentralized autonomous sensor fault detection strategy for structural health monitoring systems using neural networks. 
-By implementing the neural networks into the sensor nodes, sensor faults can be detected by individual sensor nodes.
+We realized the autonomous sensor fault detection by implementing a neural network into each sensor node.
+The neural networks are trained to predict the expected sensor measurement, which is then compared to the actual measurment, revealing potential sensor faults.
+The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can, in fact, improve the dependability and the accuracy of structural health monitoring systems.
 

### 2015-01-23 Paper/PaperJahrSchlich.tex
Änderungen aus Besprechung

@@
 
-The dependability and the accuracy of structural health monitoring systems can be affected significantly by sensor faults. 
+The dependability and the accuracy of structural health monitoring systems can be affected by sensor faults. 
 In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
@@
 The obtained sensor data is collected by sensor nodes, and then analyzed and stored on a computer system over long periods of time. 
-Analysis of the sensor data can reveal abnormal changes in material and geometric behaviour early on.
+The analysis of the sensor data can reveal abnormal changes in material and geometric behaviour early on.
 
-Traditionally, the sensor nodes are wired with cables and connected to computer systems.
-However, using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs and inaccessibility of optimal sensor location with wires.
+Traditionally, the sensor nodes are connected to computer system with cables.
+However, using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
 In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
@@
 Over their lifetime, sensors can become inaccurate, faulty, or may even break.
-A fault can be defined as a defect of a sensor, that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
+A fault can be defined as a defect of a sensor that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
 To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
 The traditional approach to fault detection is the installation of physically redundant sensors.
-Faulty sensors can easily be identified through the deviation of their measurements from the measurements of correlated sensors.
-Physical redundancy, although efficient for sensor fault detection, shows the disadvantage of increased installation and maintenance costs. 
+Faulty sensors can be identified through the deviation of their measurements from the measurements of correlated sensors.
+Physical redundancy, although efficient for sensor fault detection, causes increased installation and maintenance costs due to the multiple installation of sensors. 
 Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors.
@@
 Artificial neural networks essentially consist of interconnected data processing units called neurons. 
-Neurons are grouped in different layers; usually one input layer, a number of hidden layers and one output layer.
-Artificial neural networks have the distinct ability of learning, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy \citep{Li2011}.
+The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
+Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy \citep{Li2011}.
 
@@
 First, a wireless structural health monitoring system is designed and implemented. 
-Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of said node in order to detect sensor faults in a decentralized manner. 
+Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node in order to detect sensor faults in a decentralized manner. 
 Then, the system is tested in laboratory experiments. 
@@
 In this project, the acceleration is measured using wireless sensor nodes and a basestation of the type "Oracle Sun SPOT". 
-The 3-axis digital output accelerometer with sensitivity ranging between $ \pm $ 2G and $\pm$ 8G has a maximum sample rate of 125Hz. \cite[9]{eDemo2010}
+The 3-axis digital output accelerometer with sensitivity ranging between $ \pm 2\,$g and $\pm 8\,$g has a maximum sample rate of 125Hz. \cite[9]{eDemo2010}
 

### 2015-02-02 Paper/PaperJahrSchlich.tex
java implementation im SHM kapitel

@@
 \usepackage{graphicx}
-\usepackage[authoryear, square]{natbib}
+\usepackage[numbers, sort&compress, square]{natbib}
+\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
 
 % used for line numbering
-%\usepackage{lineno}
+\usepackage{lineno}
 
@@
 
-\baselineskip36pt
+\baselineskip24pt
 
@@
 
-%\linenumbers % Schaltet Zeilennummerierung ein
-%\modulolinenumbers[5] % nur jede 5. Zeile
+\linenumbers % Schaltet Zeilennummerierung ein
+\modulolinenumbers[5] % nur jede 5. Zeile
 \section*{Dictionary}
@@
 Structural health monitoring (SHM) systems can be deployed to evaluate the conditions and to ensure the structural stability of civil engineering structures.
-\cite[4]{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure".
+\citet{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure".
 The obtained sensor data is collected by sensor nodes, and then analyzed and stored on a computer system over long periods of time. 
@@
 Virtual sensor measurements are computed for each sensor and then compared to the actual measurements.
-For example, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor.
-\citep{Smarsly2014}
+For example, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor
+\citep{Smarsly2014}.
 
@@
 \section*{Design and implementation of the wireless structural health monitoring system}
-In the following section, a brief description of SHM systems in general is given. Furthermore, the used wireless SHM system is introduced and the Java implementation is described.
+In the following section, the wireless SHM system is introduced and the Java implementation is described.
 
-According to \cite[5]{BisbySHM}, a structural health monitoring system commonly consists of the following components:
+According to \citet{BisbySHM}, a structural health monitoring system commonly consists of the following tasks:
 1. data acquisition,
@@
 Tasks 1 to 3 are usually implemented on site, while tasks 4 to 6 are usually implemented on computer systems off site.
-A wireless SHM system installed on the building site consists of sensor nodes attached to the construction, a base station, and a local computer. 
-The base station, which can be an external device or a computer program, links the individual sensor nodes with each other and the local computer.
+A wireless SHM system installed on the building site consists of sensor nodes attached to the construction, a basestation, and a local computer. 
+The basestation, which can be an external device or a computer program, links the individual sensor nodes with each other and the local computer.
 To each sensor node, several sensors measuring different physical quantities may be connected. 
 In this project, the acceleration is measured using wireless sensor nodes and a basestation of the type "Oracle Sun SPOT". 
-The 3-axis digital output accelerometer with sensitivity ranging between $ \pm 2\,$g and $\pm 8\,$g has a maximum sample rate of 125Hz. \cite[9]{eDemo2010}
+The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sample rate of 125Hz \citep{eDemo2010}.
 
@@
 On the computer, the data is stored in a MySQL database.
+\todo{Abb}
 
-\todo[inline]{vielleicht eine Abbildung? SunSPOTs? Aufbau SHM?}
+\begin{figure}[ht]
+    \centering
+    \fbox{\includegraphics{figures/missing.jpeg}}
+    \caption{class diagram describing the structure of the implemented SHM system}
+    \label{fig:UML}
+\end{figure}
 
-\textbf{3. paragraph: UML / Java implementaion}
+\autoref{fig:UML} describes the classes of the SHM system code, programmed in Java Platform, Micro Edition. 
+The SHM system consists of two packages---\texttt{SensorNode} and \texttt{Basestation}.
 
-%According to [3], a structural health monitoring system basically consists of the following components:
-%1. data acquisition, 2. data communication, 3. data processing, 4. data storage, 5. diagnostics and 6. data retrieval.
-%
-%
-%
-%For autonomous fault detection in wireless SHM systems, an integrated fault detection and synchronization strategy is implemented into a wireless SHM system.
-The SHM system consists of wireless sensor nodes and a base station.
-The sensor nodes are installed on the monitored structure and the base station links the wireless SHM system and a local computer.
-%
+The package \texttt{SensorNode} consits of the four classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, that are embedded directly into the sensor nodes.
+The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
+There are two phases: At first, the acceleration is measured with a low samplerate.
+Once the acceleration exceeds a threshold, the samplerate is increased and the measured values are stored into an array.
+The different phases are indicatad by lighting different LEDs.
+The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
+With the transformed data, the magnitudes and the correlating frequencies of the oscillation can be calculated.
+Finally, the natural frequency can be determined by extracting the maximal magnitude.
+\todo{FFT, Magnitude, Frequenz?}
+The \texttt{Communication} class opens a radio connection between the sensor node and the basestation and transfers data from the sensor node to the basestation.
+The entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
+Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
+
+The package \texttt{Basestation} runs on the host computer and operates the basestation.
+It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
+The \texttt{Database\-Handler} class establishes a connection to a database, creates a database table, if none with the specified name is available, and inserts data into the database table.
+The entry point of the programm is the \texttt{run()} method in the \texttt{MainBase} class.
+The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
+
 %
@@
 \item neural networks check integrity
-\item if no faults detected, sensor nodes send data to base station
+\item if no faults detected, sensor nodes send data to basestation
 \end{itemize}
@@
 \item neural networks check integrity -> error!
-\item sensor node sends alert to base station
+\item sensor node sends alert to basestation
 \end{itemize}
@@
 
-%\section*{Discussion of the experimental results}
+\section*{Notes}
 
+package hyperref eingebunden: klickbare links, autoref
+
+package SPOT in SensorNode umbenennen?!
+
+citep, citet, see
 %----------------------------------------------------------------------------------------

### 2015-02-02 SHM_JahrSchlich/SPOT/src/AccelerationSampler.java
lauffähiger, aufgeräumter, finaler code

@@ -1,0 +1,71 @@
+package spot;
+
+// author Jahr&Schlich
+
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.IAccelerometer3D;
+import com.sun.spot.util.Utils;
+import java.lang.Math;
+import com.sun.spot.resources.transducers.ITriColorLED;
+
+
+
+public class AccelerationSampler 
+{
+	IAccelerometer3D accelerometer = 
+			(IAccelerometer3D) Resources.lookup(IAccelerometer3D.class);
+	
+	
+    ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED7");
+    ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED6");
+	
+	public double[] getaccelerationArray (int samplePeriodIfListening, 
+										  int samplePeriodIfMeasuring, 
+										  int arrayLength, 
+										  double threshold) 
+	{
+		
+		double[] accelerationArray = new double[arrayLength];
+		
+        try {
+        	
+        	double accely = accelerometer.getAccelY();
+        	
+            // Flash an LED to indicate a sampling event
+            led.setRGB(255, 255, 255);
+            led.setOn();
+        	
+        	// measure acceleration with low samplerate, until acceleration 
+            // exceeds threshold
+        	while (Math.abs(accely) < threshold) {
+        		Utils.sleep(samplePeriodIfListening);
+        		accely = accelerometer.getAccelY();
+			}
+            led.setOff();
+            
+            
+            // light an LED to indicate a sampling event
+            led2.setRGB(0, 255, 255);
+            led2.setOn();
+        	
+        	// measure acceleration with high samplerate, store values in array
+			for (int i = 0; i < arrayLength; i++) {
+				long start = System.currentTimeMillis();
+				accelerationArray[i] = accelerometer.getAccelY();
+				System.out.println("accY=  " + accelerationArray[i]);
+				Utils.sleep(samplePeriodIfMeasuring - 
+										(System.currentTimeMillis() - start));
+				
+			}
+			
+			// turn off LED
+            led2.setOff();
+			
+        } catch (Exception e) {
+            System.err.println("Caught " + e + " during sensor sampling.");
+        }
+       
+        return accelerationArray;
+	}
+	
+}

### 2015-02-02 SHM_JahrSchlich/SPOT/src/Communication.java

@@ -1,0 +1,43 @@
+package spot;
+
+import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.ITriColorLED;
+import com.sun.spot.util.Utils;
+
+import javax.microedition.io.*;
+
+public class Communication {
+
+        RadiogramConnection rCon = null;
+        Datagram dg = null;
+        String ourAddress = System.getProperty("IEEE_ADDRESS");
+        ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED7");
+        
+        public void EstablishConnection(int hostPort){
+            try {
+                // Open up a broadcast connection to the host port
+                // where the 'on Desktop' portion of this demo is listening
+                rCon = (RadiogramConnection) Connector.open("radiogram://broadcast:" + hostPort);
+                dg = rCon.newDatagram(50);  // only sending 12 bytes of data
+                System.out.println("Starting communication on" + ourAddress);
+            } catch (Exception e) {
+                System.err.println("Caught " + e + " in connection initialization.");
+            }	
+        }
+
+        public void SendMeasurement(double magnitude, double frequencyMax) {
+        	try {
+	        	// Pack time and sensor reading into a radio datagram, send it.
+		        dg.reset();
+		        dg.writeDouble(magnitude);
+		        // cast frequency to double
+		        float frequency = (float) frequencyMax;
+		        dg.writeFloat(frequency);
+		        rCon.send(dg);
+		        System.out.println("SENT!");
+        	 } catch (Exception e) {   
+        		 System.err.println("Caught " + e + " in sending data.");
+        	 }
+        }
+    }

### 2015-02-02 SHM_JahrSchlich/SPOT/src/FFT.java

@@ -1,0 +1,193 @@
+package spot;
+/**
+ * @author Orlando Selenu
+// author Jahr&Schlich
+ * 
+ */
+
+import java.lang.Math;
+
+import com.sun.squawk.util.MathUtils;
+
+public class FFT {
+    /**
+     * The Fast Fourier Transform (generic version, with NO optimizations).
+     * 
+     * @param inputReal
+     *            an array of length n, the real part
+     * @param inputImag
+     *            an array of length n, the imaginary part
+     * @param DIRECT
+     *            TRUE = direct transform, FALSE = inverse transform
+     * @return a new array of length 2n
+     */
+    public static double[] performFFT(final double[] inputReal, double[] inputImag,
+	    boolean DIRECT) 
+    {
+		// - n is the dimension of the problem
+		// - nu is its logarithm in base e
+		int n = inputReal.length;
+	 
+		// If n is a power of 2, then ld is an integer (_without_ decimals)
+		double ld = MathUtils.log(n) / MathUtils.log(2.0);
+	 
+		// Here I check if n is a power of 2. If exist decimals in ld, I quit
+		// from the function returning null.
+		if (((int) ld) - ld != 0) 
+		{
+		    System.out.println("The number of elements is not a power of 2.");
+		    return null;
+		}
+	 
+		// Declaration and initialization of the variables
+		// ld should be an integer, actually, so I don't lose any information in
+		// the cast
+		int nu = (int) ld;
+		int n2 = n / 2;
+		int nu1 = nu - 1;
+		double[] xReal = new double[n];
+		double[] xImag = new double[n];
+		double tReal, tImag, p, arg, c, s;
+			
+		// Here I check if I'm going to do the direct transform or the inverse
+		// transform.
+		double constant;
+		if (DIRECT)
+		    constant = -2 * Math.PI;
+		else
+		    constant = 2 * Math.PI;
+	 
+		// I don't want to overwrite the input arrays, so here I copy them. This
+		// choice adds \Theta(2n) to the complexity.
+		for (int i = 0; i < n; i++) 
+		{
+		    xReal[i] = inputReal[i];
+		    xImag[i] = inputImag[i];
+		}
+		 
+		// First phase - calculation
+		int k = 0;
+		for (int l = 1; l <= nu; l++) 
+		{
[…]

### 2015-02-02 SHM_JahrSchlich/SPOT/src/MainSpot.java

@@ -1,0 +1,89 @@
+/*
+ *Main.java
+ *
+ * Copyright (c) 2008-2010 Sun Microsystems, Inc.
+ *
+ * Permission is hereby granted, free of charge, to any person obtaining a copy
+ * of this software and associated documentation files (the "Software"), to
+ * deal in the Software without restriction, including without limitation the
+ * rights to use, copy, modify, merge, publish, distribute, sublicense, and/or
+ * sell copies of the Software, and to permit persons to whom the Software is
+ * furnished to do so, subject to the following conditions:
+ *
+ * The above copyright notice and this permission notice shall be included in
+ * all copies or substantial portions of the Software.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+ * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+ * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+ * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+ * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
+ * FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
+ * DEALINGS IN THE SOFTWARE.
+ */
+
+package spot;
+
+//author Jahr&Schlich
+
+import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.ITriColorLED;
+import com.sun.spot.util.Utils;
+import javax.microedition.io.*;
+import javax.microedition.midlet.MIDlet;
+import javax.microedition.midlet.MIDletStateChangeException;
+import com.sun.spot.resources.transducers.IAccelerometer3D;
+import com.sun.spot.service.BootloaderListenerService;
+
+
+public class MainSpot extends MIDlet {
+
+    private int HOST_PORT = 67;
+    private int SAMPLE_PERIOD_LISTENING = 1 * 1000;  // in milliseconds
+    private int SAMPLE_PERIOD_MEASURING = 25;  // in milliseconds
+    private int ARRAY_LENGTH = 512;
[…]

### 2015-02-02 SHM_JahrSchlich/basestation/src/DatabaseHandler.java

@@ -1,0 +1,77 @@
+package basestation;
+
+import java.sql.DriverManager;
+import java.sql.PreparedStatement;
+import java.sql.ResultSet;
+//import java.sql.SQLException;
+import java.sql.Statement;
+import java.sql.Timestamp;
+import java.util.Calendar;
+
+public class DatabaseHandler {
+	
+	java.sql.Connection dbConn = null;
+	
+	public void getConnection(String url, String user, String password) {
+        try {            
+            // Get a connection to the database for given user/password
+        	dbConn = DriverManager.getConnection(url, user, password);
+
+            // Display URL and connection information
+        	System.out.println("Database connection etablished.");
+            System.out.println("\tURL: " + url);
+            System.out.println("\tConnection: " + dbConn);
+            
+        } catch (Exception e) {
+            System.err.println("setUp caught " + e);
+            System.err.println("Make sure that mySQL is installed properly \n" +
+                    "and has a Test database accessible via the \n" +
+                    "default administrator settings on localhost:3306\n");            
+        }
+	}
+    
+	public void createTableIfNeeded(String tableName) throws Exception {
+		
+		// Get a list of the names of all Tables in the Database
+		ResultSet resultSet = dbConn.getMetaData().getTables(null, null, tableName, null);
+		// Check if any of the Tables already has the specified name
+		if (resultSet.next()) {
+			System.out.println("Table " + tableName + " already exists.");
+		} else {
+			// Create a table with the specified name
+			System.out.println("Creating table " + tableName + ".");
+			Statement statement = dbConn.createStatement();
+			statement.execute("CREATE TABLE "+ tableName + " ("
+					+ "Nr INT NOT NULL AUTO_INCREMENT PRIMARY KEY, "
+					+ "Time TIMESTAMP, "
+					+ "SpotID VARCHAR(19), "
+					+ "Magnitude DOUBLE, "
+					+ "Frequency FLOAT)");
+			statement.close();
+		}
+		resultSet.close();
+	}
+	
+	
+	public void writeMeasurement(String tableName, String spotID, 
+						double magnitude, float frequency) throws Exception {
+		
+		
+		//Get current systemtime as timestamp to avoid trouble with unsynchronized SunSPOTs
+		Calendar calendar = Calendar.getInstance();
+        Timestamp now = new Timestamp(calendar.getTime().getTime());
+        
[…]

### 2015-02-02 SHM_JahrSchlich/basestation/src/MainBase.java

@@ -1,0 +1,120 @@
+package basestation;
+
+/*
+ * SendDataDemoHostApplication.java
+ *
+ * Copyright (c) 2008-2009 Sun Microsystems, Inc.
+ *
+ * Permission is hereby granted, free of charge, to any person obtaining a copy
+ * of this software and associated documentation files (the "Software"), to
+ * deal in the Software without restriction, including without limitation the
+ * rights to use, copy, modify, merge, publish, distribute, sublicense, and/or
+ * sell copies of the Software, and to permit persons to whom the Software is
+ * furnished to do so, subject to the following conditions:
+ *
+ * The above copyright notice and this permission notice shall be included in
+ * all copies or substantial portions of the Software.
+ *
+ * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+ * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+ * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
+ * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+ * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
+ * FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
+ * DEALINGS IN THE SOFTWARE.
+ */
+
+import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.peripheral.ota.OTACommandServer;
+
+import java.sql.DriverManager;
+import java.sql.Connection;
+import java.sql.Timestamp;
+import java.sql.SQLException;
+import java.text.DateFormat;
+import java.util.Calendar;
+import java.util.Date;
+import java.util.Properties;
+
+import javax.microedition.io.*;
+
+import java.sql.PreparedStatement;
+import java.sql.ResultSet;
+import java.sql.Statement;
+
+/**
+ * This application is the 'on Desktop' portion of the SendDataDemo. 
+ * This host application collects sensor samples sent by the 'on SPOT'
+ * portion running on neighboring SPOTs and just prints them out. 
+ *   
+ * @author: Vipul Gupta
+ * modified: Ron Goldman
+ */
+public class MainBase {
[…]

### 2015-02-02 SHM_NN_2014/src/accelerometer/AccelerometerSampleCode.java
obsolenten Code entfernt

@@ -1,123 +1,0 @@
-/*
- * Copyright (c) 2007-2010 Sun Microsystems, Inc.
- * 
- * Permission is hereby granted, free of charge, to any person obtaining a copy
- * of this software and associated documentation files (the "Software"), to 
- * deal in the Software without restriction, including without limitation the 
- * rights to use, copy, modify, merge, publish, distribute, sublicense, and/or 
- * sell copies of the Software, and to permit persons to whom the Software is 
- * furnished to do so, subject to the following conditions:
- * 
- * The above copyright notice and this permission notice shall be included in 
- * all copies or substantial portions of the Software.
- * 
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR 
- * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, 
- * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE 
- * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER 
- * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING 
- * FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER 
- * DEALINGS IN THE SOFTWARE.
- **/
-
-package accelerometer;
-
-import com.sun.spot.resources.Resources;
-import com.sun.spot.resources.transducers.LEDColor;
-import com.sun.spot.resources.transducers.IAccelerometer3D;
-import com.sun.spot.resources.transducers.ITriColorLEDArray;
-import com.sun.spot.util.Utils;
-
-import java.io.IOException;
-
-import javax.microedition.midlet.MIDlet;
-import javax.microedition.midlet.MIDletStateChangeException;
-
-/**
- * Sample code snippet to show how to use the accelerometer
- * on the Sun SPOT General Purpose Sensor Board.
- *
- * Simple use of the accelerometer to measure tilt on the Sun SPOT.
- *
- * @author Ron Goldman<br>
- * date: August 14, 2006
- * date: March 12, 2007 - modified to use new IAccelerometer3D interface
- * date: May 26, 2010 - modified to use new Resource & Sensor APIs
- */
[…]

### 2015-02-02 SHM_NN_2014/src/fft/FFTbase.java

@@ -1,166 +1,0 @@
-package fft;
-/**
- * @author Orlando Selenu
- * 
- */
-public class FFTbase {
-    /**
-     * The Fast Fourier Transform (generic version, with NO optimizations).
-     * 
-     * @param inputReal
-     *            an array of length n, the real part
-     * @param inputImag
-     *            an array of length n, the imaginary part
-     * @param DIRECT
-     *            TRUE = direct transform, FALSE = inverse transform
-     * @return a new array of length 2n
-     */
-    public static double[] fft(final double[] inputReal, double[] inputImag,
-	    boolean DIRECT) 
-    {
-		// - n is the dimension of the problem
-		// - nu is its logarithm in base e
-		int n = inputReal.length;
-	 
-		// If n is a power of 2, then ld is an integer (_without_ decimals)
-		double ld = Math.log(n) / Math.log(2.0);
-	 
-		// Here I check if n is a power of 2. If exist decimals in ld, I quit
-		// from the function returning null.
-		if (((int) ld) - ld != 0) 
-		{
-		    System.out.println("The number of elements is not a power of 2.");
-		    return null;
-		}
-	 
-		// Declaration and initialization of the variables
-		// ld should be an integer, actually, so I don't lose any information in
-		// the cast
-		int nu = (int) ld;
-		int n2 = n / 2;
-		int nu1 = nu - 1;
-		double[] xReal = new double[n];
-		double[] xImag = new double[n];
-		double tReal, tImag, p, arg, c, s;
-			
-		// Here I check if I'm going to do the direct transform or the inverse
-		// transform.
-		double constant;
-		if (DIRECT)
-		    constant = -2 * Math.PI;
-		else
-		    constant = 2 * Math.PI;
-	 
-		// I don't want to overwrite the input arrays, so here I copy them. This
-		// choice adds \Theta(2n) to the complexity.
-		for (int i = 0; i < n; i++) 
-		{
-		    xReal[i] = inputReal[i];
-		    xImag[i] = inputImag[i];
-		}
-		 
-		// First phase - calculation
-		int k = 0;
-		for (int l = 1; l <= nu; l++) 
-		{
-		    while (k < n) 
-		    {
-				for (int i = 1; i <= n2; i++) 
-				{
-				    p = bitreverseReference(k >> nu1, nu);
[…]

### 2015-02-02 SHM_NN_2014/src/fft/GetCSV.java

@@ -1,57 +1,0 @@
-package fft;
-
-
-import java.io.File;
-import java.io.FileNotFoundException;
-import java.util.Scanner;
-
-public class GetCSV
-{
-    public static double[] Load(String file) throws FileNotFoundException
-    {
-        //Get scanner instance
-        Scanner scanner = new Scanner(new File(file));
-        //Set the delimiter used in file
-        scanner.useDelimiter(",");
-        
-        //Count elements in CSV file
-        // TO DO: there must be a better way
-        int i = 0;
-        while (scanner.hasNextLine())
-        {
-            scanner.nextLine();
-            i+=1;
-        }
-        // print number of values
-        //System.out.println(i);
-    
-        //close scanner
-        scanner.close();
-        
-        
-        //Get second scanner instance
-        // TO DO: Set scanner 1 to first position instead
-        Scanner scanner2 = new Scanner(new File(file));
-        //Set the delimiter used in file
-        scanner.useDelimiter(",");
-                
-        // initialize array of complex numbers
-        double[] y = new double[i];
-        
-        // parse text to double, create complex number, add it to array
-        int j = 0;    
-        while (scanner2.hasNextLine())
-        {
-        	double a = Double.valueOf(scanner2.nextLine());
-        	y[j]=new Double(a);
-            //System.out.print(y[j]);
-        	j+=1;
-        }
-        
-        //close scanner2 
-        scanner2.close();
-
-        // return array of complex numbers
-        return y;
-    }
-}

### 2015-02-02 SHM_NN_2014/src/fft/Main.java

@@ -1,41 +1,0 @@
-package fft;
-
-import java.io.FileNotFoundException;
-import javax.swing.JFrame;
-
-import org.math.plot.*;
-
-public class Main
-{
-	public static void main(String[] args) throws FileNotFoundException
-    {
-
-		double[] inputReal = GetCSV.Load("/media/katrin/Data/eigeneDateien/Uni/Master/5/SHM/SHMCode/simdata.txt");
-		double[] inputImag = new double[inputReal.length];
-		boolean DIRECT = Boolean.TRUE;
-		double samplerate = 512.; // example samplerate
-
-		// perform fft
-		double[] transformedBase = FFTbase.fft(inputReal, inputImag, DIRECT);
-		
-		// calculate magnitude and frequencies
-		double [] magnitude = FFTbase.calculateMagnitude(transformedBase);
-		double[] frequency = FFTbase.calculateFrequency(transformedBase, samplerate);
-        
-        // plot frequency spectrum
-        plot(frequency, magnitude);
-              
-    }
-	
-	private static void plot(double[] frequency, double[] magnitude)
-	{
-		Plot2DPanel panel = new Plot2DPanel();
-        panel.addLinePlot("Line", frequency, magnitude);
-        	
-        JFrame  frame= new JFrame("frequency spectrum");
-        frame.setContentPane(panel);
-        frame.setSize(500, 600);
-        frame.setDefaultCloseOperation(JFrame.EXIT_ON_CLOSE);
-        frame.setVisible(true);
-	}
-}

### 2015-02-02 SHM_NN_2014/src/fft/SimData.java

@@ -1,32 +1,0 @@
-package fft;
-
-import java.util.Vector;
-
-public class SimData {
-	public static void main(){
-	// generate x-vector
-	//Vector x = new Vector();
-	//x = arange(0, 20, 0.001);
-			
-		// define a sinus function
-		
-	Vector<Double> y = new Vector<Double>();
-		
-	y.addElement(1.);
-		
-		
-	/*
-	//y = sin(pi * x);
-	//another sinus function
-	z = 0.1* sin(pi * 30 * x)+1;
-	// generate noise by adding values from gaussian distribution
-	n = random.normal(-0.01,0.01,len(y));
-	// and now evanescent
-	y = exp(-0.1 * x) * sin(pi * x);
-	z = exp(-0.1 * x) * 0.1* sin(pi * 30 * x)+1;
-	// add signals
-	y = y + z + n;
-	print(y);
-	*/
-	}
-}

### 2015-02-02 SHM_NN_2014_Host/src/MersenneTwisterFast.java

@@ -1,1344 +1,0 @@
-package spot.src.org.sunspotworld.demo;
-
-
-
-import java.io.DataInputStream;
-import java.io.DataOutputStream;
-import java.io.IOException;
-import java.util.Random;
-
-import com.sun.squawk.util.MathUtils;
-
-
-
-/**
- * <h3>MersenneTwister and MersenneTwisterFast</h3>
- * <p>
- * <b>Version 16</b>, based on version MT199937(99/10/29) of the Mersenne
- * Twister algorithm found at <a
- * href="http://www.math.keio.ac.jp/matumoto/emt.html"> The Mersenne Twister
- * Home Page</a>, with the initialization improved using the new 2002/1/26
- * initialization algorithm By Sean Luke, October 2004.
- * 
- * <p>
- * <b>MersenneTwister</b> is a drop-in subclass replacement for
- * java.util.Random. It is properly synchronized and can be used in a
- * multithreaded environment. On modern VMs such as HotSpot, it is approximately
- * 1/3 slower than java.util.Random.
- * 
- * <p>
- * <b>MersenneTwisterFast</b> is not a subclass of java.util.Random. It has the
- * same public methods as Random does, however, and it is algorithmically
- * identical to MersenneTwister. MersenneTwisterFast has hard-code inlined all
- * of its methods directly, and made all of them final (well, the ones of
- * consequence anyway). Further, these methods are <i>not</i> synchronized, so
- * the same MersenneTwisterFast instance cannot be shared by multiple threads.
- * But all this helps MersenneTwisterFast achieve well over twice the speed of
- * MersenneTwister. java.util.Random is about 1/3 slower than
- * MersenneTwisterFast.
- * 
- * <h3>About the Mersenne Twister</h3>
- * <p>
- * This is a Java version of the C-program for MT19937: Integer version. The
- * MT19937 algorithm was created by Makoto Matsumoto and Takuji Nishimura, who
- * ask: "When you use this, send an email to: matumoto@math.keio.ac.jp with an
- * appropriate reference to your work". Indicate that this is a translation of
- * their algorithm into Java.
- * 
- * <p>
- * <b>Reference. </b> Makato Matsumoto and Takuji Nishimura, "Mersenne Twister:
- * A 623-Dimensionally Equidistributed Uniform Pseudo-Random Number Generator",
[…]

### 2015-02-03 Paper/PaperJahrSchlich.tex
Extended section 'laboratory experiments'

@@
 
-To validate the proposed approach, the SHM system is installed on a test structure in order to conduct laboratory experiments.
-The test structure consists of four metal sheets with an area of 20\,cm by 60\,cm and 4\,mm of thickness.
-Those metal sheets are mounted on threaded rods with a vertical clearance of approximately 40\,cm.
-At the bottom, the threaded rods are fixed into a block of solid wood with an area of 40\,cm by 60\,cm and a height of 50\,cm.
-The SHM system is installed onto the test structure by fastening a SunSPOT to each of the metal sheets.
-The finished laboratory setup is shown in figure~\ref{fig:teststructure}.
+To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure.
+The test structure is a 4-story shear-frame consisting of four steel plates with an area of 25\,cm\,$\times$\,50\,cm and a thickness of 0.8\,mm.
+Those plates are mounted on threaded rods with a vertical clearance of 23\,cm.
+At the bottom, the rods are fixed into a solid block of wood with an area of 40\,cm\,$\times$\,60\,cm and a height of 30\,cm.
+The SHM system is installed on the test structure by fastening wireless sensor nodes to each story.
+The laboratory setup is shown in figure~\ref{fig:teststructure}.
 \todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
-The structure is excited by displacing one of the metal sheets to the side and releasing it.
-This method of excitation ensures, that the structure vibrates in its Eigenfrequency with little interferences.
 
@@
 \end{figure}
+
+The structure is excited by deflecting and releasing the top of the structure.
+This excitation method ensures that the structure vibrates in its Eigenfrequency with little interferences.
+After excitation the sensor nodes automatically start measuring acceleration values with a sampling rate of 40\,Hz.
+To minimize data traffic between the sensor nodes, each node performs a fast Fourier transform (FFT) algorithm after sufficient acceleration samples have been collected.
+Generally FFTs are used to convert equidistant samples of a function into a combination of frequencies, and the corresponding magnitudes, that has the same sample values.
+In this case the acceleration samples are converted into the vibration frequencies, and the corresponding magnitudes, of the building.
+Each sensor node sends only the values of the predominating frequency, i.e. the frequency with the biggest magnitude, to the other nodes.
 

### 2015-02-03 Paper/PaperJahrSchlich.tex
Extended section 'Laboratory experiments'
Überarbeitung SHM

--- 2015-02-02 Paper/PaperJahrSchlich.tex
@@
 \usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
+
+\usepackage{color}
 
@@
 The obtained sensor data is collected by sensor nodes, and then analyzed and stored on a computer system over long periods of time. 
-The analysis of the sensor data can reveal abnormal changes in material and geometric behaviour early on.
+The analysis of the sensor data can reveal abnormal changes in material and geometric behaviour at an early stage.
 
@@
 
-Over their lifetime, sensors can become inaccurate, faulty, or may even break.
+Over \colorbox{cyan}{their} lifetime, sensors can become inaccurate, faulty, or may even break.
 A fault can be defined as a defect of a sensor that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
 To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
-The traditional approach to fault detection is the installation of physically redundant sensors.
+\colorbox{cyan}{An often used approach} to fault detection is the installation of physically redundant sensors.
 Faulty sensors can be identified through the deviation of their measurements from the measurements of correlated sensors.
@@
 The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
-Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with arbitrary accuracy \citep{Li2011}.
+Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with \colorbox{cyan}{any desired accuracy} \citep{Li2011}.
 
@@
 \section*{Design and implementation of the wireless structural health monitoring system}
-In the following section, the wireless SHM system is introduced and the Java implementation is described.
+In the following section, the wireless SHM system is introduced and \colorbox{cyan}{the} Java implementation is described.
 
@@
 Tasks 1 to 3 are usually implemented on site, while tasks 4 to 6 are usually implemented on computer systems off site.
-A wireless SHM system installed on the building site consists of sensor nodes attached to the construction, a basestation, and a local computer. 
-The basestation, which can be an external device or a computer program, links the individual sensor nodes with each other and the local computer.
+A wireless SHM system installed on the structure consists of sensor nodes, a basestation, and a local computer. 
+The basestation, which can be an external device or a computer program, links the sensor nodes with each other and with the local computer.
 To each sensor node, several sensors measuring different physical quantities may be connected. 
@@
 
-In operation, the sensor nodes acquire sensor data and may perform initial processing.
-The processed data is then transmitted via radio to the local computer, where it is stored for analysis and diagnosis.
-In this project, the data is transmitted from the sensor nodes to the basestation by radio, and from the basestation to the local computer via USB cable.
-On the computer, the data is stored in a MySQL database.
+\colorbox{cyan}{folgender Absatz geändert:}
+
+During system operation, the sensor nodes acquire acceleration values and perform a fast Fourrier transform to determine the characteristics of the oscillation of the structure.
+The processed data is then transmitted via radio to the basestation and, finally, to the local computer, where it is stored in a MySQL database for further analysis and diagnosis.
+
 \todo{Abb}
@@
 \autoref{fig:UML} describes the classes of the SHM system code, programmed in Java Platform, Micro Edition. 
-The SHM system consists of two packages---\texttt{SensorNode} and \texttt{Basestation}.
+The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
 
-The package \texttt{SensorNode} consits of the four classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, that are embedded directly into the sensor nodes.
+The package \texttt{sensornode} consits of the four classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, that are embedded directly into the sensor nodes.
 The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
@@
 The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
-With the transformed data, the magnitudes and the correlating frequencies of the oscillation can be calculated.
+With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation can be calculated.
 Finally, the natural frequency can be determined by extracting the maximal magnitude.
-\todo{FFT, Magnitude, Frequenz?}
 The \texttt{Communication} class opens a radio connection between the sensor node and the basestation and transfers data from the sensor node to the basestation.
@@
 
-The package \texttt{Basestation} runs on the host computer and operates the basestation.
+The package \texttt{basestation} runs on the host computer and operates the basestation.
 It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
-The \texttt{Database\-Handler} class establishes a connection to a database, creates a database table, if none with the specified name is available, and inserts data into the database table.
+The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
 The entry point of the programm is the \texttt{run()} method in the \texttt{MainBase} class.

### 2015-02-03 SHM_JahrSchlich/SPOT/src/AccelerationSampler.java

@@ -1,7 +1,5 @@
 package spot;
 
-// author Jahr&Schlich
-
 import com.sun.spot.resources.Resources;
 import com.sun.spot.resources.transducers.IAccelerometer3D;
 import com.sun.spot.util.Utils;

### 2015-02-03 SHM_JahrSchlich/SPOT/src/FFT.java

@@ -1,7 +1,7 @@
 package spot;
 /**
  * @author Orlando Selenu
-// author Jahr&Schlich
+ * modified by Jahr&Schlich
  * 
  */
 

### 2015-02-03 SHM_JahrSchlich/basestation/src/DatabaseHandler.java

@@ -3,7 +3,6 @@ package basestation;
 import java.sql.DriverManager;
 import java.sql.PreparedStatement;
 import java.sql.ResultSet;
-//import java.sql.SQLException;
 import java.sql.Statement;
 import java.sql.Timestamp;
 import java.util.Calendar;

### 2015-02-03 SHM_JahrSchlich/basestation/src/MainBase.java

@@ -1,29 +1,5 @@
 package basestation;
 
-/*
- * SendDataDemoHostApplication.java
- *
- * Copyright (c) 2008-2009 Sun Microsystems, Inc.
- *
- * Permission is hereby granted, free of charge, to any person obtaining a copy
- * of this software and associated documentation files (the "Software"), to
- * deal in the Software without restriction, including without limitation the
- * rights to use, copy, modify, merge, publish, distribute, sublicense, and/or
- * sell copies of the Software, and to permit persons to whom the Software is
- * furnished to do so, subject to the following conditions:
- *
- * The above copyright notice and this permission notice shall be included in
- * all copies or substantial portions of the Software.
- *
- * THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
- * IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
- * FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
- * AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
- * LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING
- * FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER
- * DEALINGS IN THE SOFTWARE.
- */
-
 import com.sun.spot.io.j2me.radiogram.*;
 import com.sun.spot.peripheral.ota.OTACommandServer;
 
@@ -42,14 +18,6 @@ import java.sql.PreparedStatement;
 import java.sql.ResultSet;
 import java.sql.Statement;
 
-/**
- * This application is the 'on Desktop' portion of the SendDataDemo. 
- * This host application collects sensor samples sent by the 'on SPOT'
- * portion running on neighboring SPOTs and just prints them out. 
- *   
- * @author: Vipul Gupta
- * modified: Ron Goldman
- */
 public class MainBase {
     // Broadcast port on which we listen for sensor samples
     private static final int HOST_PORT = 67;
@@ -104,11 +72,6 @@ public class MainBase {
         }
     }
     
-    /**
-     * Start up the host application.
-     *
-     * @param args any command line arguments
-     */
     public static void main(String[] args) throws Exception {
         // register the application's name with the OTA Command server 
     	// & start OTA running

### 2015-02-03 Paper/PaperJahrSchlich.tex
Lab Ex überarbeitet

@@
 
-\todo{Abb}
-
 \begin{figure}[ht]
@@
 
-%
-%In the present study, wireless sensor nodes of the type “Oracle Sun SPOTs” are used.
-The sensor nodes are Java-programmable enabling embedded analyses and the sensor nodes include a 3-dimensional accelerometer, temperature and other sensors (Oracle Lab, 2014).
-In this study, acceleration measurements taken from the structure are used for analyzing the structural condition.
-Furthermore, the radios of the sensor nodes ensure wireless communications between the sensor nodes as well as between the sensor nodes and the base station in order to transmit monitoring data or control commands.
-The base station is able to receive monitoring data measured by the sensor nodes, and the base station can store the data in a database installed on the local computer.
-The main Java classes are shown in an Unified Modelling Language diagram (UML diagram), which is a standard way to visualize the implementation of a application.
-The following UML diagrams (cp.
-Figure 1 and Figure 2) show just the main classes embedded into the base station and the sensor nodes.
-In the following subsection, the implementation of the main components of the fault detection strategy, including synchronization and fault detection, are shown.
-
 \newpage
@@
 
-\textbf{2. paragraph: proposed nn}
-
-\begin{itemize}
-\item SNIPE? implementation in java?
-\item activation function / identity function
-\end{itemize}
-
-\textbf{3. paragraph: learning general}
-
-\begin{itemize}
-\item set with test values
-\item iterative weight adjustment
-\end{itemize}
-
-\textbf{4. paragraph?: learning, proposed}
+%\textbf{2. paragraph: proposed nn}
+%
+%\begin{itemize}
+%\item SNIPE? implementation in java?
+%\item activation function / identity function
+%\end{itemize}
+%
+%\textbf{3. paragraph: learning general}
+%
+%\begin{itemize}
+%\item set with test values
+%\item iterative weight adjustment
+%\end{itemize}
+%
+%\textbf{4. paragraph?: learning, proposed}
 
@@
 
-%\begin{itemize}
-%\item test structure
-%\item sensor installation
-%\item type of excitation, acceleration is measured
-%\end{itemize}
-
-To validate the proposed approach, the SHM system is installed on a test structure in order to conduct laboratory experiments.
-The test structure consists of four metal sheets with an area of 20\,cm by 60\,cm and 4\,mm of thickness.
-Those metal sheets are mounted on threaded rods with a vertical clearance of approximately 40\,cm.
-At the bottom, the threaded rods are fixed into a block of solid wood with an area of 40\,cm by 60\,cm and a height of 50\,cm.
-The SHM system is installed onto the test structure by fastening a SunSPOT to each of the metal sheets.
-The finished laboratory setup is shown in figure~\ref{fig:teststructure}.
+To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure.
+The test structure is a 4-story shear-frame consisting of four steel plates with an area of 25\,cm\,$\times$\,50\,cm and a thickness of 0.8\,mm.
+Those plates are mounted on threaded rods with a vertical clearance of 23\,cm.
+At the bottom, the rods are fixed into a solid block of wood with an area of 40\,cm\,$\times$\,60\,cm and a height of 30\,cm.
+The SHM system is installed on the test structure by fastening wireless sensor nodes to each story.
+The laboratory setup is shown in \autoref{fig:teststructure}.
 \todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
-The structure is excited by displacing one of the metal sheets to the side and releasing it.
-This method of excitation ensures, that the structure vibrates in its Eigenfrequency with little interferences.
 
@@
 
-\textbf{2. paragraph: data measurement and processing}
+The structure is excited by deflecting and releasing the top of the structure.
+This excitation method ensures that the structure vibrates in a natural frequency with little interferences.
+After excitation the sensor nodes automatically start measuring acceleration values with a sampling rate of 40\,Hz.
+To minimize data traffic, each node performs a fast Fourier transform algorithm after sufficient acceleration samples have been collected.
+%Generally FFTs are used to convert equidistant samples of a function into a combination of frequencies, and the corresponding magnitudes, that has the same sample values. 
+The acceleration samples are converted into the vibration frequencies, and the corresponding magnitudes, of the building.
+Each sensor node sends only the values of the predominating frequency, i.e. the frequency with the biggest magnitude.
+The values are then stored in a MySQL database.
 
-\begin{itemize}
-\item learning phase
-\item excitation
-\item sensors start measuring
-\item sensor nodes \textbf{perform fft} - what is fft
-\item sensor nodes exchange frequencies
-\item neural networks check integrity
-\item if no faults detected, sensor nodes send data to basestation
-\end{itemize}
+%\textbf{2. paragraph: data measurement and processing}
+%
+%\emph{\begin{itemize}
+%\item learning phase
+%\item excitation
+%\item sensors start measuring
+%\item sensor nodes \textbf{perform fft} - what is fft
+%\item sensor nodes exchange frequencies
+%\item neural networks check integrity
+%\item if no faults detected, sensor nodes send data to basestation
+%\end{itemize}}
+%
+%\textbf{3. paragraph: sensor fault detection}
+%
+%\begin{itemize}
+%\item simulation of sensor fault
+%\item neural networks check integrity -> error!
+%\item sensor node sends alert to basestation
+%\end{itemize}
 
-\textbf{3. paragraph: sensor fault detection}
-
-\begin{itemize}
-\item simulation of sensor fault
-\item neural networks check integrity -> error!
-\item sensor node sends alert to basestation
-\end{itemize}
-
-%----------------------------------------------------------------------------------------
-
-\section*{Notes}
-
-package hyperref eingebunden: klickbare links, autoref
-
-package SPOT in SensorNode umbenennen?!
-
-citep, citet, see
 %----------------------------------------------------------------------------------------

### 2015-02-04 Paper/PaperJahrSchlich.tex
Änderungen Kaffee Laden

@@
 
-\baselineskip36pt
+\baselineskip20pt
 
@@
 
-% Place your abstract within the special {sciabstract} environment.
+% 
+%\sloppy
+\setlength{\emergencystretch}{3pt}
+\hyphenation{sample-rate auto-nomous}
 
@@
 
-The dependability and the accuracy of structural health monitoring systems can be affected by sensor faults. 
+The dependability and the accuracy of structural health moni\-toring systems can be affected by sensor faults. 
 In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
@@
 
-Traditionally, the sensor nodes are connected to computer system with cables.
+Traditionally, the sensor nodes are connected to computer systems with cables.
 However, using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
@@
 
-Over \colorbox{cyan}{their} lifetime, sensors can become inaccurate, faulty, or may even break.
+Over their lifetime, sensors can become inaccurate, faulty, or may even break.
 A fault can be defined as a defect of a sensor that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
 To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
-\colorbox{cyan}{An often used approach} to fault detection is the installation of physically redundant sensors.
+A well known approach to fault detection is the installation of physically redundant sensors.
 Faulty sensors can be identified through the deviation of their measurements from the measurements of correlated sensors.
@@
 The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
-Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with \colorbox{cyan}{any desired accuracy} \citep{Li2011}.
+Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.
 
@@
 First, a wireless structural health monitoring system is designed and implemented. 
-Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node in order to detect sensor faults in a decentralized manner. 
+Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
 Then, the system is tested in laboratory experiments. 
@@
 \section*{Design and implementation of the wireless structural health monitoring system}
-In the following section, the wireless SHM system is introduced and \colorbox{cyan}{the} Java implementation is described.
+In the following section, the wireless structural health monitoring system is introduced and the software implementation is described.
+The wireless SHM system consists of wireless sensor nodes and a host computer, linked with a basestation.
+The sensor nodes and the basestation are of type "Oracle Sun SPOT". 
+The Sun SPOTS are equipped with several components, see \autoref{fig:SPOT}.
+Among others, the sensor board includes an accelerometer and eight independent RGB-LEDs.
+The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125Hz \citep{eDemo2010}.
 
-According to \citet{BisbySHM}, a structural health monitoring system commonly consists of the following tasks:
+\todo{citation BILD!}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[scale=0.5]{figures/eDemoboard.png}
+    \caption{Layout of the Sun SPOT sensor board}
+    \label{fig:SPOT}
+\end{figure}
+
+The SHM system performs the following tasks:
 1. data acquisition,
-2. data transmission, 
-3. data processing,
+2. data processing,
+3. data transmission, 
 4. data storage,
@@
 6. information retrieval.
+Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a fast Fourier transform to determine the characteristics of the oscillation of the structure.
+The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
+On the host computer, tasks 4 is conducted by storing the data in a MySQL database.
+Task 5 and 6---additional analysis and diagnosis---are conducted on the host computer in further steps.
 
-Tasks 1 to 3 are usually implemented on site, while tasks 4 to 6 are usually implemented on computer systems off site.
-A wireless SHM system installed on the structure consists of sensor nodes, a basestation, and a local computer. 
-The basestation, which can be an external device or a computer program, links the sensor nodes with each other and with the local computer.
-To each sensor node, several sensors measuring different physical quantities may be connected. 
-In this project, the acceleration is measured using wireless sensor nodes and a basestation of the type "Oracle Sun SPOT". 
-The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sample rate of 125Hz \citep{eDemo2010}.
+\todo[inline]{objektorientierung}
+The implementation is object oriented, using classes. Classes are small pieces of code.
+Da werden am Ende Programme draus gemacht und gestartet.
 
-\colorbox{cyan}{folgender Absatz geändert:}
+\autoref{fig:UML} describes the classes of the SHM system code, programmed in Java.
+The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
 
-During system operation, the sensor nodes acquire acceleration values and perform a fast Fourrier transform to determine the characteristics of the oscillation of the structure.
-The processed data is then transmitted via radio to the basestation and, finally, to the local computer, where it is stored in a MySQL database for further analysis and diagnosis.
+...
 
-\begin{figure}[ht]
+The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
+The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
+There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
+Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
+The measured values are stored into an array.
+The different phases are indicatad by lighting different LEDs.
+The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
+With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
+Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
+The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
+For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
+Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
+
+The package \texttt{basestation} runs on the host computer and operates the basestation.
+It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
+The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
+The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class.
+The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
+
+\begin{figure}[htb]
     \centering
@@
 \end{figure}
-
-\autoref{fig:UML} describes the classes of the SHM system code, programmed in Java Platform, Micro Edition. 
-The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
-
-The package \texttt{sensornode} consits of the four classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, that are embedded directly into the sensor nodes.
-The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
-There are two phases: At first, the acceleration is measured with a low samplerate.
-Once the acceleration exceeds a threshold, the samplerate is increased and the measured values are stored into an array.
-The different phases are indicatad by lighting different LEDs.
-The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
-With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation can be calculated.
-Finally, the natural frequency can be determined by extracting the maximal magnitude.
-The \texttt{Communication} class opens a radio connection between the sensor node and the basestation and transfers data from the sensor node to the basestation.
-The entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
-Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
-
-The package \texttt{basestation} runs on the host computer and operates the basestation.
-It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
-The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
-The entry point of the programm is the \texttt{run()} method in the \texttt{MainBase} class.
-The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
 
@@
 The structure is excited by deflecting and releasing the top of the structure.
-This excitation method ensures that the structure vibrates in a natural frequency with little interferences.
-After excitation the sensor nodes automatically start measuring acceleration values with a sampling rate of 40\,Hz.
-To minimize data traffic, each node performs a fast Fourier transform algorithm after sufficient acceleration samples have been collected.
+This excitation method ensures a free vibration in natural frequency with little interferences.
+After excitation, the sensor nodes automatically start the acceleration with a sampling rate of 40\,Hz.
+To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.
 %Generally FFTs are used to convert equidistant samples of a function into a combination of frequencies, and the corresponding magnitudes, that has the same sample values. 
-The acceleration samples are converted into the vibration frequencies, and the corresponding magnitudes, of the building.
+The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
 Each sensor node sends only the values of the predominating frequency, i.e. the frequency with the biggest magnitude.
-The values are then stored in a MySQL database.
+The values are then stored in the MySQL database at the host computer.
 

### 2015-02-05 Paper/PaperJahrSchlich.tex
object orientation, minor changes in summary

@@
 
-\todo[inline]{objektorientierung}
-The implementation is object oriented, using classes. Classes are small pieces of code.
-Da werden am Ende Programme draus gemacht und gestartet.
+\colorbox{cyan}{neuer Absatz:}
 
-\autoref{fig:UML} describes the classes of the SHM system code, programmed in Java.
+The SHM system is programmed object-oriented in Java.
+Object orientation uses objects, that are instanciated using classes as paradigm.
+A class includes methods, allowing the objects to perform actions, and attributes, specific data.
+\autoref{fig:UML} describes the classes of the SHM system code.
 The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
-
-...
+A package organizes several Java classes that build a program.
 
@@
     \includegraphics[scale=0.3]{figures/teststructure.jpg}
-    \caption{Picture of the laboratory setup}
+    \caption{Laboratory setup}
     \label{fig:teststructure}
@@
 The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
-Each sensor node sends only the values of the predominating frequency, i.e. the frequency with the biggest magnitude.
+Each sensor node transfers only the values of the predominating frequency---i.e.
+the frequency with the biggest magnitude--- \colorbox{cyan}{to the basestation}.
 The values are then stored in the MySQL database at the host computer.
@@
 This paper presents a decentralized autonomous sensor fault detection strategy for structural health monitoring systems using neural networks. 
-We realized the autonomous sensor fault detection by implementing a neural network into each sensor node.
-The neural networks are trained to predict the expected sensor measurement, which is then compared to the actual measurment, revealing potential sensor faults.
-The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can, in fact, improve the dependability and the accuracy of structural health monitoring systems.
+Autonomous sensor fault detection has been realized by implementing a neural network into each sensor node.
+The neural networks have been trained to predict expected sensor measurements to be compared to actual measurments, in oreder to detect sensor faults.
+The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can, improve the dependability and the accuracy of structural health monitoring systems.
 

### 2015-02-06 Paper/PaperJahrSchlich.tex

@@
 \todo{citation BILD!}
-\begin{figure}[htb]
+\begin{figure}[ht]
     \centering
@@
 
-\begin{figure}[htb]
+\begin{figure}[h!]
     \centering
-    \fbox{\includegraphics{figures/missing.jpeg}}
+    \includegraphics{figures/uml.pdf}
     \caption{class diagram describing the structure of the implemented SHM system}
@@
 To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure.
-The test structure is a 4-story shear-frame consisting of four steel plates with an area of 25\,cm\,$\times$\,50\,cm and a thickness of 0.8\,mm.
-Those plates are mounted on threaded rods with a vertical clearance of 23\,cm.
-At the bottom, the rods are fixed into a solid block of wood with an area of 40\,cm\,$\times$\,60\,cm and a height of 30\,cm.
-The SHM system is installed on the test structure by fastening wireless sensor nodes to each story.
-The laboratory setup is shown in \autoref{fig:teststructure}.
+The test structure is a 4-story shear-frame consisting of four steel plates of 25\,cm\,$\times$\,50\,cm and a thickness of 0.8\,mm.
+The plates are mounted on threaded rods with a vertical clearance of 23\,cm.
+At the bottom, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm and a height of 30\,cm.
+The SHM system is installed on the test structure by fastening a wireless sensor node to each story.
+The laboratory setup after system installation is shown in \autoref{fig:teststructure}.
+
 \todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
@@
 
-The structure is excited by deflecting and releasing the top of the structure.
 This excitation method ensures a free vibration in natural frequency with little interferences.
@@
 
-This paper presents a decentralized autonomous sensor fault detection strategy for structural health monitoring systems using neural networks. 
+This paper has shown a decentralized autonomous sensor fault detection strategy for structural health monitoring systems based on neural networks. 
 Autonomous sensor fault detection has been realized by implementing a neural network into each sensor node.
-The neural networks have been trained to predict expected sensor measurements to be compared to actual measurments, in oreder to detect sensor faults.
-The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can, improve the dependability and the accuracy of structural health monitoring systems.
+The neural networks have been trained to predict expected sensor measurements to be compared to actual measurments, in order to detect sensor faults.
+The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.
 

### 2015-02-08 SHM_JahrSchlich/sensornode/src/Communication.java
natural frequency detection fixed

@@ -1,4 +1,4 @@
-package spot;
+package sensornode;
 
 import com.sun.spot.io.j2me.radiogram.*;
 import com.sun.spot.resources.Resources;

### 2015-02-08 SHM_JahrSchlich/sensornode/src/FFT.java

@@ -1,4 +1,4 @@
-package spot;
+package sensornode;
 /**
  * @author Orlando Selenu
  * modified by Jahr&Schlich
@@ -173,6 +173,12 @@ public class FFT {
 	//get maximal magnitude -> natural frequency
     public static double[] calcNaturalFreq(double[] magnitude, double[] frequency)
     {
+    	// split array in half to get rid of double frequencies
+    	double[] newMagnitude = new double[magnitude.length/2];
+    	for(int i =0;i < newMagnitude.length;i++)
+    		newMagnitude[i] = magnitude[i];
+    	
+    	// determine maximal magnitude
 		double maxMagnitude = magnitude[0];
 		int indexMax = 0;
 		for ( int i = 1; i < magnitude.length; i++) {
@@ -181,6 +187,7 @@ public class FFT {
 		    	indexMax = i;
 		    }
 		}
+		
 		//get frequency for maximal  magnitude
 		double maxFrequency = frequency[indexMax];
 		System.out.println("maxMagnitude= "+ maxMagnitude+

### 2015-02-08 Paper/PaperJahrSchlichSHM.tex

--- 2015-02-06 Paper/PaperJahrSchlich.tex
@@
 
-\title{\textbf{ \Large{Decentralized autonomous sensor fault detection using neural networks}} }
+\title{\textbf{ \Large{Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}
 
@@
 \setlength{\emergencystretch}{3pt}
-\hyphenation{sample-rate auto-nomous}
+\hyphenation{sample-rate auto-nomous moni-toring}
 
@@
 
-The dependability and the accuracy of structural health moni\-toring systems can be affected by sensor faults. 
+Structural health monitoring systems
+shm systems produce a lot of data
+we reduce the data with fft
+The dependability and the accuracy of structural health monitoring systems can be affected by sensor faults. 
 In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
 For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
-The predictions are computed using neural networks based on measured sensor data of adjacent sensor nodes.
-In laboratory experiments, devised to validate the proposed approach, several simulated sensor faults are detected.
-These results indicate that the use of neural networks for fault detection increases the dependability and the accuracy of structural health monitoring systems.
+In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced.
 
@@
 
-Over their lifetime, sensors can become inaccurate, faulty, or may even break.
-A fault can be defined as a defect of a sensor that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
-To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
-A well known approach to fault detection is the installation of physically redundant sensors.
-Faulty sensors can be identified through the deviation of their measurements from the measurements of correlated sensors.
-Physical redundancy, although efficient for sensor fault detection, causes increased installation and maintenance costs due to the multiple installation of sensors. 
-Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors.
-Virtual sensor measurements are computed for each sensor and then compared to the actual measurements.
-For example, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor
-\citep{Smarsly2014}.
-
-In this study, analytical redundancy is implemented into wireless sensor nodes based on artificial neural networks.
-Artificial neural networks essentially consist of interconnected data processing units called neurons. 
-The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
-Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.
+\begin{itemize}
+	\item problem: viel data traffic
+	\item daten können verloren gehen
+	\item lösung: embedded computing, decentralized data analysis on sensor nodes 
+	\item fft
+\end{itemize}
 
@@
 First, a wireless structural health monitoring system is designed and implemented. 
-Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
+Next, a decentralized data analysis approach, using fast Fourier transform, is developed.
 Then, the system is tested in laboratory experiments.
@@
 The SHM system is programmed object-oriented in Java. 
-Object orientation uses objects, that are instanciated using classes as paradigm.
-A class includes methods, allowing the objects to perform actions, and attributes, specific data.
-\autoref{fig:UML} describes the classes of the SHM system code.
+Object orientation uses objects which are instanciated using classes as paradigm. 
+A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
 The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
@@
 
+\autoref{fig:UML} describes the classes of the \texttt{sensornode} package.
 The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
@@
 Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
+\todo{Grafik zweitteilen - hier package sensornode}
 
+\autoref{fig:UML} describes the classes of the \texttt{basestation} package.
 The package \texttt{basestation} runs on the host computer and operates the basestation.
@@
 The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
+\todo{Grafik zweitteilen - hier package basestation}
 
@@
 
-\section*{Implementation and training of neural networks}
+\section*{Laboratory experiments}
 
-Neural networks are implemented into the sensor nodes using SNIPE\footnote{}, a open source Java Library.
-\begin{figure}[ht]
-    \centering
-    \includegraphics{figures/neuralnetwork.pdf}
-    \caption{Schematic drawing of an artificial neural network with three layers}
-    \label{fig:neuralnetwork}
-\end{figure}
-\todo[inline]{Grafik wird an unser neuronales Netz angepasst}
-
-%\begin{itemize}
-%\item neurons
-%\item layer
-%\item weights
-%\item activation function / identity function
-%\end{itemize}
-
-%\textbf{2. paragraph: proposed nn}
-%
-%\begin{itemize}
-%\item SNIPE? implementation in java?
-%\item activation function / identity function
-%\end{itemize}
-%
-%\textbf{3. paragraph: learning general}
-%
-%\begin{itemize}
-%\item set with test values
-%\item iterative weight adjustment
-%\end{itemize}
-%
-%\textbf{4. paragraph?: learning, proposed}
-
-%----------------------------------------------------------------------------------------
-
-\newpage
-\section*{Laboratory experiments}
+In the following section, the laboratory experiments are described.
+First a describtion of the test setup is given, second, the data aquisition and processing are depicted, and finally, the results are discussed.
 
@@
 The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
-Each sensor node transfers only the values of the predominating frequency---i.e.
-the frequency with the biggest magnitude--- \colorbox{cyan}{to the basestation}.
+Each sensor node transfers only the values of the predominating frequency---i.e. the frequency with the biggest magnitude--- \colorbox{cyan}{to the basestation}. 
 The values are then stored in the MySQL database at the host computer.
+
+\begin{itemize}
+	\item 
+	\item that means our paper is also wireless and very very good
+\end{itemize}
 
@@
 
-This paper has shown a decentralized autonomous sensor fault detection strategy for structural health monitoring systems based on neural networks. 
-Autonomous sensor fault detection has been realized by implementing a neural network into each sensor node.
-The neural networks have been trained to predict expected sensor measurements to be compared to actual measurments, in order to detect sensor faults.
-The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.
+\begin{itemize}
+	\item data traffic was reduced by XX\,\% 
+\end{itemize}
 

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
Abstract angepasst, UML-Diagram geteilt, kleine Änderungen am Literaturverzeichnis

@@
 
-Structural health monitoring systems
-shm systems produce a lot of data
-we reduce the data with fft
-The dependability and the accuracy of structural health monitoring systems can be affected by sensor faults. 
-In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
-For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
-In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced.
+The dependability and accuracy of wireless structural health monitoring systems can be affected by data loss during wireless data transmission.
+Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and lower the occurrence of data loss.
+In this paper, the design and implementation of a wireless structural health monitoring system, capable of data compression is presented. 
+A fast Fourier transform is used on every sensor node to analyze oscillation frequencies. 
+Only the first natural frequency of the oscillation and the correlated magnitude are transmitted to the base station.
+In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced by xx %.
 
@@
 \modulolinenumbers[5] % nur jede 5. Zeile
+
 \section*{Dictionary}
@@
 \hline
-
 \end{tabular}
@@
 
-\todo{citation BILD!}
 \begin{figure}[ht]
@@
     \includegraphics[scale=0.5]{figures/eDemoboard.png}
-    \caption{Layout of the Sun SPOT sensor board}
+    \caption{Layout of the Sun SPOT sensor board \cite{spotToO}}
     \label{fig:SPOT}
@@
 
-\autoref{fig:UML} describes the classes of the \texttt{sensornode} package.
+\autoref{fig:UML-sn} describes the classes of the \texttt{sensornode} package.
 The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
@@
 Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
-\todo{Grafik zweitteilen - hier package sensornode}
 
-\autoref{fig:UML} describes the classes of the \texttt{basestation} package.
+\begin{figure}[h!]
+    \centering
+    \includegraphics{figures/uml-sensornode.pdf}
+    \caption{class diagram of the sensornode package}
+    \label{fig:UML-sn}
+\end{figure}
+
+\autoref{fig:UML-bs} describes the classes of the \texttt{basestation} package.
 The package \texttt{basestation} runs on the host computer and operates the basestation.
@@
     \centering
-    \includegraphics{figures/uml.pdf}
-    \caption{class diagram describing the structure of the implemented SHM system}
-    \label{fig:UML}
+    \includegraphics{figures/uml-basestation.pdf}
+    \caption{class diagram of the basestation package}
+    \label{fig:UML-bs}
 \end{figure}
@@
     \centering
-    \includegraphics[scale=0.3]{figures/teststructure.jpg}
+    \includegraphics[scale=0.4]{figures/teststructure.jpg}
     \caption{Laboratory setup}
@@
 This excitation method ensures a free vibration in natural frequency with little interferences.
-After excitation, the sensor nodes automatically start the acceleration with a sampling rate of 40\,Hz.
+After excitation, the sensor nodes automatically start measuring acceleration with a sampling rate of 40\,Hz.
 To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.
-%Generally FFTs are used to convert equidistant samples of a function into a combination of frequencies, and the corresponding magnitudes, that has the same sample values. 
+
 The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
@@
 \end{itemize}
-
-%\textbf{2. paragraph: data measurement and processing}
-%
-%\emph{\begin{itemize}
-%\item learning phase
-%\item excitation
-%\item sensors start measuring
-%\item sensor nodes \textbf{perform fft} - what is fft
-%\item sensor nodes exchange frequencies
-%\item neural networks check integrity
-%\item if no faults detected, sensor nodes send data to basestation
-%\end{itemize}}
-%
-%\textbf{3. paragraph: sensor fault detection}
-%
-%\begin{itemize}
-%\item simulation of sensor fault
-%\item neural networks check integrity -> error!
-%\item sensor node sends alert to basestation
-%\end{itemize}
 

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
data transmission + FFT

--- 2015-02-08 Paper/PaperJahrSchlichSHM.tex
@@
 Traditionally, the sensor nodes are connected to computer systems with cables.
-However, using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
+Using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
 In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
 
-\begin{itemize}
-	\item problem: viel data traffic
-	\item daten können verloren gehen
-	\item lösung: embedded computing, decentralized data analysis on sensor nodes 
-	\item fft
-\end{itemize}
+However, when using wireless SHM systems, special attention has to be paid to data transfer.
+Immense amounts of raw data incure and have to be transmitted.
+Over the air transmission is error-prone and may lead to irreparable data loss.  (klingt das zu stark? Warum ist das mit FFT nicht mehr so schlimm?)
+To save transmission time and storage space, and to decrease data loss, the transmidded data can be reduced using embedded computing.
+Depending on the measuring objective, first steps of data processing can be conducted decentralized directly on the sensor nodes.
+Transmitting analysis results instead of raw data can reduce the amount of transmitted data exponentially.
+
+In this paper, the natural frequency of a building structure is determined using a fast Fourier transform.
+Deviations from the natural frequency indicate changes in the vibration behavior, which may be caused by structural failures.
+
+%\begin{itemize}
+%	\item problem: viel data traffic
+%	\item daten können verloren gehen
+%	\item lösung: embedded computing, decentralized data analysis on sensor nodes 
+%	\item fft
+%\end{itemize}
 
@@
 
-\autoref{fig:UML} describes the classes of the \texttt{sensornode} package.
-The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
-The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
+\autoref{fig:UML} describes the classes of the \texttt{sensor\-node} package.
+The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, which are embedded directly into the sensor nodes.
+The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
 There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
@@
 Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
-The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
-For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
-Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
+The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
+For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
+Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
 \todo{Grafik zweitteilen - hier package sensornode}
 
-\autoref{fig:UML} describes the classes of the \texttt{basestation} package.
-The package \texttt{basestation} runs on the host computer and operates the basestation.
-It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
+\autoref{fig:UML} describes the classes of the \texttt{base\-station} package.
+The package \texttt{base\-station} runs on the host computer and operates the basestation.
+It consists of the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
 The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
-The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class.
-The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
+The entry point of the program is the \texttt{run()} method in the \texttt{Main\-Base} class.
+The \texttt{Main\-Base} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
 \todo{Grafik zweitteilen - hier package basestation}
@@
     \centering
-    \includegraphics{figures/uml.pdf}
+    \includegraphics[width = \textwidth]{figures/uml.pdf}
     \caption{class diagram describing the structure of the implemented SHM system}

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex

@@
 
-\baselineskip20pt
+\baselineskip18pt
 
@@
 
-\newpage
-
 \section*{Design and implementation of the wireless structural health monitoring system}
@@
 
-\autoref{fig:UML} describes the classes of the \texttt{sensor\-node} package.
+\autoref{fig:UML-sn} describes the classes of the \texttt{sensor\-node} package.
 The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, which are embedded directly into the sensor nodes.
@@
 Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
-\todo{Grafik zweitteilen - hier package sensornode}
 
-\autoref{fig:UML} describes the classes of the \texttt{base\-station} package.
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
+    \caption{class diagram of the sensornode package}
+    \label{fig:UML-sn}
+\end{figure}
+
+\autoref{fig:UML-bs} describes the classes of the \texttt{base\-station} package.
 The package \texttt{base\-station} runs on the host computer and operates the basestation.
 It consists of the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
+
 The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
@@
 The \texttt{Main\-Base} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
-\todo{Grafik zweitteilen - hier package basestation}
 
@@
     \centering
-    \includegraphics[width = \textwidth]{figures/uml.pdf}
-    \caption{class diagram describing the structure of the implemented SHM system}
-    \label{fig:UML}
+    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
+    \caption{class diagram of the basestation package}
+    \label{fig:UML-bs}
 \end{figure}

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
abstract

@@
 \usepackage{lineno}
+
+\makeatletter
+\renewcommand\section{\@startsection{section}{1}{\z@}%
+                     {-3.5ex \@plus -1ex \@minus -.2ex}%
+                     {2.3ex \@plus.2ex}%
+                     {\normalfont\large\bfseries}}
+\makeatother
+
+%\usepackage{titlesec}
+%\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
 
@@
 
-\renewcommand\refname{References and Notes}
+%\renewcommand\refname{References and Notes}
 
@@
 
-\title{\textbf{ \Large{Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}
+\title{\normalfont\textbf{ \Large {Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}
 
 \author
-{Katrin Jahr, Robert Schlich\\
+{\large Katrin Jahr, Robert Schlich\\
 \\
@@
 \\
-\normalsize{katrin.jahr@uni-weimar.de}\\
+\normalsize{katrin.jahr@uni-weimar.de, }
 \normalsize{robert.schlich@uni-weimar.de}
@@
 
-\baselineskip18pt
+\baselineskip20pt
 
@@
 
-The dependability and accuracy of wireless structural health monitoring systems can be affected by data loss during wireless data transmission.
-Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and lower the occurrence of data loss.
-In this paper, the design and implementation of a wireless structural health monitoring system, capable of data compression is presented. 
-A fast Fourier transform is used on every sensor node to analyze oscillation frequencies. 
-Only the first natural frequency of the oscillation and the correlated magnitude are transmitted to the base station.
-In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced by xx %.
+Immense amounts of data incure during wireless structural health monitoring.
+Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and stored.
+In this paper, the design and implementation of a wireless structural health monitoring system, capable of dezentralized data compression, is presented. 
+A fast Fourier transform is embedded into each sensor node to analyze oscillation frequencies. 
+By only transmitting the first natural frequency and the correlating magnitude, the data stream is reduced to two values.
+The proposed approach was validated in laboratory experiments.
+
+%The dependability and the accuracy of wireless structural health monitoring systems can be affected by data loss during wireless data transmission.
+%Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and lower the occurrence of data loss.
+%In this paper, the design and implementation of a wireless structural health monitoring system, capable of data compression, is presented. 
+%A fast Fourier transform is used on every sensor node to analyze oscillation frequencies. 
+%Only the first natural frequency of the oscillation and the correlated magnitude are transmitted to the base station.
+%In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced by xx.
 
@@
 \modulolinenumbers[5] % nur jede 5. Zeile
-
-\section*{Dictionary}
-
-\begin{tabular}{|l|l|}
-\hline 
-Sensorknoten (SunSPOT) & sensor node \\ 
-\hline 
-einzelner Messsensor (Thermometer) & sensor \\ 
-\hline 
-Knoten im neuronalen Netz & neuron \\ 
-\hline 
-eine abgeschlossene Messungreihe & test run \\ 
-\hline 
-gemessene Werte & sensor data \\ 
-\hline 
-vorhergesagte Werte & predicted data \\ 
-\hline 
-durch Vorhersage erwartete Werte & expected data \\ 
-\hline 
-einzelner Messwert & measurement \\ 
-\hline 
-Test & laboratory experiments\\ 
-\hline
-tatsächliche, nicht virtuelle Messung & actual measurement \\
-\hline
-Messaufbau & test setup \\
-\hline
-\end{tabular}
+%
+%\section*{Dictionary}
+%
+%\begin{tabular}{|l|l|}
+%\hline 
+%Sensorknoten (SunSPOT) & sensor node \\ 
+%\hline 
+%einzelner Messsensor (Thermometer) & sensor \\ 
+%\hline 
+%Knoten im neuronalen Netz & neuron \\ 
+%\hline 
+%eine abgeschlossene Messungreihe & test run \\ 
+%\hline 
+%gemessene Werte & sensor data \\ 
+%\hline 
+%vorhergesagte Werte & predicted data \\ 
+%\hline 
+%durch Vorhersage erwartete Werte & expected data \\ 
+%\hline 
+%einzelner Messwert & measurement \\ 
+%\hline 
+%Test & laboratory experiments\\ 
+%\hline
+%tatsächliche, nicht virtuelle Messung & actual measurement \\
+%\hline
+%Messaufbau & test setup \\
+%\hline
+%\end{tabular}
 
@@
 Over the air transmission is error-prone and may lead to irreparable data loss.  (klingt das zu stark? Warum ist das mit FFT nicht mehr so schlimm?)
-To save transmission time and storage space, and to decrease data loss, the transmidded data can be reduced using embedded computing.
+To save transmission time and storage space, and to decrease data loss, the transmitted data can be reduced using embedded computing.
 Depending on the measuring objective, first steps of data processing can be conducted decentralized directly on the sensor nodes.
@@
 
-\newpage
+%\newpage
 

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
Extended section 'Laboratory Experiments'

--- 2015-02-09 Paper/PaperJahrSchlichSHM.tex
@@
 \usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
-
+\usepackage{titlesec}
 \usepackage{color}
@@
 
+The structure is excited by deflecting and releasing the top of the structure.
 This excitation method ensures a free vibration in natural frequency with little interferences.
-After excitation, the sensor nodes automatically start measuring acceleration with a sampling rate of 40\,Hz.
+After excitation, the sensor nodes automatically start measuring acceleration with an increased sampling rate of 50\,Hz.
 To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.
-
 The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
@@
 The values are then stored in the MySQL database at the host computer.
+
+The correct implementation of the FFT has been verified beforehand by plotting the frequency domain of various oscillation events.
+\autoref{fig:fft-diagram} shows  one of the results as an example.
+The natural frequencies are readily identifiable by the peaks in the diagram.
+The first three natural frequencies are located at approximately 2.3\,Hz (first natural frequency), 8\,Hz (second natural frequency),and 16\,Hz (third natural frequency).
+The same natural frequencies appeared throughout repeated tests with varying excitation forces, sampling rates, and number of measurements.
+
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{c c c}
+		1\textsuperscript{st} natural frequency & 2\textsuperscript{nd} natural frequency & 3\textsuperscript{rd} natural frequency \\ 
+		2,3 & 8 & 16 \\
+	\end{tabular}
+	\caption{Natural frequencies of the test structure}
+	\label{tab:nat-freqs}
+\end{table}
 

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
Abgleich

@@
 
+\clubpenalty = 10000 % schliesst Schusterjungen aus
+\widowpenalty = 10000 \displaywidowpenalty = 10000% schliesst Hurenkinder aus
+
 %The next command sets up an environment for the abstract to your paper.
@@
 {\setlength{\leftmargin}{.22in}}
-{\setlength{\labelsep}{.5em}}}
+%{\setlength{\labelsep}{.5em}}
+}
 {\end{list}}
@@
 A fast Fourier transform is embedded into each sensor node to analyze oscillation frequencies. 
-By only transmitting the first natural frequency and the correlating magnitude, the data stream is reduced to two values.
+By only transmitting the first natural frequency and the correlating magnitude, the data stream is reduced to two values per sampling event.
 The proposed approach was validated in laboratory experiments.
@@
 Immense amounts of raw data incure and have to be transmitted.
-Over the air transmission is error-prone and may lead to irreparable data loss.  (klingt das zu stark? Warum ist das mit FFT nicht mehr so schlimm?)
-To save transmission time and storage space, and to decrease data loss, the transmitted data can be reduced using embedded computing.
+Over the air transmission is error-prone and may lead to data loss.
+To save transmission time and storage space, and to decrease the probability data loss, the transmitted data can be reduced using embedded computing.
 Depending on the measuring objective, first steps of data processing can be conducted decentralized directly on the sensor nodes.
@@
 Deviations from the natural frequency indicate changes in the vibration behavior, which may be caused by structural failures.
+\todo[inline]{noch ein Satz wäre schön: nur übertragung von einem Wert...}
 
@@
 
-\begin{figure}[ht]
+\begin{figure}[htb]
     \centering
@@
 
-\colorbox{cyan}{neuer Absatz:}
-
 The SHM system is programmed object-oriented in Java. 
@@
 The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
-There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
-Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
+There are two phases: At first, the acceleration is measured with a low sampling rate.
+Once the acceleration exceeds a threshold, the second phase is entered by increasing the sampling rate. 
 The measured values are stored into an array.
@@
 With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
-Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
+Finally, the natural frequency is determined by extracting the maximal magnitude.
 The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
-For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
+For starting the operation of the sensor node, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
 Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
@@
 It consists of the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
-
 The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
@@
 
-\todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
+%\todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
 
@@
 The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
-Each sensor node transfers only the values of the predominating frequency---i.e. the frequency with the biggest magnitude--- \colorbox{cyan}{to the basestation}. 
+Each sensor node transfers only the values of the predominating frequency---i.e. the frequency with the biggest magnitude---to the basestation. 
 The values are then stored in the MySQL database at the host computer.

### 2015-02-09 Paper/PaperJahrSchlichSHM.tex
Reworked all figures to include latex's sans-serif-font

@@
 \usepackage{color}
+\usepackage{booktabs}
 
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht]
     \centering
-    \includegraphics[scale=0.4]{figures/teststructure.jpg}
+    \includegraphics[scale=0.6]{figures/testsetup.png}
     \caption{Laboratory setup}
@@
 The correct implementation of the FFT has been verified beforehand by plotting the frequency domain of various oscillation events.
-\autoref{fig:fft-diagram} shows  one of the results as an example.
+\autoref{fig:50hzfft} shows  one of the results as an example.
 The natural frequencies are readily identifiable by the peaks in the diagram.
-The first three natural frequencies are located at approximately 2.3\,Hz (first natural frequency), 8\,Hz (second natural frequency),and 16\,Hz (third natural frequency).
-The same natural frequencies appeared throughout repeated tests with varying excitation forces, sampling rates, and number of measurements.
+The first three natural frequencies are located at approximately 2.3\,Hz (1\textsuperscript{st} natural frequency), 8\,Hz (2\textsuperscript{nd} natural frequency),and 18\,Hz (3\textsuperscript{rd} natural frequency).
+The same natural frequencies appeared throughout repeated tests with varying excitation forces, sampling rates, and number of measured acceleration values.
 
-\begin{table}[ht]
+\begin{figure}[ht]
     \centering
-	\begin{tabular}{c c c}
-		1\textsuperscript{st} natural frequency & 2\textsuperscript{nd} natural frequency & 3\textsuperscript{rd} natural frequency \\ 
-		2,3 & 8 & 16 \\
-	\end{tabular}
-	\caption{Natural frequencies of the test structure}
-	\label{tab:nat-freqs}
-\end{table}
+    \includegraphics{figures/50Hz.pdf}
+    \caption{frequency domain graph of the oscillation}
+    \label{fig:50hzfft}
+\end{figure}
 
-\begin{itemize}
-	\item 
-	\item that means our paper is also wireless and very very good
-\end{itemize}
+After the
+
+%\begin{table}[ht]
+%	\centering
+%	\begin{tabular}{c c c}
+%		\toprule
+%		1\textsuperscript{st} natural frequency & 2\textsuperscript{nd} natural frequency & 3\textsuperscript{rd} natural frequency \\ 
+%		\midrule
+%		2,3 & 8 & 16 \\
+%		\bottomrule
+%	\end{tabular}
+%	\caption{Natural frequencies of the test structure}
+%	\label{tab:nat-freqs}
+%\end{table}
 

### 2015-02-10 Paper/PaperJahrSchlichSHM.tex
summary

--- 2015-02-09 Paper/PaperJahrSchlichSHM.tex
@@
 \\
-\normalsize{katrin.jahr@uni-weimar.de, }
+\normalsize{katrin.jahr@uni-weimar.de \qquad}
 \normalsize{robert.schlich@uni-weimar.de}
@@
 Among others, the sensor board includes an accelerometer and eight independent RGB-LEDs.
-The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125Hz \citep{eDemo2010}.
+The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz \citep{eDemo2010}.
 
@@
 
-\begin{itemize}
-	\item data traffic was reduced by XX\,\% 
-\end{itemize}
+In structural health monitoring, a decentralized data compression stragegy reduces data traffic and storage space.
+Using a fast Fourier transform to determine the natural frequency allows conclusions about the structural integrity of a structure.
+The transmitted data was reduced exponentially, saving up to xx MB per sampling event.
 

### 2015-02-10 SHM_JahrSchlich/basestation/src/DatabaseHandler.java

@@ -45,15 +45,19 @@ public class DatabaseHandler {
 					+ "Time TIMESTAMP, "
 					+ "SpotID VARCHAR(19), "
 					+ "Magnitude DOUBLE, "
-					+ "Frequency FLOAT)");
+					+ "Frequency FLOAT, "
+					+ "Samplerate INTEGER)");
 			statement.close();
 		}
 		resultSet.close();
 	}
 	
 	
-	public void writeMeasurement(String tableName, String spotID, 
-						double magnitude, float frequency) throws Exception {
+	public void writeMeasurement(String tableName, 
+								String spotID, 
+								double magnitude, 
+								float frequency, 
+								int samplerate) throws Exception {
 		
 		
 		//Get current systemtime as timestamp to avoid trouble with unsynchronized SunSPOTs
@@ -63,11 +67,12 @@ public class DatabaseHandler {
 		System.out.println("Writing measurement into Database");
 		PreparedStatement preparedStatement = dbConn.prepareStatement("INSERT INTO "
 				+ tableName
-				+ " (Time, SpotID, Magnitude, Frequency) VALUES (?,?,?,?)");
+				+ " (Time, SpotID, Magnitude, Frequency, Samplerate) VALUES (?,?,?,?,?)");
 			preparedStatement.setTimestamp(1, now);
 			preparedStatement.setString(2, spotID);
 			preparedStatement.setDouble(3, magnitude);
 			preparedStatement.setFloat(4, frequency);
+			preparedStatement.setFloat(5, samplerate);
 			preparedStatement.execute();
 			System.out.println(preparedStatement);
 		preparedStatement.close();

### 2015-02-10 SHM_JahrSchlich/basestation/src/MainBase.java

@@ -26,6 +26,7 @@ public class MainBase {
     private static final String DATABASE_USER = "shm";
     private static final String DATABASE_PASSWORD = "pass";
     private static final String TABLE_NAME = "measurements";
+    private int samplerate = 76;  // in milliseconds
     
     private void run() throws Exception {
     	
@@ -64,7 +65,7 @@ public class MainBase {
                 double magnitude = dg.readDouble();      // read the magnitude
                 float frequency = dg.readFloat();        // read the frequency
                 // Write measurement to Database
-                dbHandler.writeMeasurement(TABLE_NAME, addr, magnitude, frequency);
+                dbHandler.writeMeasurement(TABLE_NAME, addr, magnitude, frequency, samplerate);
             } catch (Exception e) {
                 System.err.println("Caught " + e +  " while reading sensor samples.");
                 throw e;

### 2015-02-10 SHM_JahrSchlich/sensornode/src/FFT.java

@@ -190,8 +190,8 @@ public class FFT {
 		
 		//get frequency for maximal  magnitude
 		double maxFrequency = frequency[indexMax];
-		System.out.println("maxMagnitude= "+ maxMagnitude+
-											", maxFrequency= " + maxFrequency);
+		//System.out.println("maxMagnitude= "+ maxMagnitude+
+		//									", maxFrequency= " + maxFrequency);
 		
 		double naturalFreq[] = {maxMagnitude, maxFrequency};
     		

### 2015-02-10 SHM_JahrSchlich/sensornode/src/MainSpot.java

@@ -9,16 +9,17 @@ import javax.microedition.midlet.MIDlet;
 import javax.microedition.midlet.MIDletStateChangeException;
 import com.sun.spot.resources.transducers.IAccelerometer3D;
 import com.sun.spot.service.BootloaderListenerService;
+//import java.io.Writer;
 
 
 public class MainSpot extends MIDlet {
 
     private int HOST_PORT = 67;
-    private int SAMPLE_PERIOD_LISTENING = 1 * 1000;  // in milliseconds
-    private int SAMPLE_PERIOD_MEASURING = 100;  // in milliseconds
-    private int ARRAY_LENGTH = 128;
+    private int SAMPLE_PERIOD_LISTENING = 1 * 500;  // in milliseconds
+    private int SAMPLE_PERIOD_MEASURING = 13;  // in milliseconds
+    private int ARRAY_LENGTH = 512;
     private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
-    private double THRESHOLD = 0.1;
+    private double THRESHOLD = 0.2;
     
     protected void startApp() throws MIDletStateChangeException {
 
@@ -45,6 +46,7 @@ public class MainSpot extends MIDlet {
 			double[] frequency = FFT.calcFreq(transform, SAMPLERATE);
 			double[] naturalFreq =  FFT.calcNaturalFreq(magnitude, frequency);
 			
+			
 			// print frequency and magnitude array on console
 			System.out.println("Magnitude");
 			for (int i = 0 ; i < ARRAY_LENGTH; i++){
@@ -54,8 +56,30 @@ public class MainSpot extends MIDlet {
 			for (int i = 0 ; i < ARRAY_LENGTH; i++){
 				System.out.println(frequency[i]);
 			}
-
-			//send processed data to database
+			System.out.println("natural magnitude: " + naturalFreq[0]);
+			System.out.println("natural Frequency: " + naturalFreq[1]);
+			
+			
+			/*
+			
+			// print data to file
+			Writer.PrintWriter writer = new Writer.PrintWriter("measurement.txt", "UTF-8");
+			
+			writer.println("Magnitude");
+			for (int i = 0 ; i < ARRAY_LENGTH; i++){
+				writer.println(magnitude[i]);
+			}	
+			writer.println("Frequency");
+			for (int i = 0 ; i < ARRAY_LENGTH; i++){
+				writer.println(frequency[i]);
+			}
[…]

### 2015-02-10 Paper/PaperJahrSchlichSHM.tex
finished lab ex

@@
 \documentclass[12pt,a4paper]{scrartcl}
-\usepackage[ngerman, english]{babel}
+\usepackage[english]{babel}
 \usepackage[T1]{fontenc}
@@
 
-\title{\normalfont\textbf{ \Large {Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}
+\title{\textsf\textbf{ \Large {Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}
 
@@
 This excitation method ensures a free vibration in natural frequency with little interferences.
-After excitation, the sensor nodes automatically start measuring acceleration with an increased sampling rate of 50\,Hz.
+After excitation, the sensor nodes automatically start measuring acceleration with an increased sampling rate.
 To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.
 The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
-Each sensor node transfers only the values of the predominating frequency---i.e. the frequency with the biggest magnitude---to the basestation. 
+Each sensor node transfers only the values of the first natural frequency---i.e. the frequency with the biggest magnitude---to the basestation. 
 The values are then stored in the MySQL database at the host computer.
@@
 \autoref{fig:50hzfft} shows  one of the results as an example.
-The natural frequencies are readily identifiable by the peaks in the diagram.
+The natural frequencies are readily identifiable by the peaks in the graph.
 The first three natural frequencies are located at approximately 2.3\,Hz (1\textsuperscript{st} natural frequency), 8\,Hz (2\textsuperscript{nd} natural frequency),and 18\,Hz (3\textsuperscript{rd} natural frequency).
 The same natural frequencies appeared throughout repeated tests with varying excitation forces, sampling rates, and number of measured acceleration values.
+The chosen sensor nodes were unable to handle sampling rates above 76\,Hz.
+Furthermore test runs with more than 512 measured values showed no significant increase in precision.
+Therefore a sampling rate of 76\,Hz and 512 measured values were assumed reasonable for the following test runs.
 
@@
 
-After the
+In a next step multiple sensor nodes are installed to different positions on the test structure in order to test the SHM system as a whole.
+After the sensors are installed, the test structure was subjected to varying degrees of excitation.
+The calculated natural frequencies, and corresponding magnitude, are safed in the MySQL database and analyzed manually.
+The results of these test runs are shown in \autoref{tab:lab-ex} in the appendix.
 
-%\begin{table}[ht]
-%	\centering
-%	\begin{tabular}{c c c}
-%		\toprule
-%		1\textsuperscript{st} natural frequency & 2\textsuperscript{nd} natural frequency & 3\textsuperscript{rd} natural frequency \\ 
-%		\midrule
-%		2,3 & 8 & 16 \\
-%		\bottomrule
-%	\end{tabular}
-%	\caption{Natural frequencies of the test structure}
-%	\label{tab:nat-freqs}
-%\end{table}
+Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
+The sensors installed on the top of the structure always calculates the highest corresponding magnitude and magnitudes decreased from top to bottom.
+These results are plausible and indicate, that the SHM system is working correctly.
+
+Through FFT analysis of oscillations the data traffic is decreased from 512 measured acceleration values per oscillation to two values, the natural frequency and the corresponding magnitude.
+A number stored in the used data format double occupies 8\,bytes of space.
+Therefore the data traffic was decreased by $510\times8\,B = 4080\,B = 3,98\,MB$ per sampling event
+This equates a 99.6\,\% decrease of data traffic.
 
@@
 Using a fast Fourier transform to determine the natural frequency allows conclusions about the structural integrity of a structure.
-The transmitted data was reduced exponentially, saving up to xx MB per sampling event.
+The transmitted data was reduced exponentially, saving up to 3,96\,MB per sampling event.
 

### 2015-02-10 Paper/PaperJahrSchlichSHM.tex
Dataflow diagram, layout

--- 2015-02-10 Paper/PaperJahrSchlichSHM.tex
@@
 \renewcommand\section{\@startsection{section}{1}{\z@}%
-                     {-3.5ex \@plus -1ex \@minus -.2ex}%
+                     {-7.5ex \@plus -1ex \@minus -10.2ex}%
                      {2.3ex \@plus.2ex}%
-                     {\normalfont\large\bfseries}}
+                     {\sffamily\large\bfseries}}
 \makeatother
 
-%\usepackage{titlesec}
-%\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
+\usepackage{geometry}
+\geometry{a4paper, top=30mm, left=30mm, right=30mm, bottom=40mm,
+headsep=10mm, footskip=12mm}
 
-% The following parameters seem to provide a reasonable page setup.
-\topmargin -1.0cm
-\oddsidemargin 0.2cm
-\textwidth 15cm 
-\textheight 23cm
-\footskip 1.0cm
+%% The following parameters seem to provide a reasonable page setup.
+%\topmargin -1.0cm
+%\oddsidemargin 0.2cm
+%\textwidth 15cm 
+%\textheight 23cm
+%\footskip 1.0cm
 
-\clubpenalty = 10000 % schliesst Schusterjungen aus
-\widowpenalty = 10000 \displaywidowpenalty = 10000% schliesst Hurenkinder aus
-
-%The next command sets up an environment for the abstract to your paper.
-\newenvironment{sciabstract}{%
-\begin{quote} \bfseries}
-{\end{quote}}
-
-% If your reference list includes text notes as well as references,
-% include the following line; otherwise, comment it out.
-
-%\renewcommand\refname{References and Notes}
-
-% The following lines set up an environment for the last note in the
-% reference list, which commonly includes acknowledgments of funding,
-% help, etc.  It's intended for users of BibTeX or the {thebibliography}
-% environment.  Users who are hand-coding their references at the end
-% using a list environment such as {enumerate} can simply add another
-% item at the end, and it will be numbered automatically.
-
-\newcounter{lastnote}
-\newenvironment{scilastnote}{%
-\setcounter{lastnote}{\value{enumiv}}%
-\addtocounter{lastnote}{+1}%
-\begin{list}%
-{\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
-%{\setlength{\labelsep}{.5em}}
-}
-{\end{list}}
-
-% Include your paper's title here
-
-\title{\normalfont\textbf{ \Large {Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}
-
+\title{\textbf{ \Large {Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}
 \author
@@
 }
-
-% Include the date command, but leave its argument blank.
 \date{}
 
-%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
+\clubpenalty = 10000 % schliesst Schusterjungen aus
+\widowpenalty = 10000 \displaywidowpenalty = 10000% schliesst Hurenkinder aus
+
+\newenvironment{sciabstract}{%
+\begin{quote} \itshape}
+{\end{quote}}
 
@@
 \maketitle 
+\vspace{-2em}
 
@@
 \setlength{\emergencystretch}{3pt}
-\hyphenation{sample-rate auto-nomous moni-toring}
+\hyphenation{sample-rate auto-nomous moni-toring methods}
 
 \begin{sciabstract}
-
 Immense amounts of data incure during wireless structural health monitoring.
@@
 
-%The dependability and the accuracy of wireless structural health monitoring systems can be affected by data loss during wireless data transmission.
-%Applying adequate data compression methods can significantly reduce the amount of data that has to be transmitted and lower the occurrence of data loss.
-%In this paper, the design and implementation of a wireless structural health monitoring system, capable of data compression, is presented. 
-%A fast Fourier transform is used on every sensor node to analyze oscillation frequencies. 
-%Only the first natural frequency of the oscillation and the correlated magnitude are transmitted to the base station.
-%In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced by xx.
-
 \end{sciabstract}
-
-%----------------------------------------------------------------------------------------
-
-\linenumbers % Schaltet Zeilennummerierung ein
-\modulolinenumbers[5] % nur jede 5. Zeile
-%
-%\section*{Dictionary}
-%
-%\begin{tabular}{|l|l|}
-%\hline 
-%Sensorknoten (SunSPOT) & sensor node \\ 
-%\hline 
-%einzelner Messsensor (Thermometer) & sensor \\ 
-%\hline 
-%Knoten im neuronalen Netz & neuron \\ 
-%\hline 
-%eine abgeschlossene Messungreihe & test run \\ 
-%\hline 
-%gemessene Werte & sensor data \\ 
-%\hline 
-%vorhergesagte Werte & predicted data \\ 
-%\hline 
-%durch Vorhersage erwartete Werte & expected data \\ 
-%\hline 
-%einzelner Messwert & measurement \\ 
-%\hline 
-%Test & laboratory experiments\\ 
-%\hline
-%tatsächliche, nicht virtuelle Messung & actual measurement \\
-%\hline
-%Messaufbau & test setup \\
-%\hline
-%\end{tabular}
 
@@
 6. information retrieval.
+The dataflow is illustrated in \autoref{fig:flow}.
 Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a fast Fourier transform to determine the characteristics of the oscillation of the structure.
@@
 Task 5 and 6---additional analysis and diagnosis---are conducted on the host computer in further steps.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics{figures/dataflow_tasks.eps}
+    \caption{Dataflow and associated SHM tasks}
+    \label{fig:flow}
+\end{figure}
 
@@
     \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
-    \caption{class diagram of the sensornode package}
+    \caption{Class diagram of the \texttt{sensornode} package}
     \label{fig:UML-sn}
@@
     \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
-    \caption{class diagram of the basestation package}
+    \caption{Class diagram of the \texttt{basestation} package}
     \label{fig:UML-bs}
@@
 
-%\begin{table}[ht]
-%	\centering
-%	\begin{tabular}{c c c}
-%		\toprule
-%		1\textsuperscript{st} natural frequency & 2\textsuperscript{nd} natural frequency & 3\textsuperscript{rd} natural frequency \\ 
-%		\midrule
-%		2,3 & 8 & 16 \\
-%		\bottomrule
-%	\end{tabular}
-%	\caption{Natural frequencies of the test structure}
-%	\label{tab:nat-freqs}
-%\end{table}
+\begin{table}[ht]
+	\centering
+	\begin{tabular}{l l c c c}
+		\toprule
+		ID & Position & Magnitude & Natural frequency & Sampling rate\\ 
+		\midrule
+		1 & top & 3.78 & 2.2461 & 50\\ 
+		1 & middle & 3.23 & 2.2461 & 50\\ 
+		1 & bottom & 1.97 & 2.2461 & 50\\ 
+		\midrule
+		2 & top & 4.57 & 2.2461 & 50\\ 
+		2 & middle & 3.54 & 2.2461 & 50\\
+		2 & bottom & 2.43 & 2.2461 & 50\\ 
+		\midrule
+		3 & top & 8.20 & 2.2536 & 76\\
+		3 & middle & 6.35 & 2.2536 & 76\\ 
+		3 & bottom & 3.79 & 2.2536 & 76\\
+		\midrule	
+		4 & top & 5.74 & 2.2536 & 76\\  
+		4 & middle & 4.38 & 2.2536 & 76\\ 
+		4 & bottom & 2.55 & 2.2536 & 76\\
+		\midrule		
+		5 & top & 5.53 & 2.2536 & 76\\  
+		5 & middle & 4.62 & 2.2536 & 76\\ 
+		5 & bottom & 2.74 & 2.2536 & 76\\ 		
+		\bottomrule
+	\end{tabular}
+	\caption{Natural frequencies of the test structure}
+	\label{tab:lab-ex}
+\end{table}
 

### 2015-02-10 Paper/PaperJahrSchlichSHM.tex

@@
 headsep=10mm, footskip=12mm}
-
-%% The following parameters seem to provide a reasonable page setup.
-%\topmargin -1.0cm
-%\oddsidemargin 0.2cm
-%\textwidth 15cm 
-%\textheight 23cm
-%\footskip 1.0cm
 
@@
 
-%\todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
-
 \begin{figure}[ht]
@@
 \end{figure}
-
-After the
 
@@
 
+In a next step multiple sensor nodes are installed to different positions on the test structure in order to test the SHM system as a whole.
+After the sensors are installed, the test structure was subjected to varying degrees of excitation.
+The calculated natural frequencies, and corresponding magnitude, are safed in the MySQL database and analyzed manually.
+The results of these test runs are shown in \autoref{tab:lab-ex} in the appendix.
+
+Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
+The sensors installed on the top of the structure always calculates the highest corresponding magnitude and magnitudes decreased from top to bottom.
+These results are plausible and indicate, that the SHM system is working correctly.
+
+Through FFT analysis of oscillations the data traffic is decreased from 512 measured acceleration values per oscillation to two values, the natural frequency and the corresponding magnitude.
+A number stored in the used data format double occupies 8\,bytes of space.
+Therefore the data traffic was decreased by $510\times8\,B = 4080\,B = 3,98\,MB$ per sampling event
+This equates a 99.6\,\% decrease of data traffic.
+
 %----------------------------------------------------------------------------------------

### 2015-02-10 Paper/PaperJahrSchlichSHM.tex
Korrekturen

--- 2015-02-10 Paper/PaperJahrSchlichSHM.tex
@@
 Deviations from the natural frequency indicate changes in the vibration behavior, which may be caused by structural failures.
-\todo[inline]{noch ein Satz wäre schön: nur übertragung von einem Wert...}
-
-%\begin{itemize}
-%	\item problem: viel data traffic
-%	\item daten können verloren gehen
-%	\item lösung: embedded computing, decentralized data analysis on sensor nodes 
-%	\item fft
-%\end{itemize}
 
@@
 First, a wireless structural health monitoring system is designed and implemented. 
-Next, a decentralized data analysis approach, using fast Fourier transform, is developed.
+Next, a decentralized data analysis approach, using fast Fourier transform (FFT), is developed.
 Then, the system is tested in laboratory experiments.
@@
 The sensor nodes and the basestation are of type "Oracle Sun SPOT". 
-The Sun SPOTS are equipped with several components, see \autoref{fig:SPOT}.
+The Sun SPOTS are equipped with several components.
 Among others, the sensor board includes an accelerometer and eight independent RGB-LEDs.
 The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz \citep{eDemo2010}.
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[scale=0.5]{figures/eDemoboard.png}
-    \caption{Layout of the Sun SPOT sensor board \cite{spotToO}}
-    \label{fig:SPOT}
-\end{figure}
 
@@
 The dataflow is illustrated in \autoref{fig:flow}.
-Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a fast Fourier transform to determine the characteristics of the oscillation of the structure.
+Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a FFT to determine the characteristics of the oscillation of the structure.
 The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
@@
 The different phases are indicatad by lighting different LEDs.
-The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
+The \texttt{FFT} class performs a FFT on the measured accelerations. 
 With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
@@
 In the following section, the laboratory experiments are described.
-First a describtion of the test setup is given, second, the data aquisition and processing are depicted, and finally, the results are discussed.
+First, a description of the test setup is given, second, the data aquisition and processing are depicted, and finally, the results are discussed.
 
@@
 At the bottom, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm and a height of 30\,cm.
-The SHM system is installed on the test structure by fastening a wireless sensor node to each story.
+The SHM system is installed on the test structure by fastening a wireless sensor node to the top three stories.
 The laboratory setup after system installation is shown in \autoref{fig:teststructure}.
-
-%\todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
 
@@
     \centering
-    \includegraphics[scale=0.6]{figures/testsetup.png}
+    \includegraphics[scale=0.4]{figures/testsetup.png}
     \caption{Laboratory setup}
@@
 This excitation method ensures a free vibration in natural frequency with little interferences.
-After excitation, the sensor nodes automatically start measuring acceleration with an increased sampling rate of 50\,Hz.
-To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.
-The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
-Each sensor node transfers only the values of the predominating frequency---i.e. the frequency with the biggest magnitude---to the basestation. 
-The values are then stored in the MySQL database at the host computer.
+After excitation, when the acceleration threshold is exceeded, the sensor nodes automatically start measuring the acceleration.
+To minimize the wireless data traffic, each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
+By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building.
 
-The correct implementation of the FFT has been verified beforehand by plotting the frequency domain of various oscillation events.
-\autoref{fig:50hzfft} shows  one of the results as an example.
-The natural frequencies are readily identifiable by the peaks in the diagram.
-The first three natural frequencies are located at approximately 2.3\,Hz (1\textsuperscript{st} natural frequency), 8\,Hz (2\textsuperscript{nd} natural frequency),and 18\,Hz (3\textsuperscript{rd} natural frequency).
-The same natural frequencies appeared throughout repeated tests with varying excitation forces, sampling rates, and number of measured acceleration values.
+The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
+The natural frequencies can be identified by the peaks in the magnitude graph.
+The first two natural frequencies are located at approximately 2.3\,Hz (1\textsuperscript{st} natural frequency) and 8\,Hz (2\textsuperscript{nd} natural frequency).
+Each sensor node transfers the values of the first natural frequency and the corresponding magnitude to the basestation. 
+The values are stored in the MySQL database and analyzed.
 
-\begin{figure}[ht]
+\begin{figure}[htb]
     \centering
-    \includegraphics{figures/50Hz.pdf}
+    \includegraphics[scale=0.9]{figures/50Hz.pdf}
     \caption{frequency domain graph of the oscillation}
@@
 
-After the
+The natural frequencies have been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values---and the results of a FE analysis.
+The highest possible sampling rate was determined at 76\,Hz.
+Higher samping rates could not be handled by the sensor nodes.
+Test runs with more than 512 measured values showed no significant increase in precision.
+Therefore, a sampling rate of 76\,Hz and 512 measured values were chosen.
 
-\begin{table}[ht]
+Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
+Example results are shown in \autoref{tab:lab-ex}.
+The identified magnitudes increases with ascending sensor node position from top to bottom, validating the SHM system.
+
+Through FFT analysis of sampling events, the data traffic is decreased from 512 measured acceleration values per sampling event to two values.
+A number stored in the used data format double occupies 8\,bytes of space.
+The data traffic is decreased by 510$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.
+
+\begin{table}[hb]
 	\centering
@@
 		\midrule
-		1 & top & 3.78 & 2.2461 & 50\\ 
-		1 & middle & 3.23 & 2.2461 & 50\\ 
-		1 & bottom & 1.97 & 2.2461 & 50\\ 
+		1 & top & 3.78 & 2.2461 & 76\\ 
+		1 & middle & 3.23 & 2.2461 & 76\\ 
+		1 & bottom & 1.97 & 2.2461 & 76\\ 
 		\midrule	
-		2 & top & 4.57 & 2.2461 & 50\\ 
-		2 & middle & 3.54 & 2.2461 & 50\\
-		2 & bottom & 2.43 & 2.2461 & 50\\ 
+		2 & top & 5.74 & 2.2536 & 76\\  
+		2 & middle & 4.38 & 2.2536 & 76\\ 
+		2 & bottom & 2.55 & 2.2536 & 76\\
 		\midrule
@@
 		3 & bottom & 3.79 & 2.2536 & 76\\
-		\midrule	
-		4 & top & 5.74 & 2.2536 & 76\\  
-		4 & middle & 4.38 & 2.2536 & 76\\ 
-		4 & bottom & 2.55 & 2.2536 & 76\\
-		\midrule		
-		5 & top & 5.53 & 2.2536 & 76\\  
-		5 & middle & 4.62 & 2.2536 & 76\\ 
-		5 & bottom & 2.74 & 2.2536 & 76\\ 		
 		\bottomrule
@@
 
-In structural health monitoring, a decentralized data compression stragegy reduces data traffic and storage space.
+In structural health monitoring, a decentralized data compression strategy reduces data traffic and storage space.
 Using a fast Fourier transform to determine the natural frequency allows conclusions about the structural integrity of a structure.
-The transmitted data was reduced exponentially, saving up to xx MB per sampling event.
+The transmitted data was reduced exponentially, saving up to 3,96\,MB per sampling event.
 

### 2015-02-10 Paper/PaperJahrSchlichSHM.tex
minor finfishing touches

@@
 To minimize the wireless data traffic, each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
-By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building.
+By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building \cite{rao2011fast}.
 
@@
     \centering
-    \includegraphics[scale=0.9]{figures/50Hz.pdf}
+    \includegraphics[scale=0.9]{figures/76HzHalf.pdf}
     \caption{frequency domain graph of the oscillation}
-    \label{fig:50hzfft}
+    \label{fig:76hzfft}
 \end{figure}

### 2015-02-10 Paper/PaperJahrSchlichSHM.tex

@@
 Using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
-In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
+In wireless SHM systems, the sensor nodes communicate with each other and with computer systems via wireless transceivers, eradicating wiring-specific problems.
 
@@
 
-In this paper, the natural frequency of a building structure is determined using a fast Fourier transform.
+In this paper, the natural frequency of a building structure is determined using a fast Fourier transform (FFT).
 Deviations from the natural frequency indicate changes in the vibration behavior, which may be caused by structural failures.
@@
 First, a wireless structural health monitoring system is designed and implemented. 
-Next, a decentralized data analysis approach, using fast Fourier transform (FFT), is developed.
+Next, a decentralized data analysis approach, using fast Fourier transform, is developed.
 Then, the system is tested in laboratory experiments.
@@
 The SHM system is programmed object-oriented in Java. 
-Object orientation uses objects which are instanciated using classes as paradigm. 
+Object orientation uses objects, which are instanciated using classes as paradigm. 
 A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
@@
 To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure.
-The test structure is a 4-story shear-frame consisting of four steel plates of 25\,cm\,$\times$\,50\,cm and a thickness of 0.8\,mm.
+The test structure is a 4-story shear-frame consisting of four steel plates of 25\,cm\,$\times$\,50\,cm\,$\times$\,0.8\,mm.
 The plates are mounted on threaded rods with a vertical clearance of 23\,cm.
-At the bottom, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm and a height of 30\,cm.
-The SHM system is installed on the test structure by fastening a wireless sensor node to the top three stories.
-The laboratory setup after system installation is shown in \autoref{fig:teststructure}.
+At the bottom, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm\,$\times$\,30\,cm.
+The SHM system is installed on the test structure by fastening one wireless sensor node to each of the top three stories.
+The laboratory setup is shown in \autoref{fig:teststructure}.
 
@@
 Each sensor node transfers the values of the first natural frequency and the corresponding magnitude to the basestation. 
-The values are stored in the MySQL database and analyzed.
+The values are stored in a MySQL database and analyzed.
 
@@
 Example results are shown in \autoref{tab:lab-ex}.
-The identified magnitudes increases with ascending sensor node position from top to bottom, validating the SHM system.
+The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.
 
-Through FFT analysis of sampling events, the data traffic is decreased from 512 measured acceleration values per sampling event to two values.
+Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
 A number stored in the used data format double occupies 8\,bytes of space.
-The data traffic is decreased by 510$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.
+The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.
 
@@
 		\toprule
-		ID & Position & Magnitude & Natural frequency & Sampling rate\\ 
+		ID & Position & Magnitude & Natural frequency [Hz] & Sampling rate [Hz]\\ 
 		\midrule

### 2015-02-12 Paper/PaperJahrSchlichSHM.tex
Final draft
Final Version
merge

@@
 To minimize the wireless data traffic, each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
-By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building \cite{rao2011fast}.
+By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building.
 
@@
     \centering
-    \includegraphics[scale=0.9]{figures/76HzHalf.pdf}
+    \includegraphics[scale=0.9]{figures/50Hz.pdf}
     \caption{frequency domain graph of the oscillation}
-    \label{fig:76hzfft}
+    \label{fig:50hzfft}
 \end{figure}

### 2015-02-12 Paper/PaperJahrSchlich.tex
WA Paper auf den neusten Stand gebracht.

@@
 \usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
-
+\usepackage{titlesec}
 \usepackage{color}
+\usepackage{booktabs}
 
@@
 
-% The following parameters seem to provide a reasonable page setup.
-\topmargin -1.0cm
-\oddsidemargin 0.2cm
-\textwidth 15cm 
-\textheight 23cm
-\footskip 1.0cm
+\makeatletter
+\renewcommand\section{\@startsection{section}{1}{\z@}%
+                     {-7.5ex \@plus -1ex \@minus -10.2ex}%
+                     {2.3ex \@plus.2ex}%
+                     {\sffamily\large\bfseries}}
+\makeatother
 
-%The next command sets up an environment for the abstract to your paper.
-\newenvironment{sciabstract}{%
-\begin{quote} \bfseries}
-{\end{quote}}
+\usepackage{geometry}
+\geometry{a4paper, top=35mm, left=30mm, right=30mm, bottom=35mm,
+headsep=10mm, footskip=12mm}
 
-% If your reference list includes text notes as well as references,
-% include the following line; otherwise, comment it out.
-
-\renewcommand\refname{References and Notes}
-
-% The following lines set up an environment for the last note in the
-% reference list, which commonly includes acknowledgments of funding,
-% help, etc.  It's intended for users of BibTeX or the {thebibliography}
-% environment.  Users who are hand-coding their references at the end
-% using a list environment such as {enumerate} can simply add another
-% item at the end, and it will be numbered automatically.
-
-\newcounter{lastnote}
-\newenvironment{scilastnote}{%
-\setcounter{lastnote}{\value{enumiv}}%
-\addtocounter{lastnote}{+1}%
-\begin{list}%
-{\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
-{\setlength{\labelsep}{.5em}}}
-{\end{list}}
-
-% Include your paper's title here
-
-\title{\textbf{ \Large{Decentralized autonomous sensor fault detection using neural networks}} }
-
+\title{\textbf{ \Large {Decentralized autonomous sensor fault detection using neural networks}}}
 \author
-{Katrin Jahr, Robert Schlich\\
+{\large Katrin Jahr, Robert Schlich\\
 \\
@@
 \\
-\normalsize{katrin.jahr@uni-weimar.de}\\
+\normalsize{katrin.jahr@uni-weimar.de \qquad}
 \normalsize{robert.schlich@uni-weimar.de}
 }
-
-% Include the date command, but leave its argument blank.
 \date{}
 
-%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%
+\clubpenalty = 10000 % schliesst Schusterjungen aus
+\widowpenalty = 10000 \displaywidowpenalty = 10000% schliesst Hurenkinder aus
+
+\newenvironment{sciabstract}{%
+\begin{quote} \itshape}
+{\end{quote}}
 
@@
 \setlength{\emergencystretch}{3pt}
-\hyphenation{sample-rate auto-nomous}
+\hyphenation{sample-rate auto-nomous moni-toring methods}
 
@@
 Traditionally, the sensor nodes are connected to computer systems with cables.
-However, using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
+Using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
 In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
@@
 The sensor nodes and the basestation are of type "Oracle Sun SPOT". 
-The Sun SPOTS are equipped with several components, see \autoref{fig:SPOT}.
+The Sun SPOTS are equipped with several components.
 Among others, the sensor board includes an accelerometer and eight independent RGB-LEDs.
-The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125Hz \citep{eDemo2010}.
-
-\todo{citation BILD!}
-\begin{figure}[ht]
-    \centering
-    \includegraphics[scale=0.5]{figures/eDemoboard.png}
-    \caption{Layout of the Sun SPOT sensor board}
-    \label{fig:SPOT}
-\end{figure}
+The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz \citep{eDemo2010}.
 
@@
 6. information retrieval.
-Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a fast Fourier transform to determine the characteristics of the oscillation of the structure.
+The dataflow is illustrated in \autoref{fig:flow}.
+Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a FFT to determine the characteristics of the oscillation of the structure.
 The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
@@
 
-\colorbox{cyan}{neuer Absatz:}
+\begin{figure}[htb]
+    \centering
+    \includegraphics{figures/dataflow_tasks.eps}
+    \caption{Dataflow and associated SHM tasks}
+    \label{fig:flow}
+\end{figure}
 
 The SHM system is programmed object-oriented in Java. 
-Object orientation uses objects, that are instanciated using classes as paradigm.
-A class includes methods, allowing the objects to perform actions, and attributes, specific data.
-\autoref{fig:UML} describes the classes of the SHM system code.
+Object orientation uses objects, which are instanciated using classes as paradigm. 
+A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
 The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
@@
 
-The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
-The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
-There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
-Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
+\autoref{fig:UML-sn} describes the classes of the \texttt{sensor\-node} package.
+The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, which are embedded directly into the sensor nodes.
+The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
+There are two phases: At first, the acceleration is measured with a low sampling rate.
+Once the acceleration exceeds a threshold, the second phase is entered by increasing the sampling rate. 
 The measured values are stored into an array.
 The different phases are indicatad by lighting different LEDs.
-The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
+The \texttt{FFT} class performs a FFT on the measured accelerations. 
 With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
-Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
-The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
-For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
-Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
+Finally, the natural frequency is determined by extracting the maximal magnitude.
+The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
+For starting the operation of the sensor node, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
+Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
 
-The package \texttt{basestation} runs on the host computer and operates the basestation.
-It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
-The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
-The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class.
-The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
-
-\begin{figure}[h!]
+\begin{figure}[htb]
     \centering
-    \includegraphics{figures/uml.pdf}
-    \caption{class diagram describing the structure of the implemented SHM system}
-    \label{fig:UML}
+    \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
+    \caption{Class diagram of the \texttt{sensornode} package}
+    \label{fig:UML-sn}
 \end{figure}
 
-\newpage
+\autoref{fig:UML-bs} describes the classes of the \texttt{base\-station} package.
+The package \texttt{base\-station} runs on the host computer and operates the basestation.
+It consists of the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
+The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
+The entry point of the program is the \texttt{run()} method in the \texttt{Main\-Base} class.
+The \texttt{Main\-Base} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
+    \caption{Class diagram of the \texttt{basestation} package}
+    \label{fig:UML-bs}
+\end{figure}
 
@@
 
-Neural networks are implemented into the sensor nodes using SNIPE\footnote{}, a open source Java Library.
+Neural networks are implemented into the sensor nodes using SNIPE, a open source Java Library.
 \begin{figure}[ht]
@@
 
-\newpage
 \section*{Laboratory experiments}
 
+In the following section, the laboratory experiments are described.
+First, a description of the test setup is given, second, the data aquisition and processing are depicted, and finally, the results are discussed.
+
 To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure.
-The test structure is a 4-story shear-frame consisting of four steel plates of 25\,cm\,$\times$\,50\,cm and a thickness of 0.8\,mm.
+The test structure is a 4-story shear-frame consisting of four steel plates of 25\,cm\,$\times$\,50\,cm\,$\times$\,0.8\,mm.
 The plates are mounted on threaded rods with a vertical clearance of 23\,cm.
-At the bottom, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm and a height of 30\,cm.
-The SHM system is installed on the test structure by fastening a wireless sensor node to each story.
-The laboratory setup after system installation is shown in \autoref{fig:teststructure}.
+At the bottom, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm\,$\times$\,30\,cm.
+The SHM system is installed on the test structure by fastening one wireless sensor node to each of the top three stories.
+The laboratory setup is shown in \autoref{fig:teststructure}.
 
-\todo{Bild ersetzen durch ein Bild mit den fertig installierten SPOTS}
-
-\begin{figure}[h!]
+\begin{figure}[ht]
     \centering
-    \includegraphics[scale=0.3]{figures/teststructure.jpg}
-    \caption{Laboratory setup}
+    \includegraphics[scale=0.4]{figures/testsetup.png}
+    \caption{Laboratory setup (Source: own photograph)}
     \label{fig:teststructure}
@@
 
+The structure is excited by deflecting and releasing the top of the structure.
 This excitation method ensures a free vibration in natural frequency with little interferences.
-After excitation, the sensor nodes automatically start the acceleration with a sampling rate of 40\,Hz.
-To minimize the wireless data communication, each node performs a fast Fourier transform algorithm once sufficient acceleration measurements have been collected.
-%Generally FFTs are used to convert equidistant samples of a function into a combination of frequencies, and the corresponding magnitudes, that has the same sample values. 
-The acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the building.
-Each sensor node transfers only the values of the predominating frequency---i.e.
-the frequency with the biggest magnitude--- \colorbox{cyan}{to the basestation}.
-The values are then stored in the MySQL database at the host computer.
+After excitation, when the acceleration threshold is exceeded, the sensor nodes automatically start measuring the acceleration.
+To minimize the wireless data traffic, each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
+By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building \cite{rao2011fast}.
 
-%\textbf{2. paragraph: data measurement and processing}
-%
-%\emph{\begin{itemize}
-%\item learning phase
-%\item excitation
-%\item sensors start measuring
-%\item sensor nodes \textbf{perform fft} - what is fft
-%\item sensor nodes exchange frequencies
-%\item neural networks check integrity
-%\item if no faults detected, sensor nodes send data to basestation
-%\end{itemize}}
-%
-%\textbf{3. paragraph: sensor fault detection}
-%
-%\begin{itemize}
-%\item simulation of sensor fault
-%\item neural networks check integrity -> error!
-%\item sensor node sends alert to basestation
-%\end{itemize}
+The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
+The natural frequencies can be identified by the peaks in the magnitude graph.
+The first two natural frequencies are located at approximately 2.3\,Hz (1\textsuperscript{st} natural frequency) and 8\,Hz (2\textsuperscript{nd} natural frequency).
+Each sensor node transfers the values of the first natural frequency and the corresponding magnitude to the basestation. 
+The values are stored in a MySQL database and analyzed.
+
+\begin{figure}[htb]
+    \centering
+    \includegraphics[scale=0.9]{figures/76HzHalf.pdf}
+    \caption{frequency domain graph of the oscillation}
+    \label{fig:76hzfft}
+\end{figure}
+
+The natural frequencies have been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values---and the results of a FE analysis.
+The highest possible sampling rate was determined at 76\,Hz.
+Higher samping rates could not be handled by the sensor nodes.
+Test runs with more than 512 measured values showed no significant increase in precision.
+Therefore, a sampling rate of 76\,Hz and 512 measured values were chosen.
+
+Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
+Example results are shown in \autoref{tab:lab-ex}.
+The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.
+
+Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
+A number stored in the used data format double occupies 8\,bytes of space.
+The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.
+
+\begin{table}[hb]
+	\centering
+	\begin{tabular}{l l c c c}
+		\toprule
+		ID & Position & Magnitude & Natural frequency [Hz] & Sampling rate [Hz]\\ 
+		\midrule
+		1 & top & 3.78 & 2.2461 & 76\\ 
+		1 & middle & 3.23 & 2.2461 & 76\\ 
+		1 & bottom & 1.97 & 2.2461 & 76\\ 
+		\midrule	
+		2 & top & 5.74 & 2.2536 & 76\\  
+		2 & middle & 4.38 & 2.2536 & 76\\ 
+		2 & bottom & 2.55 & 2.2536 & 76\\
+		\midrule
+		3 & top & 8.20 & 2.2536 & 76\\
+		3 & middle & 6.35 & 2.2536 & 76\\ 
+		3 & bottom & 3.79 & 2.2536 & 76\\
+		\bottomrule
+	\end{tabular}
+	\caption{Natural frequencies of the test structure}
+	\label{tab:lab-ex}
+\end{table}
 

### 2015-02-12 Paper/PaperJahrSchlich.tex
blaue Änderungsmarkierungen hinzugefügt

@@
     \centering
-    \includegraphics{figures/dataflow_tasks.eps}
+    \includegraphics{figures/dataflow_tasks_nn.eps}
     \caption{Dataflow and associated SHM tasks}
@@
 
+\colorbox{cyan}{neuer Absatz:}
 The SHM system is programmed object-oriented in Java. 
@@
 The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
-There are two phases: At first, the acceleration is measured with a low sampling rate.
-Once the acceleration exceeds a threshold, the second phase is entered by increasing the sampling rate. 
+There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
+Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
 The measured values are stored into an array.
@@
 With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
-Finally, the natural frequency is determined by extracting the maximal magnitude.
+Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
 The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
-For starting the operation of the sensor node, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
+For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
 Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
@@
 
-Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
+\textit{Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
 Example results are shown in \autoref{tab:lab-ex}.
-The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.
+The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.}
 
-Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
+\textit{Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
 A number stored in the used data format double occupies 8\,bytes of space.
-The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.
+The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.}
 

### 2015-02-15 SHM_JahrSchlich/basestation/src/DatabaseHandler.java
neuronale Netze implementiert, Kommunikation angepasst

@@ -46,7 +46,8 @@ public class DatabaseHandler {
 					+ "SpotID VARCHAR(19), "
 					+ "Magnitude DOUBLE, "
 					+ "Frequency FLOAT, "
-					+ "Samplerate INTEGER)");
+					+ "ExpectedMagnitude DOUBLE, "
+					+ "Error INT)");
 			statement.close();
 		}
 		resultSet.close();
@@ -54,11 +55,7 @@ public class DatabaseHandler {
 	
 	
 	public void writeMeasurement(String tableName, 
-								String spotID, 
-								double magnitude, 
-								float frequency, 
-								int samplerate) throws Exception {
-		
+								Measurement meas) throws Exception {
 		
 		//Get current systemtime as timestamp to avoid trouble with unsynchronized SunSPOTs
 		Calendar calendar = Calendar.getInstance();
@@ -67,12 +64,12 @@ public class DatabaseHandler {
 		System.out.println("Writing measurement into Database");
 		PreparedStatement preparedStatement = dbConn.prepareStatement("INSERT INTO "
 				+ tableName
-				+ " (Time, SpotID, Magnitude, Frequency, Samplerate) VALUES (?,?,?,?,?)");
+				+ " (Time, SpotID, Magnitude, Frequency, ExpectedMagnitude, Error) VALUES (?,?,?,?,?,?)");
 			preparedStatement.setTimestamp(1, now);
-			preparedStatement.setString(2, spotID);
-			preparedStatement.setDouble(3, magnitude);
-			preparedStatement.setFloat(4, frequency);
-			preparedStatement.setFloat(5, samplerate);
+			preparedStatement.setString(2, meas.address);
+			preparedStatement.setDouble(3, meas.magnitude);
+			preparedStatement.setFloat(4, meas.frequency);
+			preparedStatement.setFloat(5, meas.error);
 			preparedStatement.execute();
 			System.out.println(preparedStatement);
 		preparedStatement.close();

### 2015-02-15 SHM_JahrSchlich/basestation/src/MainBase.java

@@ -1,8 +1,12 @@
 package basestation;
 
 import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.io.j2me.radiostream.RadiostreamConnection;
 import com.sun.spot.peripheral.ota.OTACommandServer;
+import com.sun.spot.util.Utils;
 
+import java.io.DataInputStream;
+import java.io.DataOutputStream;
 import java.sql.DriverManager;
 import java.sql.Connection;
 import java.sql.Timestamp;
@@ -20,7 +24,7 @@ import java.sql.Statement;
 
 public class MainBase {
     // Broadcast port on which we listen for sensor samples
-    private static final int HOST_PORT = 67;
+    private static final int HOST_PORT = 68;
     private static final String DATABASE_URL = "jdbc:mysql://localhost:3306/";
     private static final String DATABASE_NAME = "SunSPOT";
     private static final String DATABASE_USER = "shm";
@@ -28,60 +32,82 @@ public class MainBase {
     private static final String TABLE_NAME = "measurements";
     private int samplerate = 76;  // in milliseconds
     
+    private String otherAddress = "0014.4F01.0000.763F";
+    
     String ourAddress = System.getProperty("IEEE_ADDRESS");
     
     private void run() throws Exception {
     	
     	//Establish a radio connection to the sensor nodes
-    	RadiogramConnection rCon;
-        Datagram dg;
-        DateFormat fmt = DateFormat.getTimeInstance();
-         
-        try {
-            // Open up a server-side broadcast radiogram connection
-            // to listen for sensor readings being sent by different SPOTs
-            rCon = (RadiogramConnection) Connector.open("radiogram://:" + HOST_PORT);
-            dg = rCon.newDatagram(rCon.getMaximumLength());
-        } catch (Exception e) {
-             System.err.println("setUp caught " + e.getMessage());
-             throw e;
-        }
-        
-    	//Create a DatabaseHandler, the DatabaseHandler will take care of 
-        // database interaction
-    	DatabaseHandler dbHandler = new DatabaseHandler();
-    	
-    	//Establish a connection with the MySQL-Database
-    	String url = DATABASE_URL + DATABASE_NAME;
-    	dbHandler.getConnection(url, DATABASE_USER, DATABASE_PASSWORD);
+        RadiostreamConnection conn;
+        DataInputStream dis;
+        DataOutputStream dos; 
 
-        //Create table with specified name
[…]

### 2015-02-15 SHM_JahrSchlich/basestation/src/Measurement.java

@@ -1,4 +1,5 @@
-package sensornode;
+package basestation;
+
 public class Measurement{
 	
 	public String address;

### 2015-02-15 SHM_JahrSchlich/sensornode/src/CommunicationBase.java

@@ -1,52 +1,0 @@
-package sensornode;
-
-import com.sun.spot.io.j2me.radiogram.*;
-import com.sun.spot.resources.Resources;
-import com.sun.spot.resources.transducers.ITriColorLED;
-import com.sun.spot.util.Utils;
-
-import javax.microedition.io.*;
-
-public class CommunicationBase {
-		
-		// open radiogramm connection
-        RadiogramConnection rCon = null;
-        Datagram dg = null;
-        String ourAddress = System.getProperty("IEEE_ADDRESS");
-        ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED1");
-        
-        public void EstablishConnection(int hostPort){
-            try {
-                // Open up a broadcast connection to the host port
-                // where the 'on Desktop' portion of this demo is listening
-                rCon = (RadiogramConnection) Connector.open("radiogram://broadcast" + hostPort);
-                dg = rCon.newDatagram(rCon.getMaximumLength());
-                System.out.println("Starting communication on " + ourAddress);
-            } catch (Exception e) {
-                System.err.println("Caught " + e + " in connection initialization.");
-            }	
-        }
-        
-        public void SendMeasurement(double magnitude, double frequencyMax) {
-        	try {
-	        	// Pack time and sensor reading into a radio datagram, send it.
-		        dg.reset();
-		        dg.writeDouble(magnitude);
-		        // cast frequency to double
-		        float frequency = (float) frequencyMax;
-		        dg.writeFloat(frequency);
-		        rCon.send(dg);
-		        System.out.println("SENT!");
-		        
-		        // blink first LED
-		        led.setRGB(255, 255, 0);
-	            led.setOn();
-				Utils.sleep(10);
-	            led.setOff();
-        	 } catch (Exception e) {   
-        		 System.err.println("Caught " + e + " in sending data.");
-        	 }
-        }
-        
-}
-        

### 2015-02-15 SHM_JahrSchlich/sensornode/src/CommunicationNormal.java

@@ -21,28 +21,8 @@ public class CommunicationNormal {
         DataInputStream dis;
         DataOutputStream dos; 
         
-        public void EstablishStreamConnection(int hostPort, String otherAddress){
-            try {
-                // Open up a broadcast connection to the host port
-                // where the 'on Desktop' portion of this demo is listening
-                conn = (RadiostreamConnection)Connector.open("radiostream://" + otherAddress + ":" + hostPort);
-                
-                DataInputStream dis = conn.openDataInputStream();
-                DataOutputStream dos = conn.openDataOutputStream();
-                
-                System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
-                
-                dos.writeUTF("HALLO");
-                
-                
-            } catch (Exception e) {
-                System.err.println("Caught " + e + " in connection initialization.");
-            }	
-        }
+        public void ExchangeData (Measurement ourMeas, String otherAddress, int hostPort) {
 
-        public Measurement ExchangeData (Measurement ourMeas, String otherAddress, int hostPort) {
-            //object for other measurement
-            Measurement othMeas = new Measurement();
         	try {     
         		
         		System.out.println("starting ExchangeData. ");
@@ -63,6 +43,7 @@ public class CommunicationNormal {
         			try{
 	            		dos.writeFloat(ourMeas.frequency);
 	    				dos.writeDouble(ourMeas.magnitude);
+	    				dos.writeInt(ourMeas.error);
 	    				dos.flush();
 	    				System.out.println("datastream flushed." );
 	    				
@@ -96,11 +77,7 @@ public class CommunicationNormal {
         		
         	} catch (Exception e) {   
         		 System.err.println("Caught " + e + " in exchanging data.");
-        	} finally {
-
-	    		return othMeas;
-        	}
-
         	
+        	}
         }
 }

### 2015-02-15 SHM_JahrSchlich/sensornode/src/FFT.java

@@ -193,7 +193,7 @@ public class FFT {
 		//System.out.println("maxMagnitude= "+ maxMagnitude+
 		//									", maxFrequency= " + maxFrequency);
 		
-		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency);
+		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency, 0);
         return measurement;
     }
 }

### 2015-02-15 SHM_JahrSchlich/sensornode/src/MainSpot.java

@@ -35,13 +35,9 @@ public class MainSpot extends MIDlet {
     	// initalize communication between spots
 		CommunicationNormal communication = new CommunicationNormal();
 		
-		//Radiogram
-		//communication.EstablishConnection(HOST_PORT);
-		
         //initialize AccelerationSampler
         AccelerationSampler sensorSampler = new AccelerationSampler();
 		
-        
        
         while (true){
 	        // get acceleration for this sensornode
@@ -59,38 +55,8 @@ public class MainSpot extends MIDlet {
 			System.out.println("I did my FFT!");
 			
 			// exchange measurement with measurement of other sensors
-			Measurement othMeas = communication.ExchangeData(ownMeas, otherAddress, HOST_PORT);
+			communication.ExchangeData(ownMeas, otherAddress, HOST_PORT);
 			
-			
-			
-			/*
-			
-			// get FFT from other sensors			
-			double[] allmagnitudes = new double[NUMBER_OF_OTHER_SENSORS];
-			for (int i = 0; i < NUMBER_OF_OTHER_SENSORS; i++){
-				Measurement measurement = communication.RetrieveSensorData();
-				// store recieved data in array
-				allmagnitudes[i] = measurement.magnitude;
-				System.out.println("I got " + i + " other FFTs!");
-			}
-			
-			*/
-			
-			
-			
-			/*
-			
-			
-			TRAIN NETWORK
-			
-			
-			*/
-
-			
-			
-			
-		
-			//send processed data to basestation
 
 			Utils.sleep(2000);
 		}   

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/CommunicationBase.java

@@ -1,52 +1,0 @@
-package sensornodeNN;
-
-import com.sun.spot.io.j2me.radiogram.*;
-import com.sun.spot.resources.Resources;
-import com.sun.spot.resources.transducers.ITriColorLED;
-import com.sun.spot.util.Utils;
-
-import javax.microedition.io.*;
-
-public class CommunicationBase {
-		
-		// open radiogramm connection
-        RadiogramConnection rCon = null;
-        Datagram dg = null;
-        String ourAddress = System.getProperty("IEEE_ADDRESS");
-        ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED1");
-        
-        public void EstablishConnection(int hostPort){
-            try {
-                // Open up a broadcast connection to the host port
-                // where the 'on Desktop' portion of this demo is listening
-                rCon = (RadiogramConnection) Connector.open("radiogram://broadcast" + hostPort);
-                dg = rCon.newDatagram(rCon.getMaximumLength());
-                System.out.println("Starting communication on " + ourAddress);
-            } catch (Exception e) {
-                System.err.println("Caught " + e + " in connection initialization.");
-            }	
-        }
-        
-        public void SendMeasurement(double magnitude, double frequencyMax) {
-        	try {
-	        	// Pack time and sensor reading into a radio datagram, send it.
-		        dg.reset();
-		        dg.writeDouble(magnitude);
-		        // cast frequency to double
-		        float frequency = (float) frequencyMax;
-		        dg.writeFloat(frequency);
-		        rCon.send(dg);
-		        System.out.println("SENT!");
-		        
-		        // blink first LED
-		        led.setRGB(255, 255, 0);
-	            led.setOn();
-				Utils.sleep(10);
-	            led.setOff();
-        	 } catch (Exception e) {   
-        		 System.err.println("Caught " + e + " in sending data.");
-        	 }
-        }
-        
-}
-        

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/CommunicationNN.java

@@ -2,6 +2,7 @@ package sensornodeNN;
 
 import com.sun.spot.io.j2me.radiogram.*;
 import com.sun.spot.io.j2me.radiostream.*;
+
 import java.io.DataInputStream;
 import java.io.DataOutputStream;
 
@@ -16,31 +17,13 @@ public class CommunicationNN {
 	    String ourAddress = System.getProperty("IEEE_ADDRESS");
         ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED2");
         ITriColorLED led3 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED3");
+        ITriColorLED led5 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED5");
         
         RadiostreamConnection conn;
         DataInputStream dis;
         DataOutputStream dos; 
         
-        public void EstablishStreamConnection(int hostPort, String otherAddress){
-            try {
-                // Open up a broadcast connection to the host port
-                // where the 'on Desktop' portion of this demo is listening
-                conn = (RadiostreamConnection)Connector.open("radiostream://" + otherAddress + ":" + hostPort);
-                
-                DataInputStream dis = conn.openDataInputStream();
-                DataOutputStream dos = conn.openDataOutputStream();
-                
-                System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
-                
-                dos.writeUTF("HALLO");
-                
-                
-            } catch (Exception e) {
-                System.err.println("Caught " + e + " in connection initialization.");
-            }	
-        }
-
-        public Measurement ReceiveData(Measurement ourMeas, String otherAddress, int hostPort) {
+        public Measurement ReceiveData(String otherAddress, int hostPort) {
             //object for other measurement
             Measurement othMeas = new Measurement();
         	try {     
@@ -54,21 +37,20 @@ public class CommunicationNN {
                  
         		System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
 
-				System.out.println("try to recieve." );	            
-			
+				System.out.println("try to recieve." );	
+				// receive data from stream
 	            othMeas.address = otherAddress;
 	    		othMeas.magnitude = dis.readDouble();
 	    		othMeas.frequency = dis.readFloat();
+	    		othMeas.error = dis.readInt();
 	    		
 				System.out.println("datastream received." );
 	    		
 		        // blink third LED
[…]

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/ErrorMeasurement.java

@@ -1,0 +1,200 @@
+package sensornodeNN;
+
+
+
+
+/**
+ * 
+ * This class contains several methods of error measurement for use in a static
+ * way.
+ * 
+ * @author David Kriesel / dkriesel.com
+ * 
+ */
+public class ErrorMeasurement {
+
+	/**
+	 * Calculates the sum of absolute errors of the given network over all
+	 * samples in the training lesson.
+	 * 
+	 * @param net
+	 * @param lesson
+	 * 
+	 * @return calculated error
+	 */
+	public static double getErrorAbsoluteSum(NeuralNetwork net,
+			TrainingSampleLesson lesson) {
+
+		double[][] inputs = lesson.getInputs();
+		double[][] desiredOutputs = lesson.getDesiredOutputs();
+
+		double error = 0.0;
+
+		for (int i = 0; i < inputs.length; i++) {
+
+			double[] outputs = net.propagate(inputs[i]);
+
+			for (int j = 0; j < outputs.length; j++) {
+				double temp = 0.0;
+				temp = desiredOutputs[i][j] - outputs[j];
+				temp = Math.abs(temp);
+				error += temp;
+			}
+		}
+		return error;
+
+	}
+
+	/**
+	 * Calculates the sum of euclidean errors of the given network over all
+	 * samples in the training lesson.
+	 * 
+	 * @param net
+	 * @param lesson
+	 * 
+	 * @return calculated error
+	 */
+	public static double getErrorEuclideanSum(NeuralNetwork net,
+			TrainingSampleLesson lesson) {
+
+		double[][] inputs = lesson.getInputs();
+		double[][] desiredOutputs = lesson.getDesiredOutputs();
+
+		double error = 0.0;
+
+		for (int i = 0; i < inputs.length; i++) {
+
+			double[] outputs = net.propagate(inputs[i]);
+
+			for (int j = 0; j < outputs.length; j++) {
+				double temp = 0.0;
+				temp = desiredOutputs[i][j] - outputs[j];
+				temp = temp * temp;
+				temp = Math.sqrt(temp);
+				error += temp;
+			}
+		}
+		return error;
+	}
+
+	/**
+	 * Calculates the Squared percentage error of the given network over all
+	 * samples in the training lesson. This method is proposed by Prechelt
+	 * (Prechelt 1994, Proben1 - A set of neural Network Benchmarking Problems
+	 * and benchmarking rules) and seeks for an error measurement that is
[…]

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/FFT.java

@@ -193,7 +193,7 @@ public class FFT {
 		//System.out.println("maxMagnitude= "+ maxMagnitude+
 		//									", maxFrequency= " + maxFrequency);
 		
-		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency);
+		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency,0);
         return measurement;
     }
 }

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/Identity.java

@@ -1,0 +1,40 @@
+package sensornodeNN;
+
+
+/**
+ * Implements the Identity Function and its derivative. Note: The
+ * getAbsoluteMaximumLocationOfSecondDerivative() function just returns 1 in
+ * order to not cause malfunctions for example in synapse initialization.
+ * 
+ * @author David Kriesel / dkriesel.com
+ * 
+ */
+public class Identity implements NeuronBehavior{
+
+//	@Override
+	public double computeDerivative(double x) {
+		return 1;
+	}
+
+//	@Override
+	public double computeActivation(double x) {
+		return x;
+	}
+
+//	@Override
+	public NeuronBehavior getDedicatedInstance() {
+		return new Identity();
+	}
+
+//	@Override
+	public boolean needsDedicatedInstancePerNeuron() {
+		return false;
+	}
+
+
+//	@Override
+	public double getAbsoluteMaximumLocationOfSecondDerivative() {
+		return 1;
+	}
+
+}

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/MainSpot.java

@@ -11,24 +11,33 @@ import javax.microedition.midlet.MIDletStateChangeException;
 
 import com.sun.spot.resources.transducers.IAccelerometer3D;
 import com.sun.spot.service.BootloaderListenerService;
-//import java.io.Writer;
+
 
 
 public class MainSpot extends MIDlet {
-
-    private int HOST_PORT = 67;
+	
+	// measurement
     private int SAMPLE_PERIOD_LISTENING = 1 * 500;  // in milliseconds
     private int SAMPLE_PERIOD_MEASURING = 13;  // in milliseconds
     private int ARRAY_LENGTH = 128;
     private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
     private double THRESHOLD = 0.2;
 
+    // communication
+    private String BASE_NAME = "0014.4F01.0000.74AD";
     private String[] SENSOR_NAMES = {"0014.4F01.0000.7840",
     								"0014.4F01.0000.792D"};
 //    								"0014.4F01.0000.7AFA"}
     private String ourAddress = System.getProperty("IEEE_ADDRESS");
+    private int HOST_PORT = 67;
+    private int HOST_PORT_BASE = 68;
     
     
+    // neural network
+    private int HIDDEN_UNITS = 3;
+    
+    private int trainingEvents = 3;
+
     protected void startApp() throws MIDletStateChangeException {
 
         // Listen for downloads/commands over USB connection
@@ -36,15 +45,35 @@ public class MainSpot extends MIDlet {
     	// initalize communication between spots
 		CommunicationNN communication = new CommunicationNN();
 		
-		//Radiogram
+		//Radiogram for communication to basestation
 		//communication.EstablishConnection(HOST_PORT);
 		
         //initialize AccelerationSampler
         AccelerationSampler sensorSampler = new AccelerationSampler();
 		
-        
-       
-        while (true){
+    	ITriColorLED led4 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED4");
+    	
+		int numberOfSensors = SENSOR_NAMES.length;	
+		
+		// NEURAL NETWORK CREATION
+		// descriptor
+		NeuralNetworkDescriptor descriptorNN = new NeuralNetworkDescriptor( numberOfSensors, HIDDEN_UNITS,  1);
+    	descriptorNN.setSettingsTopologyFeedForward();
+
+    	// create neural network
[…]

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/Measurement.java

@@ -2,17 +2,18 @@ package sensornodeNN;
 
 public class Measurement{
 	
-	
 		public String address;
 		public float frequency;
 		public double magnitude;
+		public int error;
 		
 		// initiates a measurement
-		Measurement(String a, double m, float f){
+		Measurement(String a, double m, float f, int e){
 			
 			address = a;
 			magnitude = m;
 			frequency = f;
+			error = e;
 		}
 		
 		// initiates a measurement
@@ -21,8 +22,6 @@ public class Measurement{
 			address = "a";
 			magnitude = 0;
 			frequency = 0;
+			error = 0;
 		}		
-		
-		 
-        
-    }
+}

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/NetworkTraining.java

@@ -1,0 +1,31 @@
+package sensornodeNN;
+
+public class NetworkTraining {
+	  
+
+	public static void Train(NeuralNetwork net, TrainingSampleLesson lesson) {
+
+		 
+		int runs = 20000; // number of random patterns to train
+		double eta = 0.2; //learning rate
+		double threshold = 0.99;
+	 
+		
+		// Train neural network until error falls below limit
+		double error;
+		double newerror;
+		int numOfPhases = 0;
+	
+		do{
+			numOfPhases++;
+			error = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
+			net.trainBackpropagationOfError(lesson, runs, eta);
+			newerror = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
+			System.out.println("Error after training phase "+ numOfPhases + " = " + newerror );
+		}while(newerror / error > threshold);
+		
+		System.out.println(numOfPhases + "training phases. ");
+
+		}
+
+	}

### 2015-02-15 SHM_JahrSchlich/sensornodeNN/src/TrainingSampleLesson.java

@@ -1,4 +1,4 @@
-package spot.src.org.sunspotworld.demo.neuralnetwork;
+package sensornodeNN;
 
 
 //import java.io.Serializable;

### 2015-02-15 Paper/PaperJahrSchlich.tex
nn in shm implementation

@@
 %----------------------------------------------------------------------------------------
-
-\linenumbers % Schaltet Zeilennummerierung ein
-\modulolinenumbers[5] % nur jede 5. Zeile
-\section*{Dictionary}
-
-\begin{tabular}{|l|l|}
-\hline 
-Sensorknoten (SunSPOT) & sensor node \\ 
-\hline 
-einzelner Messsensor (Thermometer) & sensor \\ 
-\hline 
-Knoten im neuronalen Netz & neuron \\ 
-\hline 
-eine abgeschlossene Messungreihe & test run \\ 
-\hline 
-gemessene Werte & sensor data \\ 
-\hline 
-vorhergesagte Werte & predicted data \\ 
-\hline 
-durch Vorhersage erwartete Werte & expected data \\ 
-\hline 
-einzelner Messwert & measurement \\ 
-\hline 
-Test & laboratory experiments\\ 
-\hline
-tatsächliche, nicht virtuelle Messung & actual measurement \\
-\hline
-Messaufbau & test setup \\
-\hline
-
-\end{tabular}
+%
+%\linenumbers % Schaltet Zeilennummerierung ein
+%\modulolinenumbers[5] % nur jede 5. Zeile
+%\section*{Dictionary}
+%
+%\begin{tabular}{|l|l|}
+%\hline 
+%Sensorknoten (SunSPOT) & sensor node \\ 
+%\hline 
+%einzelner Messsensor (Thermometer) & sensor \\ 
+%\hline 
+%Knoten im neuronalen Netz & neuron \\ 
+%\hline 
+%eine abgeschlossene Messungreihe & test run \\ 
+%\hline 
+%gemessene Werte & sensor data \\ 
+%\hline 
+%vorhergesagte Werte & predicted data \\ 
+%\hline 
+%durch Vorhersage erwartete Werte & expected data \\ 
+%\hline 
+%einzelner Messwert & measurement \\ 
+%\hline 
+%Test & laboratory experiments\\ 
+%\hline
+%tatsächliche, nicht virtuelle Messung & actual measurement \\
+%\hline
+%Messaufbau & test setup \\
+%\hline
+%
+%\end{tabular}
 
@@
 
-\colorbox{cyan}{neuer Absatz:}
+\colorbox{cyan}{section ab hier teilweise neu (NN):}
 The SHM system is programmed object-oriented in Java. 
@@
 \autoref{fig:UML-sn} describes the classes of the \texttt{sensor\-node} package.
-The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, which are embedded directly into the sensor nodes.
+The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Measure\-ment}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the Java library "SNIPE" \todo{Quelle für SNIPE!} that implement a framework for neural networks.
+The classes are embedded directly into the sensor nodes.
+
 The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
@@
 Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
-The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
+
+The \texttt{Measure\-ment} class defines analyzed measurements by storing the address of the sensor node, the maximal magnitude, the natural frequency and if a sensor defect was detected.
+For every measurement, one object of the type \texttt{Measure\-ment} is instantiated.
+Using the SNIPE library, a neural network is defined and created for fault detection.
+The \texttt{Network\-Training} class defines a routine to train the neural network with prior measured \texttt{Measure\-ment} objects.
+The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer the \texttt{Measure\-ment} objects from the sensor node to the basestation.
 For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
-Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
+Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE \texttt{Neural\-Network} class, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the measurement.
 

### 2015-02-16 Paper/PaperJahrSchlich.tex
Changed drawing of neural network, minor changes to the paper
new dataflow figure, notes to labex

@@
 
-The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
+\textit{The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
 The natural frequencies can be identified by the peaks in the magnitude graph.
@@
 Each sensor node transfers the values of the first natural frequency and the corresponding magnitude to the basestation. 
-The values are stored in a MySQL database and analyzed.
+The values are stored in a MySQL database and analyzed.}
 
@@
 
-The natural frequencies have been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values---and the results of a FE analysis.
+\textit{The natural frequencies have been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values---and the results of a FE analysis.
 The highest possible sampling rate was determined at 76\,Hz.
@@
 Test runs with more than 512 measured values showed no significant increase in precision.
-Therefore, a sampling rate of 76\,Hz and 512 measured values were chosen.
+Therefore, a sampling rate of 76\,Hz and 512 measured values were chosen.}
 
-\textit{Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
-Example results are shown in \autoref{tab:lab-ex}.
-The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.}
+\begin{itemize}
+\item neuronales Netz
+\item TRAININGSPHASE
+\item Austausch der gemessenen FFT zwischen den Sensoren
+\item trainieren des neuronalen Netzes, bis keine Verbesserung bei Vorhersagegenauigkeit eintritt
+\item MESSPHASE
+\item Austausch der gemessenen FFT zwischen den Sensoren
+\item Nutzen des neuronalen Netzes, um eventuellen Fehler zu finden
+\item KÜNSTLICHEN FEHLER ERZEUGEN
+\item TABELLE MIT ERMITTELTEN / GEMESSENEN WERTEN; MIT / OHNE FEHLER
+\end{itemize}
 
-\textit{Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
-A number stored in the used data format double occupies 8\,bytes of space.
-The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.}
+%\textit{Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
+%Example results are shown in \autoref{tab:lab-ex}.
+%The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.}
+%
+%\textit{Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
+%A number stored in the used data format double occupies 8\,bytes of space.
+%The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.}
 

### 2015-02-16 Paper/PaperJahrSchlich.tex
new dataflow figure, notes to labex
letzte Zwischenabgabe

@@
 \usepackage[numbers, sort&compress, square]{natbib}
-\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
+
 \usepackage{titlesec}
@@
 
-\makeatletter
-\renewcommand\section{\@startsection{section}{1}{\z@}%
-                     {-7.5ex \@plus -1ex \@minus -10.2ex}%
-                     {2.3ex \@plus.2ex}%
-                     {\sffamily\large\bfseries}}
-\makeatother
+\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
+
+%\makeatletter
+%\renewcommand\section{\@startsection{section}{1}{\z@}%
+%                     {-7.5ex \@plus -1ex \@minus -10.2ex}%
+%                     {2.3ex \@plus.2ex}%
+%                     {\sffamily\large\bfseries}}
+%\makeatother
 
@@
 In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
+\colorbox{cyan}{hinzugefügt: }However, when using wireless SHM systems, special attention has to be paid to data transfer.
+To save transmission time and storage space, and to decrease the probability of data loss, the transmitted data is reduced using embedded computing.
 
@@
 The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
-Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.
+Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.
 
 This paper is organized as follows:
-First, a wireless structural health monitoring system is designed and implemented. 
+First, a dezentralized wireless structural health monitoring system is designed and implemented. 
 Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
@@
 %----------------------------------------------------------------------------------------
-
-\newpage
 
@@
 5. diagnostics and 
-6. information retrieval.
+6. information retrieval, see \citet{BisbySHM}.
 The dataflow is illustrated in \autoref{fig:flow}.
 Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a FFT to determine the characteristics of the oscillation of the structure.
+The results of the FFT are fed into a neural network to detect sensor faults.
 The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
@@
 
-\begin{figure}[htb]
+\begin{figure}[hbtp]
     \centering
@@
 
-\colorbox{cyan}{section ab hier teilweise neu (NN):}
+The neural network represents the structure of the SHM system.
+Each sensor node of the SHM system is represented by either an input neuron or an output neuron, as shown in \autoref{fig:neuralnetwork}.
+During computation, each sensor node embeds a distinct neural network. 
+Measurements of other sensor nodes are fed into the input neurons of the neural network; the output neuron returns the predicted measurement of the embedding sensor node. 
+Hidden neurons are used for computation and do not represent a physical counterpart.
+
+\begin{figure}[hbtp]
+    \centering
+    \includegraphics{figures/neuralnetwork.pdf}
+    \caption{Scheme of the neural network}
+    \label{fig:neuralnetwork}
+\end{figure}
+
 The SHM system is programmed object-oriented in Java. 
@@
 \autoref{fig:UML-sn} describes the classes of the \texttt{sensor\-node} package.
-The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Measure\-ment}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the Java library "SNIPE" \todo{Quelle für SNIPE!} that implement a framework for neural networks.
+The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Measure\-ment}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the Java library "SNIPE" \citep{SNIPE} that implement a framework for neural networks.
 The classes are embedded directly into the sensor nodes.
@@
 Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
-
-The \texttt{Measure\-ment} class defines analyzed measurements by storing the address of the sensor node, the maximal magnitude, the natural frequency and if a sensor defect was detected.
+The \texttt{Measure\-ment} class defines analyzed measurements by storing the address of the sensor node, the maximal magnitude, the natural frequency and if a sensor fault was detected during runtime.
 For every measurement, one object of the type \texttt{Measure\-ment} is instantiated.
@@
 For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
-Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE \texttt{Neural\-Network} class, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the measurement.
+Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the measurement.
 
-\begin{figure}[htb]
+\begin{figure}[htbp]
     \centering
     \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
-    \caption{Class diagram of the \texttt{sensornode} package}
+    \caption{Class diagram of the \texttt{sensornode} package ((Update auf NN folgt))}
     \label{fig:UML-sn}
@@
 
-\begin{figure}[htb]
+\begin{figure}[htbp]
     \centering
@@
 
-%----------------------------------------------------------------------------------------
-
-\section*{Implementation and training of neural networks}
-
-Neural networks are implemented into the sensor nodes using SNIPE, a open source Java Library.
-\begin{figure}[ht]
-    \centering
-    \includegraphics{figures/neuralnetwork.pdf}
-    \caption{Schematic drawing of an artificial neural network with three layers}
-    \label{fig:neuralnetwork}
-\end{figure}
-\todo[inline]{Grafik wird an unser neuronales Netz angepasst}
-
-%\begin{itemize}
-%\item neurons
-%\item layer
-%\item weights
-%\item activation function / identity function
-%\end{itemize}
-
-%\textbf{2. paragraph: proposed nn}
-%
-%\begin{itemize}
-%\item SNIPE? implementation in java?
-%\item activation function / identity function
-%\end{itemize}
-%
-%\textbf{3. paragraph: learning general}
-%
-%\begin{itemize}
-%\item set with test values
-%\item iterative weight adjustment
-%\end{itemize}
-%
-%\textbf{4. paragraph?: learning, proposed}
-
-%----------------------------------------------------------------------------------------
-
 \section*{Laboratory experiments}
@@
 In the following section, the laboratory experiments are described.
-First, a description of the test setup is given, second, the data aquisition and processing are depicted, and finally, the results are discussed.
+First, a description of the test setup is given, second, the data aquisition, data processing and sensor fault detection are depicted, and, finally, the results are discussed.
 
@@
 After excitation, when the acceleration threshold is exceeded, the sensor nodes automatically start measuring the acceleration.
-To minimize the wireless data traffic, each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
+Each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
 By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building \cite{rao2011fast}.
 
-\textit{The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
-The natural frequencies can be identified by the peaks in the magnitude graph.
-The first two natural frequencies are located at approximately 2.3\,Hz (1\textsuperscript{st} natural frequency) and 8\,Hz (2\textsuperscript{nd} natural frequency).
-Each sensor node transfers the values of the first natural frequency and the corresponding magnitude to the basestation. 
-The values are stored in a MySQL database and analyzed.}
-
-\begin{figure}[htb]
-    \centering
-    \includegraphics[scale=0.9]{figures/76HzHalf.pdf}
-    \caption{frequency domain graph of the oscillation}
-    \label{fig:76hzfft}
-\end{figure}
-
-\textit{The natural frequencies have been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values---and the results of a FE analysis.
-The highest possible sampling rate was determined at 76\,Hz.
-Higher samping rates could not be handled by the sensor nodes.
-Test runs with more than 512 measured values showed no significant increase in precision.
-Therefore, a sampling rate of 76\,Hz and 512 measured values were chosen.}
-
 \begin{itemize}
-\item neuronales Netz
+\item \textbf{neuronales Netz}
 \item TRAININGSPHASE
@@
 \item KÜNSTLICHEN FEHLER ERZEUGEN
-\item TABELLE MIT ERMITTELTEN / GEMESSENEN WERTEN; MIT / OHNE FEHLER
+\item TABELLE MIT ERMITTELTEN / GEMESSENEN WERTEN; MIT / OHNE FEHLERN
 \end{itemize}
 
-%\textit{Every sensor calculated the same natural frequency to a precision of four decimal points regardless of sensor position and degree of excitation.
+%The neural network has been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values.
+
+%The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
+%The natural frequencies can be identified by the peaks in the magnitude graph.
+%Each sensor node stores the calculated first natural frequency and the corresponding magnitude in a \texttt{Measurement} object and transmits the object to the other sensor nodes.
+
+%\begin{figure}[htb]
+%    \centering
+%    \includegraphics{figures/76HzHalf.pdf}
+%    \caption{frequency domain graph of the oscillation}
+%    \label{fig:76hzfft}
+%\end{figure}
+
+After the sensor fault detection, the \texttt{Measurement} objects of all sensor nodes ar transmitted to the basestation.
+The values and detected sensor faults are stored in a MySQL database and availeble for further analysis.
+
 %Example results are shown in \autoref{tab:lab-ex}.
-%The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.}
+%The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.
 %
@@
 
-\begin{table}[hb]
-	\centering
-	\begin{tabular}{l l c c c}
-		\toprule
-		ID & Position & Magnitude & Natural frequency [Hz] & Sampling rate [Hz]\\ 
-		\midrule
-		1 & top & 3.78 & 2.2461 & 76\\ 
-		1 & middle & 3.23 & 2.2461 & 76\\ 
-		1 & bottom & 1.97 & 2.2461 & 76\\ 
-		\midrule	
-		2 & top & 5.74 & 2.2536 & 76\\  
-		2 & middle & 4.38 & 2.2536 & 76\\ 
-		2 & bottom & 2.55 & 2.2536 & 76\\
-		\midrule
-		3 & top & 8.20 & 2.2536 & 76\\
-		3 & middle & 6.35 & 2.2536 & 76\\ 
-		3 & bottom & 3.79 & 2.2536 & 76\\
-		\bottomrule
-	\end{tabular}
-	\caption{Natural frequencies of the test structure}
-	\label{tab:lab-ex}
-\end{table}
+%\begin{table}[hb]
+%	\centering
+%	\begin{tabular}{l l c c c}
+%		\toprule
+%		ID & Position & Magnitude & Natural frequency [Hz] & Sampling rate [Hz]\\ 
+%		\midrule
+%		1 & top & 3.78 & 2.2461 & 76\\ 
+%		1 & middle & 3.23 & 2.2461 & 76\\ 
+%		1 & bottom & 1.97 & 2.2461 & 76\\ 
+%		\midrule	
+%		2 & top & 5.74 & 2.2536 & 76\\  
+%		2 & middle & 4.38 & 2.2536 & 76\\ 
+%		2 & bottom & 2.55 & 2.2536 & 76\\
+%		\midrule
+%		3 & top & 8.20 & 2.2536 & 76\\
+%		3 & middle & 6.35 & 2.2536 & 76\\ 
+%		3 & bottom & 3.79 & 2.2536 & 76\\
+%		\bottomrule
+%	\end{tabular}
+%	\caption{Natural frequencies of the test structure}
+%	\label{tab:lab-ex}
+%\end{table}
 

### 2015-02-16 SHM_JahrSchlich/sensornode/src/CommunicationNormal.java
Basestationübertragung läuft bei Robert

@@ -36,9 +36,6 @@ public class CommunicationNormal {
         		System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
 
         		boolean okay = false;
-        		
-        		        		
-        		
         		while(okay == false){
         			try{
 	            		dos.writeFloat(ourMeas.frequency);

### 2015-02-16 SHM_JahrSchlich/sensornodeNN/src/CommunicationNN.java

@@ -18,6 +18,7 @@ public class CommunicationNN {
         ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED2");
         ITriColorLED led3 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED3");
         ITriColorLED led5 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED5");
+        ITriColorLED led6 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED6");
         
         RadiostreamConnection conn;
         DataInputStream dis;
@@ -64,8 +65,7 @@ public class CommunicationNN {
         		dos.close();
         		conn.close();
         		
-        		// wait half a second, turn off the light
-				Utils.sleep(500);
+        		// turn off the light
         		led2.setOff();
 				led3.setOff();
 	    		        		
@@ -81,11 +81,19 @@ public class CommunicationNN {
         public void StoreData(Measurement[] allMeas, int hostPort, String baseAddress) {
         	//object for other measurement
         	
-        	for (int meas = 0; meas < allMeas.length; meas++){
+        	System.out.println("allMeas.length= "+ allMeas.length);
+        	
+        	int meas = 0;
+        	
+        	while ( meas < allMeas.length){
         		
-	        	try {     
+	        	try {   
+	            	System.out.println("meas= "+ meas);
+	            		        		
+	        		
 	        		conn = (RadiostreamConnection)Connector.open("radiostream://" + baseAddress + ":" + hostPort);
-
+	        		conn.setTimeout(300); 
+	        		
 	        		DataInputStream dis = conn.openDataInputStream();
 	        		DataOutputStream dos = conn.openDataOutputStream();
 	        		
@@ -98,7 +106,14 @@ public class CommunicationNN {
 	        		// try to send data until basestation sent okay
 	        		while(okay == false){
 	        			try{
-	    	                 // send Data
+
+	        		        led6.setRGB(255, 0, 0);
+	        	            led6.setOn();
+	        				Utils.sleep(1000);
+	        	            led6.setOff();	        				
+	        	            
+	        	            // send Data
+	        				System.out.println(allMeas[meas].address);
 	    	        		dos.writeUTF(allMeas[meas].address);
 	    	        		dos.writeFloat(allMeas[meas].frequency);
 	    	        		dos.writeDouble(allMeas[meas].magnitude);
[…]

### 2015-02-16 SHM_JahrSchlich/sensornodeNN/src/MainSpot.java

@@ -24,9 +24,9 @@ public class MainSpot extends MIDlet {
     private double THRESHOLD = 0.2;
 
     // communication
-    private String BASE_NAME = "0014.4F01.0000.74AD";
-    private String[] SENSOR_NAMES = {"0014.4F01.0000.7840",
-    								"0014.4F01.0000.792D"};
+    private String BASE_NAME = "0014.4F01.0000.77BA";
+    private String[] SENSOR_NAMES = {"0014.4F01.0000.792D",
+    								"0014.4F01.0000.7840"};
 //    								"0014.4F01.0000.7AFA"}
     private String ourAddress = System.getProperty("IEEE_ADDRESS");
     private int HOST_PORT = 67;
@@ -36,7 +36,7 @@ public class MainSpot extends MIDlet {
     // neural network
     private int HIDDEN_UNITS = 3;
     
-    private int trainingEvents = 3;
+    private int trainingEvents = 1;
 
     protected void startApp() throws MIDletStateChangeException {
 
@@ -169,7 +169,7 @@ public class MainSpot extends MIDlet {
 			Measurement[] allMeas = new Measurement[othMeas.length+1];
 			allMeas[0]=ownMeas;
 			
-			for (int i=1; i<othMeas.length;i++) allMeas[i]=othMeas[i-1];
+			for (int i=1; i<othMeas.length+1;i++) allMeas[i]=othMeas[i-1];
 			
 			
 			communication.StoreData(allMeas, HOST_PORT_BASE, BASE_NAME);

### 2015-02-17 Paper/PaperJahrSchlich.tex
Frequencies und Überblick in Introduction

@@
 % 
-%\sloppy
+\sloppy
 \setlength{\emergencystretch}{3pt}
@@
 In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
-\colorbox{cyan}{hinzugefügt: }However, when using wireless SHM systems, special attention has to be paid to data transfer.
-To save transmission time and storage space, and to decrease the probability of data loss, the transmitted data is reduced using embedded computing.
+However, when using wireless SHM systems, special attention has to be paid to data transfer.
+To reduce transmission time, storage space and power consumption, and to decrease the probability of data loss, the transmitted data can be reduced using embedded computing.
 
-Over their lifetime, sensors can become inaccurate, faulty, or may even break.
+As data reduction strategy, the natural frequency of a structure can be calculated on the sensor nodes instead of transmitting the raw measurements.
+Changes in the natural frequency result from changes in the vibration behavior, which indicate structural damage \citep{salawu1997detection}.
+The natural frequency can be determined by analysing the frequency spectrum of a vibration of a structure.
+The frequency spectrum shows the relationship between the frequencies and magnitudes---magnitudes represent the influence a certain frequency has in a vibration.
+The natural frequency is defined as the frequency with the highest magnitude.
+While the frequencies are specific for a structure, the scale of the magnitudes are dependent on the structure, the degree of the vibration and the observed location on the structure.
+During a vibration, the magnitudes on different locations are correlated with each other.
+\citep{rao2011fast}.
+
+Over their lifetime, the wireless sensor nodes can become inaccurate, faulty, or may even break.
 A fault can be defined as a defect of a sensor that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
@@
 Physical redundancy, although efficient for sensor fault detection, causes increased installation and maintenance costs due to the multiple installation of sensors. 
-Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors.
+Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors \citep{Smarsly2014}.
 Virtual sensor measurements are computed for each sensor and then compared to the actual measurements. 
-For example, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor
-\citep{Smarsly2014}.
+If the setup of a structure is known, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor.
 
-In this study, analytical redundancy is implemented into wireless sensor nodes based on artificial neural networks.
+Without a priori knowledge, analytical redundancy can be implemented into wireless sensor nodes based on artificial neural networks, that are used to predict measurements.
 Artificial neural networks essentially consist of interconnected data processing units called neurons. 
@@
 Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.
+
+In this study, a wireless SHM system capable of decentralized autonomous fault detection, is presented.
+The sensor nodes calculate the natural frequency of the structure to monitor the structural integrity.
+The corresponding magnitudes are used for sensor fault detection:
+A neural network is implemented into each sensor node and trained---with magnitudes of past vibrations---to map the correlation of the magnitudes calculated by different sensor nodes.
+During runtime, the neural networks compute expected magnitudes for the embedding sensor node.
+A deviation between the expected and the measured magnitude indicates a sensor fault.
 
@@
 \section*{Design and implementation of the wireless structural health monitoring system}
-In the following section, the wireless structural health monitoring system is introduced and the software implementation is described.
-The wireless SHM system consists of wireless sensor nodes and a host computer, linked with a basestation.
+In this section, the wireless structural health monitoring system is introduced and the software implementation is described.
+The wireless SHM system is composed of wireless sensor nodes and a host computer, both linked through a basestation.
 The sensor nodes and the basestation are of type "Oracle Sun SPOT". 
-The Sun SPOTS are equipped with several components.
-Among others, the sensor board includes an accelerometer and eight independent RGB-LEDs.
-The 3-axis digital output accelerometer with sensitivity ranging between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz \citep{eDemo2010}.
+The Sun SPOTS are equipped with several components, e.g. the sensor nodes include an internal accelerometer and eight independent RGB-LEDs.
+The "3-axis digital output accelerometer with three sensitivity ranges" \citep{eDemo2010} between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz.
 
-The SHM system performs the following tasks:
+The SHM system performs the following monitoring tasks:
 1. data acquisition,
@@
 The dataflow is illustrated in \autoref{fig:flow}.
-Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and perform a FFT to determine the characteristics of the oscillation of the structure.
-The results of the FFT are fed into a neural network to detect sensor faults.
+Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and
+determine the natural frequency and the corresponding magnitude of the structure using a fast Fourier transform (FFT).
+The magnitudes of the natural frequencies are fed into a neural network to detect sensor faults.
 The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
-On the host computer, tasks 4 is conducted by storing the data in a MySQL database.
-Task 5 and 6---additional analysis and diagnosis---are conducted on the host computer in further steps.
+On the host computer, task 4 is conducted by storing the data in a MySQL database.
+Task 5 and 6 are conducted on the host computer in further steps.
 
@@
 The neural network represents the structure of the SHM system.
-Each sensor node of the SHM system is represented by either an input neuron or an output neuron, as shown in \autoref{fig:neuralnetwork}.
-During computation, each sensor node embeds a distinct neural network. 
+Each sensor node of the SHM system embeds a distinct neural network; a sensor node is represented by either an input neuron or an output neuron, as shown in \autoref{fig:neuralnetwork}.
 Measurements of other sensor nodes are fed into the input neurons of the neural network; the output neuron returns the predicted measurement of the embedding sensor node.
-Hidden neurons are used for computation and do not represent a physical counterpart.
+Hidden neurons are used for computation and do not have a physical counterpart.
 
@@
 
-The SHM system is programmed object-oriented in Java. 
-Object orientation uses objects, which are instanciated using classes as paradigm. 
+The SHM system is implemented in an object-oriented way using Java programming language. 
+Object orientation uses objects, which are instanciated using classes. 
 A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
-The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
-A package organizes several Java classes that build a program.
+The SHM system is organized in two packages---\texttt{sensornode} and \texttt{basestation} that organize several Java classes belonging to a program.
 
-\autoref{fig:UML-sn} describes the classes of the \texttt{sensor\-node} package.
+\autoref{fig:UML-sn} illustrates the classes of the \texttt{sensor\-node} package.
 The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Measure\-ment}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the Java library "SNIPE" \citep{SNIPE} that implement a framework for neural networks.
@@
 The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
-There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
-Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
-The measured values are stored into an array.
+There are two phases: first, the acceleration is measured with a low sampling rate.
+Once the acceleration exceeds a threshold, the second phase is entered by increasing the sampling rate. 
+The measured values are stored in the sensors nodes in an array.
 The different phases are indicatad by lighting different LEDs.
-The \texttt{FFT} class performs a FFT on the measured accelerations. 
-With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
-Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
-The \texttt{Measure\-ment} class defines analyzed measurements by storing the address of the sensor node, the maximal magnitude, the natural frequency and if a sensor fault was detected during runtime.
-For every measurement, one object of the type \texttt{Measure\-ment} is instantiated.
+The \texttt{FFT} class performs a FFT on the measured accelerations to transform the measured values from time domain into frequency domain. 
+With the transformed data, the magnitudes and the correlating frequencies of the measured vibrations are calculated.
+Finally, the natural frequency is determined by extracting the maximal magnitude.
+The \texttt{Measure\-ment} class defines analyzed measurements by storing the following parameters: the address of the sensor node, the maximal magnitude, the natural frequency and a boolean value that indicates if a sensor fault was detected during runtime.
 Using the SNIPE library, a neural network is defined and created for fault detection.
-The \texttt{Network\-Training} class defines a routine to train the neural network with prior measured \texttt{Measure\-ment} objects.
-The \texttt{Communi\-cation} class opens a radio connection between the sensor node and the basestation to transfer the \texttt{Measure\-ment} objects from the sensor node to the basestation.
-For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
-Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the measurement.
+The \texttt{Network\-Training} class defines a routine to train the neural network with previously measured \texttt{Measure\-ment} objects.
+The \texttt{Communi\-cation} class ensures the wireless communication by opening a radio connection between the sensor node and the basestation to transfer the \texttt{Measure\-ment} objects from the sensor node to the basestation.
+For starting the operation of the sensor node, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
+Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the monitoring tasks.
 
@@
 
-\autoref{fig:UML-bs} describes the classes of the \texttt{base\-station} package.
+\autoref{fig:UML-bs} illustrates the classes of the \texttt{base\-station} package.
 The package \texttt{base\-station} runs on the host computer and operates the basestation.
-It consists of the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
+It includes the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
 The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
@@
 \section*{Laboratory experiments}
+In this section, laboratory experiments, devised to validate the proposed SHM system, are described.
+First, a description of the test setup is given; second, the data aquisition, data processing and sensor fault detection are illustrated; and, finally, the results are discussed.
 
-In the following section, the laboratory experiments are described.
-First, a description of the test setup is given, second, the data aquisition, data processing and sensor fault detection are depicted, and, finally, the results are discussed.
+To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure, as shown in \autoref{fig:teststructure}.
+The test structure is a 4-story shear-frame structure consisting of four steel plates of 25\,cm\,$\times$\,50\,cm\,$\times$\,0.8\,mm.
+The plates are mounted on threaded rods with a vertical clearance of 23\,cm.
+At the bottom of the structure, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm\,$\times$\,30\,cm.
+The SHM system is installed on the test structure by mounting one wireless sensor node to each of the three top stories.
 
-To validate the proposed approach in laboratory experiments, the wireless SHM system is installed on a test structure.
-The test structure is a 4-story shear-frame consisting of four steel plates of 25\,cm\,$\times$\,50\,cm\,$\times$\,0.8\,mm.
-The plates are mounted on threaded rods with a vertical clearance of 23\,cm.
-At the bottom, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm\,$\times$\,30\,cm.
-The SHM system is installed on the test structure by fastening one wireless sensor node to each of the top three stories.
-The laboratory setup is shown in \autoref{fig:teststructure}.
-
-\begin{figure}[ht]
+\begin{figure}
 \centering
-    \includegraphics[scale=0.4]{figures/testsetup.png}
+\setlength\fboxsep{1.5em}\fcolorbox{white}{lightgray}{\includegraphics[scale=0.4]{figures/testsetup.png}}
 \caption{Laboratory setup (Source: own photograph)}
@@
 
-The structure is excited by deflecting and releasing the top of the structure.
-This excitation method ensures a free vibration in natural frequency with little interferences.
-After excitation, when the acceleration threshold is exceeded, the sensor nodes automatically start measuring the acceleration.
+Ensuring a free vibration in natural frequency with little interferences, the structure is excited by a deflection of the top of the structure.
+
+After excitation, i.e. when the acceleration threshold is exceeded as detected by the sensor nodes, the sensor nodes automatically start measuring the acceleration.
 Each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
-By the use of the FFT, the acceleration measurements are converted into the oscillation frequencies and the corresponding magnitudes of the building \cite{rao2011fast}.
+By applying the FFT, the acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the structure \cite{rao2011fast}.
 
@@
 
-%The implementation of the FFT has been verified by plotting the frequency domain of various oscillation events, see \autoref{fig:76hzfft}.
+%The implementation of the FFT has been verified by plotting the frequency domain of various vibration events, see \autoref{fig:76hzfft}.
 %The natural frequencies can be identified by the peaks in the magnitude graph.
@@
 %    \includegraphics{figures/76HzHalf.pdf}
-%    \caption{frequency domain graph of the oscillation}
+%    \caption{frequency domain graph of the vibration}
 %    \label{fig:76hzfft}
@@
 
+The dependability and the accuracy of structural health moni\-toring systems can be affected by sensor faults. 
+In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
+For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
+The predictions are computed using neural networks based on measured sensor data of adjacent sensor nodes.
+In laboratory experiments, devised to validate the proposed approach, several simulated sensor faults are detected.
+These results indicate that the use of neural networks for fault detection increases the dependability and the accuracy of structural health monitoring systems.
+
+In this study, a wireless SHM system capable of decentralized autonomous fault detection, is presented.
+The sensor nodes calculate the natural frequency of the structure to monitor the structural integrity.
+The corresponding magnitudes are used for sensor fault detection:
+A neural network is implemented into each sensor node and trained---with magnitudes of past vibrations---to map the correlation of the magnitudes calculated by different sensor nodes.
+During runtime, the neural networks compute expected magnitudes for the embedding sensor node.
+A deviation between the expected and the measured magnitude indicates a sensor fault.
+
+This paper is organized as follows:
+First, a dezentralized wireless structural health monitoring system is designed and implemented. 
+Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
+Then, the system is tested in laboratory experiments. 
+Finally, the experimental results are discussed and future research directions are proposed.
+
 %----------------------------------------------------------------------------------------

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/CommunicationNN.java
consolenoutput reduziert

@@ -29,23 +29,23 @@ public class CommunicationNN {
             Measurement othMeas = new Measurement();
         	try {     
         		
-        		System.out.println("starting ExchangeData. ");
+        		//System.out.println("starting ExchangeData. ");
         		conn = (RadiostreamConnection)Connector.open("radiostream://" + otherAddress + ":" + hostPort);
         		// set timeout for connection
         		// conn.setTimeout(10000); 
         		DataInputStream dis = conn.openDataInputStream();
         		DataOutputStream dos = conn.openDataOutputStream();
                  
-        		System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
+        		//System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
 
-				System.out.println("try to recieve." );	
+				//System.out.println("try to recieve." );	
 				// receive data from stream
 	            othMeas.address = otherAddress;
 	    		othMeas.magnitude = dis.readDouble();
 	    		othMeas.frequency = dis.readFloat();
 	    		othMeas.error = dis.readInt();
 	    		
-				System.out.println("datastream received." );
+				//System.out.println("datastream received." );
 	    		
 		        // blink third LED
 		        led3.setRGB(0, 255, 255);
@@ -54,7 +54,7 @@ public class CommunicationNN {
 	            // send "okay"
         		dos.writeBoolean(true);
 				dos.flush();
-				System.out.println("received message flushed." );
+				//System.out.println("received message flushed." );
 				
 		        // blink second LED
 		        led2.setRGB(0, 255, 0);
@@ -81,14 +81,14 @@ public class CommunicationNN {
         public void StoreData(Measurement[] allMeas, int hostPort, String baseAddress) {
         	//object for other measurement
         	
-        	System.out.println("allMeas.length= "+ allMeas.length);
+        	//System.out.println("allMeas.length= "+ allMeas.length);
         	
         	int meas = 0;
         	
         	while ( meas < allMeas.length){
         		
 	        	try {   
-	            	System.out.println("meas= "+ meas);
+	            	//System.out.println("meas= "+ meas);
 	            		        		
 	        		
 	        		conn = (RadiostreamConnection)Connector.open("radiostream://" + baseAddress + ":" + hostPort);
@@ -113,16 +113,15 @@ public class CommunicationNN {
 	        	            led6.setOff();	        				
 	        	            
 	        	            // send Data
[…]

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/MainSpot.java

@@ -86,7 +86,7 @@ public class MainSpot extends MIDlet {
 			double[] frequency = FFT.calcFreq(transform, SAMPLERATE);
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
-			System.out.println("I did my FFT!");
+			//System.out.println("I did my FFT!");
 			
 			// write magnitude to NN output array
 			desiredOutputs[event][0] = ownMeas.magnitude; 
@@ -111,7 +111,7 @@ public class MainSpot extends MIDlet {
 		// NEURAL NETWORK TRAINING
     	// create lessonMeasurement
     	TrainingSampleLesson lesson = new TrainingSampleLesson(inputs, desiredOutputs);
-    	System.out.println("sample lesson created");	    	
+    	//System.out.println("sample lesson created");	    	
 
         // train the neural network
         NetworkTraining training = new NetworkTraining();
@@ -135,7 +135,7 @@ public class MainSpot extends MIDlet {
 			double[] frequency = FFT.calcFreq(transform, SAMPLERATE);
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
-			System.out.println("I did my FFT!");
+			//System.out.println("I did my FFT!");
 			
 			double[] magnitudes = new double[numberOfSensors];
 			// COMMUNICATION

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/MainSpot.java
Addded NeualNetwork Measurement to datastream

@@ -161,20 +161,21 @@ public class MainSpot extends MIDlet {
 			double threshold = 0.01;
 			
 			if (errorNN > threshold) ownMeas.error = -1;
+			Measurement NNmeas = new Measurement("NeuralNetwork", expMeas, 0, 0);
+			
+			
+			if (errorNN > threshold) ownMeas.error = -1;
 			else ownMeas.error = 1;
-			
-			
+
 			//SEND DATA TO BASESTATION
+			Measurement[] allMeas = new Measurement[othMeas.length+2];
+			allMeas[0]=NNmeas;
+			allMeas[1]=ownMeas;
 			
-			Measurement[] allMeas = new Measurement[othMeas.length+1];
-			allMeas[0]=ownMeas;
-			
-			for (int i=1; i<othMeas.length+1;i++) allMeas[i]=othMeas[i-1];
-			
+			for (int i=2; i<othMeas.length+2;i++) allMeas[i]=othMeas[i-2];
 			
 			communication.StoreData(allMeas, HOST_PORT_BASE, BASE_NAME);
-
-			Utils.sleep(2000);			
+			Utils.sleep(2000);		
         }
     }
     

### 2015-02-17 SHM_JahrSchlich/sensornode/src/FFT.java
kaputte FFT gefixt

@@ -179,11 +179,12 @@ public class FFT {
     		newMagnitude[i] = magnitude[i];
     	
     	// determine maximal magnitude
-		double maxMagnitude = magnitude[0];
+		double maxMagnitude = newMagnitude[0];
 		int indexMax = 0;
-		for ( int i = 1; i < magnitude.length; i++) {
-		    if ( magnitude[i] > maxMagnitude) {
-		    	maxMagnitude = magnitude[i];
+		for ( int i = 1; i < newMagnitude.length; i++) {
+			//System.out.println(newMagnitude[i]);
+		    if ( newMagnitude[i] > maxMagnitude) {
+		    	maxMagnitude = newMagnitude[i];
 		    	indexMax = i;
 		    }
 		}
@@ -193,7 +194,7 @@ public class FFT {
 		//System.out.println("maxMagnitude= "+ maxMagnitude+
 		//									", maxFrequency= " + maxFrequency);
 		
-		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency, 0);
+		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency,0);
         return measurement;
     }
 }

### 2015-02-17 SHM_JahrSchlich/sensornode/src/MainSpot.java

@@ -18,10 +18,10 @@ public class MainSpot extends MIDlet {
 
     private int HOST_PORT = 67;
     private int SAMPLE_PERIOD_LISTENING = 1 * 500;  // in milliseconds
-    private int SAMPLE_PERIOD_MEASURING = 13;  // in milliseconds
-    private int ARRAY_LENGTH = 128;
+    private int SAMPLE_PERIOD_MEASURING = 20;  // in milliseconds
+    private int ARRAY_LENGTH = 256;
     private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
-    private double THRESHOLD = 0.2;
+    private double THRESHOLD = 0.15;
     private int NUMBER_OF_OTHER_SENSORS = 1;
     
     

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/FFT.java

@@ -179,11 +179,12 @@ public class FFT {
     		newMagnitude[i] = magnitude[i];
     	
     	// determine maximal magnitude
-		double maxMagnitude = magnitude[0];
+		double maxMagnitude = newMagnitude[0];
 		int indexMax = 0;
-		for ( int i = 1; i < magnitude.length; i++) {
-		    if ( magnitude[i] > maxMagnitude) {
-		    	maxMagnitude = magnitude[i];
+		for ( int i = 1; i < newMagnitude.length; i++) {
+			//System.out.println(newMagnitude[i]);
+		    if ( newMagnitude[i] > maxMagnitude) {
+		    	maxMagnitude = newMagnitude[i];
 		    	indexMax = i;
 		    }
 		}

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/MainSpot.java

@@ -18,10 +18,10 @@ public class MainSpot extends MIDlet {
 	
 	// measurement
     private int SAMPLE_PERIOD_LISTENING = 1 * 500;  // in milliseconds
-    private int SAMPLE_PERIOD_MEASURING = 13;  // in milliseconds
-    private int ARRAY_LENGTH = 128;
+    private int SAMPLE_PERIOD_MEASURING = 20;  // in milliseconds
+    private int ARRAY_LENGTH = 256;
     private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
-    private double THRESHOLD = 0.2;
+    private double THRESHOLD = 0.15;
 
     // communication
     private String BASE_NAME = "0014.4F01.0000.77BA";
@@ -36,7 +36,7 @@ public class MainSpot extends MIDlet {
     // neural network
     private int HIDDEN_UNITS = 3;
     
-    private int trainingEvents = 1;
+    private int trainingEvents = 3;
 
     protected void startApp() throws MIDletStateChangeException {
 
@@ -87,6 +87,7 @@ public class MainSpot extends MIDlet {
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
 			//System.out.println("I did my FFT!");
+			System.out.println("training: magnitude: " + ownMeas.magnitude + " frequency: " + ownMeas.frequency);
 			
 			// write magnitude to NN output array
 			desiredOutputs[event][0] = ownMeas.magnitude; 
@@ -99,6 +100,8 @@ public class MainSpot extends MIDlet {
 				//talk to next sensor in list
 				String otherAddress = SENSOR_NAMES[sensor];
 				othMeas[sensor] = communication.ReceiveData(otherAddress, HOST_PORT);
+				
+				System.out.println("communication: "+othMeas[sensor].address+ "magnitude: " + othMeas[sensor].magnitude + " frequency: " + othMeas[sensor].frequency);
 				
 				// write magnitude to NN input array
 				inputs[event][sensor] = othMeas[sensor].magnitude;	
@@ -136,6 +139,7 @@ public class MainSpot extends MIDlet {
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
 			//System.out.println("I did my FFT!");
[…]

### 2015-02-17 SHM_JahrSchlich/basestation/src/basestation/DatabaseHandler.java
Verzeichnisse bearbeitet, communication repaired

@@ -46,7 +46,6 @@ public class DatabaseHandler {
 					+ "SpotID VARCHAR(19), "
 					+ "Magnitude DOUBLE, "
 					+ "Frequency FLOAT, "
-					+ "ExpectedMagnitude DOUBLE, "
 					+ "Error INT)");
 			statement.close();
 		}
@@ -64,7 +63,7 @@ public class DatabaseHandler {
 		System.out.println("Writing measurement into Database");
 		PreparedStatement preparedStatement = dbConn.prepareStatement("INSERT INTO "
 				+ tableName
-				+ " (Time, SpotID, Magnitude, Frequency, ExpectedMagnitude, Error) VALUES (?,?,?,?,?,?)");
+				+ " (Time, SpotID, Magnitude, Frequency, Error) VALUES (?,?,?,?,?)");
 			preparedStatement.setTimestamp(1, now);
 			preparedStatement.setString(2, meas.address);
 			preparedStatement.setDouble(3, meas.magnitude);

### 2015-02-17 SHM_JahrSchlich/basestation/src/basestation/MainBase.java

@@ -17,7 +17,7 @@ import javax.microedition.io.*;
 public class MainBase {
     // Broadcast port on which we listen for sensor samples
     private static final int HOST_PORT = 68;
-    private static final String DATABASE_URL = "jdbc:mysql://localhost:3306/";
+    private static final String DATABASE_URL = "jdbc:mysql://141.54.51.238:3306/";
     private static final String DATABASE_NAME = "SunSPOT";
     private static final String DATABASE_USER = "shm";
     private static final String DATABASE_PASSWORD = "pass";
@@ -52,10 +52,13 @@ public class MainBase {
 				// receive data from stream
 	            othMeas.address = dis.readUTF();
 	            System.out.println("Adress: " + othMeas.address);
-	    		othMeas.magnitude = dis.readDouble();
-	    		System.out.println("Magnitude: " + othMeas.magnitude);
+	            
 	    		othMeas.frequency = dis.readFloat();
 	    		System.out.println("Frequency: " + othMeas.frequency);
+	    		
+	    		othMeas.magnitude = dis.readDouble();
+	    		System.out.println("Magnitude: " + othMeas.magnitude);
+	    		
 	    		othMeas.error = dis.readInt();
 	    		System.out.println("Error: " + othMeas.error);
 	    		
@@ -73,7 +76,6 @@ public class MainBase {
 	    		dos.close();
 	    		conn.close();
 				
-	    		/*
 				//DATABASE
 				//Create a DatabaseHandler, the DatabaseHandler will take care of 
 		        // database interaction
@@ -89,12 +91,12 @@ public class MainBase {
 		        // Main data collection loop
 	            try {
 	                // Write measurement to Database
+	            	System.out.println("Writing to database!");
 	                dbHandler.writeMeasurement(TABLE_NAME, othMeas);
 	            } catch (Exception e) {
 	                System.err.println("Caught " + e +  " while reading sensor samples.");
 	                throw e;
 	            }
-	            */
 			
 	    	} catch (Exception e) {   
 	    		 System.err.println("Caught " + e + " in exchanging data.");

### 2015-02-17 SHM_JahrSchlich/sensornode/src/sensornode/CommunicationNormal.java

@@ -38,7 +38,8 @@ public class CommunicationNormal {
         		boolean okay = false;
         		while(okay == false){
         			try{
-	            		dos.writeFloat(ourMeas.frequency);
+	            		
+	    				dos.writeFloat(ourMeas.frequency);
 	    				dos.writeDouble(ourMeas.magnitude);
 	    				dos.writeInt(ourMeas.error);
 	    				dos.flush();

### 2015-02-17 SHM_JahrSchlich/sensornode/src/sensornode/MainSpot.java

@@ -52,7 +52,7 @@ public class MainSpot extends MIDlet {
 			double[] frequency = FFT.calcFreq(transform, SAMPLERATE);
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
-			System.out.println("I did my FFT!");
+			System.out.println("magnitude: " + ownMeas.magnitude + " frequency: " + ownMeas.frequency);
 			
 			// exchange measurement with measurement of other sensors
 			communication.ExchangeData(ownMeas, otherAddress, HOST_PORT);

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/CommunicationNN.java

@@ -40,9 +40,10 @@ public class CommunicationNN {
 
 				//System.out.println("try to recieve." );	
 				// receive data from stream
+        		
 	            othMeas.address = otherAddress;
+	            othMeas.frequency = dis.readFloat();
 	    		othMeas.magnitude = dis.readDouble();
-	    		othMeas.frequency = dis.readFloat();
 	    		othMeas.error = dis.readInt();
 	    		
 				//System.out.println("datastream received." );

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/FFT.java

@@ -179,12 +179,11 @@ public class FFT {
     		newMagnitude[i] = magnitude[i];
     	
     	// determine maximal magnitude
-		double maxMagnitude = newMagnitude[0];
+		double maxMagnitude = magnitude[0];
 		int indexMax = 0;
-		for ( int i = 1; i < newMagnitude.length; i++) {
-			//System.out.println(newMagnitude[i]);
-		    if ( newMagnitude[i] > maxMagnitude) {
-		    	maxMagnitude = newMagnitude[i];
+		for ( int i = 1; i < magnitude.length; i++) {
+		    if ( magnitude[i] > maxMagnitude) {
+		    	maxMagnitude = magnitude[i];
 		    	indexMax = i;
 		    }
 		}

### 2015-02-17 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/MainSpot.java

@@ -21,7 +21,7 @@ public class MainSpot extends MIDlet {
     private int SAMPLE_PERIOD_MEASURING = 20;  // in milliseconds
     private int ARRAY_LENGTH = 256;
     private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
-    private double THRESHOLD = 0.15;
+    private double THRESHOLD = 0.2;
 
     // communication
     private String BASE_NAME = "0014.4F01.0000.77BA";
@@ -36,7 +36,7 @@ public class MainSpot extends MIDlet {
     // neural network
     private int HIDDEN_UNITS = 3;
     
-    private int trainingEvents = 3;
+    private int trainingEvents = 1;
 
     protected void startApp() throws MIDletStateChangeException {
 
@@ -87,7 +87,6 @@ public class MainSpot extends MIDlet {
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
 			//System.out.println("I did my FFT!");
-			System.out.println("training: magnitude: " + ownMeas.magnitude + " frequency: " + ownMeas.frequency);
 			
 			// write magnitude to NN output array
 			desiredOutputs[event][0] = ownMeas.magnitude; 
@@ -100,8 +99,6 @@ public class MainSpot extends MIDlet {
 				//talk to next sensor in list
 				String otherAddress = SENSOR_NAMES[sensor];
 				othMeas[sensor] = communication.ReceiveData(otherAddress, HOST_PORT);
-				
-				System.out.println("communication: "+othMeas[sensor].address+ "magnitude: " + othMeas[sensor].magnitude + " frequency: " + othMeas[sensor].frequency);
 				
 				// write magnitude to NN input array
 				inputs[event][sensor] = othMeas[sensor].magnitude;	
@@ -139,7 +136,6 @@ public class MainSpot extends MIDlet {
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
 			//System.out.println("I did my FFT!");
-			System.out.println("measurement: magnitude: " + ownMeas.magnitude + " frequency: " + ownMeas.frequency);
 			
 			double[] magnitudes = new double[numberOfSensors];
 			// COMMUNICATION

### 2015-02-18 SHM_JahrSchlich/sensornode/src/sensornode/Measurement.java
attribut prediction added to measurment

@@ -5,6 +5,7 @@ public class Measurement{
 	public float frequency;
 	public double magnitude;
 	public int error;
+	public double prediction = 0;
 	
 	// initiates a measurement
 	Measurement(String a, double m, float f, int e){

### 2015-02-18 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/CommunicationNN.java

@@ -117,6 +117,7 @@ public class CommunicationNN {
 	        				dos.writeUTF(allMeas[meas].address);
 	    	        		dos.writeFloat(allMeas[meas].frequency);
 	    	        		dos.writeDouble(allMeas[meas].magnitude);
+	    	        		dos.writeDouble(allMeas[meas].prediction);
 	    	        		dos.writeInt(allMeas[meas].error);
 	    	        		
 	    					dos.flush();

### 2015-02-18 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/Measurement.java

@@ -2,26 +2,27 @@ package sensornodeNN;
 
 public class Measurement{
 	
-		public String address;
-		public float frequency;
-		public double magnitude;
-		public int error;
+	public String address;
+	public float frequency;
+	public double magnitude;
+	public int error;
+	public double prediction = 0;
+	
+	// initiates a measurement
+	Measurement(String a, double m, float f, int e){
 		
-		// initiates a measurement
-		Measurement(String a, double m, float f, int e){
-			
-			address = a;
-			magnitude = m;
-			frequency = f;
-			error = e;
-		}
+		address = a;
+		magnitude = m;
+		frequency = f;
+		error = e;
+	}
+	
+	// initiates a measurement
+	Measurement(){
 		
-		// initiates a measurement
-		Measurement(){
-			
-			address = "a";
-			magnitude = 0;
-			frequency = 0;
-			error = 0;
-		}		
+		address = "a";
+		magnitude = 0;
+		frequency = 0;
+		error = 0;
+	}		
 }

### 2015-02-18 SHM_JahrSchlich/basestation/src/basestation/Measurement.java
Modified NN measurement handling

@@ -6,6 +6,7 @@ public class Measurement{
 	public float frequency;
 	public double magnitude;
 	public int error;
+	public double prediction = 0;
 	
 	// initiates a measurement
 	Measurement(String a, double m, float f, int e){

### 2015-02-18 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/CommunicationNN.java

@@ -142,7 +142,5 @@ public class CommunicationNN {
 	        		 System.err.println("Caught " + e + " in exchanging data 2.");
 	        	}
         	}
-
-        	
         }
 }

### 2015-02-18 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/FFT.java

@@ -179,11 +179,11 @@ public class FFT {
     		newMagnitude[i] = magnitude[i];
     	
     	// determine maximal magnitude
-		double maxMagnitude = magnitude[0];
+		double maxMagnitude = newMagnitude[0];
 		int indexMax = 0;
-		for ( int i = 1; i < magnitude.length; i++) {
-		    if ( magnitude[i] > maxMagnitude) {
-		    	maxMagnitude = magnitude[i];
+		for ( int i = 1; i < newMagnitude.length; i++) {
+		    if ( newMagnitude[i] > maxMagnitude) {
+		    	maxMagnitude = newMagnitude[i];
 		    	indexMax = i;
 		    }
 		}

### 2015-02-18 Paper/PaperJahrSchlich.tex
lab ex outline

@@
 %----------------------------------------------------------------------------------------
-
 \section*{Introduction}
@@
 
-\begin{figure}[hbtp]
+\begin{figure}[hbt]
     \centering
@@
 The \texttt{Measure\-ment} class defines analyzed measurements by storing the following parameters: the address of the sensor node, the maximal magnitude, the natural frequency and a boolean value that indicates if a sensor fault was detected during runtime.
-Using the SNIPE library, a neural network is defined and created for fault detection.
+Using the SNIPE library, a neural network is described and created for fault detection.
 The \texttt{Network\-Training} class defines a routine to train the neural network with previously measured \texttt{Measure\-ment} objects.
@@
 
-\begin{figure}[htbp]
-    \centering
-    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
-    \caption{Class diagram of the \texttt{basestation} package}
-    \label{fig:UML-bs}
-\end{figure}
+%\begin{figure}[htbp]
+%    \centering
+%    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
+%    \caption{Class diagram of the \texttt{basestation} package}
+%    \label{fig:UML-bs}
+%\end{figure}
 
@@
 The SHM system is installed on the test structure by mounting one wireless sensor node to each of the three top stories.
+Ensuring a free vibration in natural frequency with little interferences, the structure is excited by a deflection of the top of the structure.
 
@@
 
-Ensuring a free vibration in natural frequency with little interferences, the structure is excited by a deflection of the top of the structure.
+Several test runs were conducted.
+During each test run, all sensor nodes run simultaneously.
+A test run includes a training phase and a data collection phase.
+The training phase consists of the description, the creation and the training of a neural network.
+During the training of the neural network, several measurements are performed and used as training input.
+A measurement includes the excitation of the structure, acceleration sampling, calculation of the natural frequency and corresponding magnitude, and data exchange with the other sensor nodes.
 
-After excitation, i.e. when the acceleration threshold is exceeded as detected by the sensor nodes, the sensor nodes automatically start measuring the acceleration.
-Each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
-By applying the FFT, the acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the structure \cite{rao2011fast}.
+The data collection phase consists of any desired number of measurements, sensor fault detection, and data storage.
+After every measurement, the measured magnitudes of the other sensor nodes are used as input of the neural network to predict the expected magnitude of the embedding sensor nodes for sensor fault detection.
+The deviation of the expected magnitude and the measured magnitude is calculated.
+If the deviation exceeds a threshold, a sensor fault is assumed.
+After the sensor fault detection, the following are stored into the database: the time of the measurement, the natural frequency, the measured magnitude, the expected magnitude, the deviation,  and a boolean value that indicates if a sensor fault was detected.
+
+In several test runs, the properties of the neural network are optimized.
+
+has been tested with varying numbers of training measurements (\autoref{fig:traininputs}).
+With increasing numbers of training measurement, the prediction accuracy increases.
+
+The neural network has been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values.
+
+The implementation of the FFT has been verified by plotting the frequency sprectrum of various vibration measurements, (\autoref{fig:76hzfft}).
+The natural frequencies can be identified by the peaks in the magnitude graph.
+Each sensor node stores the calculated first natural frequency and the corresponding magnitude in a \texttt{Measurement} object and transmits the object to the other sensor nodes.
 
@@
 \item \textbf{neuronales Netz}
-\item TRAININGSPHASE
-\item Austausch der gemessenen FFT zwischen den Sensoren
-\item trainieren des neuronalen Netzes, bis keine Verbesserung bei Vorhersagegenauigkeit eintritt
-\item MESSPHASE
-\item Austausch der gemessenen FFT zwischen den Sensoren
-\item Nutzen des neuronalen Netzes, um eventuellen Fehler zu finden
+%\item TRAININGSPHASE
+%\item Austausch der gemessenen FFT zwischen den Sensoren
+%\item trainieren des neuronalen Netzes, bis keine Verbesserung bei Vorhersagegenauigkeit eintritt
+%\item MESSPHASE
+%\item Austausch der gemessenen FFT zwischen den Sensoren
+%\item Nutzen des neuronalen Netzes, um eventuellen Fehler zu finden
 \item KÜNSTLICHEN FEHLER ERZEUGEN
@@
 
-%The implementation of the FFT has been verified by plotting the frequency domain of various vibration events, see \autoref{fig:76hzfft}.
-%The natural frequencies can be identified by the peaks in the magnitude graph.
-%Each sensor node stores the calculated first natural frequency and the corresponding magnitude in a \texttt{Measurement} object and transmits the object to the other sensor nodes.
+%After excitation, i.e. when the acceleration threshold is exceeded as detected by the sensor nodes, the sensor nodes automatically start measuring the acceleration.
+%Each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
+%By applying the FFT, the acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the structure \cite{rao2011fast}.
 
@@
 
-The dependability and the accuracy of structural health moni\-toring systems can be affected by sensor faults. 
-In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
-For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
-The predictions are computed using neural networks based on measured sensor data of adjacent sensor nodes.
-In laboratory experiments, devised to validate the proposed approach, several simulated sensor faults are detected.
-These results indicate that the use of neural networks for fault detection increases the dependability and the accuracy of structural health monitoring systems.
-
 In this study, a wireless SHM system capable of decentralized autonomous fault detection, is presented.
@@
 
-This paper is organized as follows:
-First, a dezentralized wireless structural health monitoring system is designed and implemented. 
-Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
-Then, the system is tested in laboratory experiments. 
-Finally, the experimental results are discussed and future research directions are proposed.
-
 %----------------------------------------------------------------------------------------

### 2015-02-18 Paper/PaperJahrSchlich.tex
lab tests, mean error figure

@@
 
-%\begin{figure}[htbp]
-%    \centering
-%    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
-%    \caption{Class diagram of the \texttt{basestation} package}
-%    \label{fig:UML-bs}
-%\end{figure}
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width = \textwidth]{figures/uml-basestation.pdf}
+    \caption{Class diagram of the \texttt{basestation} package}
+    \label{fig:UML-bs}
+\end{figure}
 
@@
 
-In several test runs, the properties of the neural network are optimized.
+The neural network has been tested with varying numbers of training measurements (\autoref{fig:traininputs}).
+With increasing numbers of training measurements, the prediction accuracy of the expected magnitude increases.
 
-has been tested with varying numbers of training measurements (\autoref{fig:traininputs}).
-With increasing numbers of training measurement, the prediction accuracy increases.
+\begin{figure}[htb]
+    \centering
+    %\includegraphics{figures/traininputs.pdf}
+    \caption{correlation between number of traing measurements and mean error during collection phase}
+    \label{fig:traininputs}
+\end{figure}
 
-The neural network has been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values.
+To validate the autonomous sensor fault detection, a sensor fault was simulated by shifting and rotating one of the sensor nodes after completion of the training phase.
+The expected magnitudes deviated from the measured magnitudes (\autoref{tab:sensorfault}.
+The SHM system recognized the deviation and the boolean was set to "TRUE" accordingly.
 
-The implementation of the FFT has been verified by plotting the frequency sprectrum of various vibration measurements, (\autoref{fig:76hzfft}).
-The natural frequencies can be identified by the peaks in the magnitude graph.
-Each sensor node stores the calculated first natural frequency and the corresponding magnitude in a \texttt{Measurement} object and transmits the object to the other sensor nodes.
+\begin{table}[hb]
+	\centering
+	\caption{Deviation between measured magnitude and expected magnitude with and without simulated sensor faults}
+	\begin{tabular}{c c c c}
+		\toprule
+		Sensor fault & Measured magnitude & Expected magnitude & Deviation [\%]\\ 
+		\midrule
+		  & 2.1502 & 2.1389 & 0.5292 \\
+		FALSE & 2.0817 & 2.0890 & 0.3529 \\
+		  & 1.8663 & 1.8631 & 0.1726 \\
+		\midrule	
+		 & 1.8161 & 2.0725 & 12.3715\\
+		TRUE & 1.6665 & 1.9404 & 14.1156\\
+		 & 1.8238 & 2.1687 & 15.9035\\
+		\bottomrule
+	\end{tabular}
+	\label{tab:sensorfault}
+\end{table}
 
-\begin{itemize}
-\item \textbf{neuronales Netz}
-%\item TRAININGSPHASE
-%\item Austausch der gemessenen FFT zwischen den Sensoren
-%\item trainieren des neuronalen Netzes, bis keine Verbesserung bei Vorhersagegenauigkeit eintritt
-%\item MESSPHASE
-%\item Austausch der gemessenen FFT zwischen den Sensoren
-%\item Nutzen des neuronalen Netzes, um eventuellen Fehler zu finden
-\item KÜNSTLICHEN FEHLER ERZEUGEN
-\item TABELLE MIT ERMITTELTEN / GEMESSENEN WERTEN; MIT / OHNE FEHLERN
-\end{itemize}
-
-%The neural network has been validated by repeated tests---using varying excitation forces, varying sampling rates, and varying quantities of measured values.
-
-%After excitation, i.e. when the acceleration threshold is exceeded as detected by the sensor nodes, the sensor nodes automatically start measuring the acceleration.
-%Each sensor node performs a FFT, once sufficient acceleration measurements have been collected.
-%By applying the FFT, the acceleration measurements are converted into the vibration frequencies and the corresponding magnitudes of the structure \cite{rao2011fast}.
-
-%\begin{figure}[htb]
-%    \centering
-%    \includegraphics{figures/76HzHalf.pdf}
-%    \caption{frequency domain graph of the vibration}
-%    \label{fig:76hzfft}
-%\end{figure}
-
-After the sensor fault detection, the \texttt{Measurement} objects of all sensor nodes ar transmitted to the basestation.
-The values and detected sensor faults are stored in a MySQL database and availeble for further analysis.
-
-%Example results are shown in \autoref{tab:lab-ex}.
-%The identified magnitudes increase with ascending sensor node position from top to bottom, corresponding to the higher deflection at the top of the structure.
-%
-%\textit{Through FFT analysis of sampling events, the data traffic is decreased from 512 doubles per sampling event to two doubles per sampling event.
-%A number stored in the used data format double occupies 8\,bytes of space.
-%The data traffic is decreased by (512$-$2)$\times$8\,B\,=\,4080\,B\,=\,3,98\,MB per sampling event, decreasing data traffic by 99.6\,\%.}
-
-%\begin{table}[hb]
-%	\centering
-%	\begin{tabular}{l l c c c}
-%		\toprule
-%		ID & Position & Magnitude & Natural frequency [Hz] & Sampling rate [Hz]\\ 
-%		\midrule
-%		1 & top & 3.78 & 2.2461 & 76\\ 
-%		1 & middle & 3.23 & 2.2461 & 76\\ 
-%		1 & bottom & 1.97 & 2.2461 & 76\\ 
-%		\midrule	
-%		2 & top & 5.74 & 2.2536 & 76\\  
-%		2 & middle & 4.38 & 2.2536 & 76\\ 
-%		2 & bottom & 2.55 & 2.2536 & 76\\
-%		\midrule
-%		3 & top & 8.20 & 2.2536 & 76\\
-%		3 & middle & 6.35 & 2.2536 & 76\\ 
-%		3 & bottom & 3.79 & 2.2536 & 76\\
-%		\bottomrule
-%	\end{tabular}
-%	\caption{Natural frequencies of the test structure}
-%	\label{tab:lab-ex}
-%\end{table}
+Through the embedded FFT analysis of sampling events, the data traffic is decreased from 512 values per measurement and sensor node to 4 values per sampling event.
+A number stored in the used data format double occupies 8\,bytes of space.
+The data traffic is decreased by (512$-$4)$\times$8\,B\,=\,4064\,B\,=\,3.97\,MB per sampling event, decreasing data traffic by 99.3\,\%.
 

### 2015-02-18 Paper/PaperJahrSchlich.tex
Vorbereitung für Bilder
siehe oben

@@
 
+\usepackage{subcaption}
+
 \usepackage{titlesec}
@@
 
-\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue]{hyperref}
+\usepackage[colorlinks, linkcolor=black, citecolor=blue, urlcolor=blue, bookmarks]{hyperref}
 
@@
 
-\begin{figure}
+\begin{figure}[htb]
+
 	\centering
-\setlength\fboxsep{1.5em}\fcolorbox{white}{lightgray}{\includegraphics[scale=0.4]{figures/testsetup.png}}
-\caption{Laboratory setup (Source: own photograph)}
+
+	\begin{subfigure}{0.49\textwidth}
+		\includegraphics[width=\textwidth]{figures/testsetup.png}
+    	\caption{Regular sensor arrangement}
 		\label{fig:teststructure}
+    \end{subfigure}
+  	\begin{subfigure}{0.49\textwidth}
+		\includegraphics[width=\textwidth]{figures/testsetup.png}
+    	\caption{Simulated sensor fault}
+        \label{fig:sensorfault}
+	\end{subfigure}
+  	\caption{Laboratory setup (Source: own photographs)}
 \end{figure}
@@
 To validate the autonomous sensor fault detection, a sensor fault was simulated by shifting and rotating one of the sensor nodes after completion of the training phase.
-The expected magnitudes deviated from the measured magnitudes (\autoref{tab:sensorfault}.
+The expected magnitudes deviated from the measured magnitudes (\autoref{tab:sensorfault}).
 The SHM system recognized the deviation and the boolean was set to "TRUE" accordingly.

### 2015-02-18 Paper/PaperJahrSchlich.tex
Final corrections of the figures
summary

@@
 With increasing numbers of training measurements, the prediction accuracy of the expected magnitude increases.
+With 24 training measurements, a minimum deviation---between the expected magnitude and the measured magnitude---of average 0.47\% has been reached.
 
@@
 The neural networks have been trained to predict expected sensor measurements to be compared to actual measurments, in order to detect sensor faults.
-The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.
-
-In this study, a wireless SHM system capable of decentralized autonomous fault detection, is presented.
-The sensor nodes calculate the natural frequency of the structure to monitor the structural integrity.
-The corresponding magnitudes are used for sensor fault detection:
-A neural network is implemented into each sensor node and trained---with magnitudes of past vibrations---to map the correlation of the magnitudes calculated by different sensor nodes.
-During runtime, the neural networks compute expected magnitudes for the embedding sensor node.
-A deviation between the expected and the measured magnitude indicates a sensor fault.
+The structural integrity of the monitored structure are detected by observing the natural frequency.
+To verify the proposed approach, the SHM system has been installed on a test structure to be tested in laboratory experiments.
+During several testruns, the prediction accuracy could be reduced up to a deviation of average 0.47\%.
+Simulated sensor faults have been detected reliably.
+Through decentralized data processing, embedded in each sensor node, the data traffic has been reduced by 99.3\,\%.
+It has been proven, that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.
 

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/ErrorMeasurement.java
letzte Änderung im Code

@@ -65,12 +65,22 @@ public class ErrorMeasurement {
 		for (int i = 0; i < inputs.length; i++) {
 
 			double[] outputs = net.propagate(inputs[i]);
+			
+			System.out.println("outputs: " + outputs[0]);
 
 			for (int j = 0; j < outputs.length; j++) {
 				double temp = 0.0;
+				
 				temp = desiredOutputs[i][j] - outputs[j];
-				temp = temp * temp;
-				temp = Math.sqrt(temp);
+
+				System.out.println( i + " temp: " +temp);
+				
+				
+				temp = Math.abs(temp);
+				
+				//temp = temp * temp;
+				//temp = Math.sqrt(temp);
+				
 				error += temp;
 			}
 		}

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/MainSpot.java

@@ -17,7 +17,7 @@ import com.sun.spot.service.BootloaderListenerService;
 public class MainSpot extends MIDlet {
 	
 	// measurement
-    private int SAMPLE_PERIOD_LISTENING = 1 * 500;  // in milliseconds
+    private int SAMPLE_PERIOD_LISTENING = 1 * 100;  // in milliseconds
     private int SAMPLE_PERIOD_MEASURING = 20;  // in milliseconds
     private int ARRAY_LENGTH = 256;
     private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
@@ -36,7 +36,7 @@ public class MainSpot extends MIDlet {
     // neural network
     private int HIDDEN_UNITS = 3;
     
-    private int trainingEvents = 1;
+    private int trainingEvents = 6;
 
     protected void startApp() throws MIDletStateChangeException {
 
@@ -90,6 +90,7 @@ public class MainSpot extends MIDlet {
 			
 			// write magnitude to NN output array
 			desiredOutputs[event][0] = ownMeas.magnitude; 
+			System.out.println(ownMeas.address +" magnitude: " + ownMeas.magnitude);
 			
 			// COMMUNICATION
         	
@@ -102,6 +103,12 @@ public class MainSpot extends MIDlet {
 				
 				// write magnitude to NN input array
 				inputs[event][sensor] = othMeas[sensor].magnitude;	
+
+				
+				System.out.println(othMeas[sensor].address +" magnitude: " + othMeas[sensor].magnitude);
+
+				Utils.sleep(1000);	
+			
 			}
 			
 			event++;
@@ -111,7 +118,7 @@ public class MainSpot extends MIDlet {
 		// NEURAL NETWORK TRAINING
     	// create lessonMeasurement
     	TrainingSampleLesson lesson = new TrainingSampleLesson(inputs, desiredOutputs);
-    	//System.out.println("sample lesson created");	    	
+    	System.out.println("sample lesson created");	    	
 
         // train the neural network
         NetworkTraining training = new NetworkTraining();
@@ -152,29 +159,25 @@ public class MainSpot extends MIDlet {
 			
 			
 			//FAULT DETECTION
-			double expMeas = nn.propagate(magnitudes)[0];
-			double errorNN = Math.abs(expMeas-ownMeas.magnitude) /  ownMeas.magnitude;
[…]

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/NetworkTraining.java

@@ -3,12 +3,12 @@ package sensornodeNN;
 public class NetworkTraining {
 	  
 
-	public static void Train(NeuralNetwork net, TrainingSampleLesson lesson) {
+	public void Train(NeuralNetwork net, TrainingSampleLesson lesson) {
 
 		 
-		int runs = 20000; // number of random patterns to train
-		double eta = 0.2; //learning rate
-		double threshold = 0.99;
+		int runs = 100000; // number of random patterns to train
+		double eta = 0.5; //learning rate
+		double threshold = 1e-5;
 	 
 		
 		// Train neural network until error falls below limit
@@ -18,13 +18,29 @@ public class NetworkTraining {
 	
 		do{
 			numOfPhases++;
+			
 			error = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
+			
+			System.out.println("Error 1: "+ error);	
+			
+			eta = eta/2;
+			
 			net.trainBackpropagationOfError(lesson, runs, eta);
+			
+			//net.trainResilientBackpropagation(lesson, runs, false);
+			
+			
+			
 			newerror = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
+			
 			System.out.println("Error after training phase "+ numOfPhases + " = " + newerror );
-		}while(newerror / error > threshold);
+			//System.out.println("Abbruchbedingung "+ Math.abs(newerror-error)*100 / newerror);
+			
+			if (numOfPhases > 10) break;
+			
+		}while(newerror > threshold);
 		
-		System.out.println(numOfPhases + "training phases. ");
+		System.out.println(numOfPhases + " training phases. ");
 
 		}
 

### 2015-02-19 SHM_JahrSchlich/sensornode/src/sensornode/AccelerationSampler.java
code aufgeräumt sensor nodes

@@ -1,3 +1,8 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornode;
 
 import com.sun.spot.resources.Resources;
@@ -6,13 +11,9 @@ import com.sun.spot.util.Utils;
 import java.lang.Math;
 import com.sun.spot.resources.transducers.ITriColorLED;
 
-
-
-public class AccelerationSampler 
-{
+public class AccelerationSampler {
 	IAccelerometer3D accelerometer = 
 			(IAccelerometer3D) Resources.lookup(IAccelerometer3D.class);
-	
 	
     ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED7");
     ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED6");
@@ -20,18 +21,12 @@ public class AccelerationSampler
 	public double[] getaccelerationArray (int samplePeriodIfListening, 
 										  int samplePeriodIfMeasuring, 
 										  int arrayLength, 
-										  double threshold) 
-	{
-		
+										  double threshold) 	{ 
 		double[] accelerationArray = new double[arrayLength];
-		
-        try {
-        	
+		try {
         	double accely = accelerometer.getAccelY();
-        	
-            // Flash an LED to indicate a sampling event
-            led.setRGB(255, 255, 0);
-            led.setOn();
+        	// Flash an LED to indicate a sampling event
+            led.setRGB(255, 255, 0); led.setOn();
         	
         	// measure acceleration with low samplerate, until acceleration 
             // exceeds threshold
@@ -40,10 +35,8 @@ public class AccelerationSampler
         		accely = accelerometer.getAccelY();
 			}
             led.setOff();
-            
             // light an LED to indicate a sampling event
-            led2.setRGB(0, 255, 255);
-            led2.setOn();
+            led2.setRGB(0, 255, 255); led2.setOn();
         	
         	// measure acceleration with high samplerate, store values in array
 			for (int i = 0; i < arrayLength; i++) {
@@ -52,7 +45,6 @@ public class AccelerationSampler
 				//System.out.println("accY=  " + accelerationArray[i]);
 				Utils.sleep(samplePeriodIfMeasuring - 
 										(System.currentTimeMillis() - start));
-				
 			}
 			
 			// turn off LED
@@ -61,8 +53,6 @@ public class AccelerationSampler
         } catch (Exception e) {
[…]

### 2015-02-19 SHM_JahrSchlich/sensornode/src/sensornode/CommunicationNormal.java

@@ -1,3 +1,8 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornode;
 
 import com.sun.spot.io.j2me.radiogram.*;
@@ -12,7 +17,6 @@ import com.sun.spot.util.Utils;
 import javax.microedition.io.*;
 
 public class CommunicationNormal {
-		
 	    String ourAddress = System.getProperty("IEEE_ADDRESS");
         ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED2");
         ITriColorLED led3 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED3");
@@ -24,21 +28,17 @@ public class CommunicationNormal {
         public void ExchangeData (Measurement ourMeas, String otherAddress, int hostPort) {
 
         	try {     
-        		
         		System.out.println("starting ExchangeData. ");
         		conn = (RadiostreamConnection)Connector.open("radiostream://" + otherAddress + ":" + hostPort);
         		// set timeout for connection
         		conn.setTimeout(5000); 
-        		
+        		// open streams
         		DataInputStream dis = conn.openDataInputStream();
         		DataOutputStream dos = conn.openDataOutputStream();
-                 
         		System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
-
         		boolean okay = false;
         		while(okay == false){
         			try{
-	            		
 	    				dos.writeFloat(ourMeas.frequency);
 	    				dos.writeDouble(ourMeas.magnitude);
 	    				dos.writeInt(ourMeas.error);
@@ -46,36 +46,23 @@ public class CommunicationNormal {
 	    				System.out.println("datastream flushed." );
 	    				
 	    		        // blink second LED
-	    		        led2.setRGB(0, 255, 0);
-	    	            led2.setOn();
+	    		        led2.setRGB(0, 255, 0); led2.setOn();
 	    	            
 	    				okay = dis.readBoolean();
-	    				
 	    				System.out.println("okay received." );
-	    				
-	    		        // blink third LED
-	    		        led3.setRGB(0, 255, 255);
-	    	            led3.setOn();
-	    	          
+	    				 // blink third LED
+	    		        led3.setRGB(0, 255, 255); led3.setOn();
+	    		        
     				} catch(Exception e){
     					continue;
-       			
     				}
         		}
-	            
-        		dis.close();
-        		dos.close();
-        		conn.close();
-        		
[…]

### 2015-02-19 SHM_JahrSchlich/sensornode/src/sensornode/FFT.java

@@ -1,12 +1,11 @@
-package sensornode;
 /**
  * @author Orlando Selenu
- * modified by Jahr&Schlich
+ * additional methods by Jahr&Schlich
  * 
  */
+package sensornode;
 
 import java.lang.Math;
-
 import com.sun.squawk.util.MathUtils;
 
 public class FFT {
@@ -145,7 +144,7 @@ public class FFT {
     }
     
     
-	// calculate magnitude as norm of imaginary and real part 
+	// calculate magnitudes as norm of imaginary and real part 
     public static double[] calcMagnitude(double[] transform)
     {
     	int nummeas = transform.length / 2;
@@ -158,7 +157,7 @@ public class FFT {
 		return magnitude;
     }
     
-    // calculate frequencies from samplerate and sampled values
+    // calculate frequencies from samplerate and number of sampled values
     public static double[] calcFreq(double[] transform, double samplerate)
     {
     	int nummeas = transform.length / 2;
@@ -170,30 +169,26 @@ public class FFT {
 		return frequency;
     }
     
-	//get maximal magnitude -> natural frequency
+	//get maximal magnitude and of natural frequency
     public static Measurement calcNaturalFreq (double[] magnitude, double[] frequency, String ourAddress)
     {
     	// split array in half to get rid of double frequencies
     	double[] newMagnitude = new double[magnitude.length/2];
     	for(int i =0;i < newMagnitude.length;i++)
     		newMagnitude[i] = magnitude[i];
-    	
+    	    	
     	// determine maximal magnitude
 		double maxMagnitude = newMagnitude[0];
 		int indexMax = 0;
 		for ( int i = 1; i < newMagnitude.length; i++) {
-			//System.out.println(newMagnitude[i]);
 		    if ( newMagnitude[i] > maxMagnitude) {
 		    	maxMagnitude = newMagnitude[i];
 		    	indexMax = i;
 		    }
 		}
-		
 		//get frequency for maximal  magnitude
 		float maxFrequency = (float) frequency[indexMax];
-		//System.out.println("maxMagnitude= "+ maxMagnitude+
-		//									", maxFrequency= " + maxFrequency);
-		
+
 		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency,0);
         return measurement;
     }

### 2015-02-19 SHM_JahrSchlich/sensornode/src/sensornode/MainSpot.java

@@ -1,3 +1,8 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornode;
 
 import com.sun.spot.io.j2me.radiogram.*;
@@ -13,7 +18,6 @@ import com.sun.spot.resources.transducers.IAccelerometer3D;
 import com.sun.spot.service.BootloaderListenerService;
 //import java.io.Writer;
 
-
 public class MainSpot extends MIDlet {
 
     private int HOST_PORT = 67;
@@ -24,7 +28,6 @@ public class MainSpot extends MIDlet {
     private double THRESHOLD = 0.15;
     private int NUMBER_OF_OTHER_SENSORS = 1;
     
-    
     private String ourAddress = System.getProperty("IEEE_ADDRESS");
     private String otherAddress = "0014.4F01.0000.763F";
     
@@ -34,11 +37,9 @@ public class MainSpot extends MIDlet {
     	new BootloaderListenerService().getInstance().start();
     	// initalize communication between spots
 		CommunicationNormal communication = new CommunicationNormal();
-		
-        //initialize AccelerationSampler
+	    //initialize AccelerationSampler
         AccelerationSampler sensorSampler = new AccelerationSampler();
-		
-       
+	   
         while (true){
 	        // get acceleration for this sensornode
 	        double[] accelerationArray = sensorSampler.getaccelerationArray
@@ -52,11 +53,11 @@ public class MainSpot extends MIDlet {
 			double[] frequency = FFT.calcFreq(transform, SAMPLERATE);
 			// instanziate und initialize measurement object
 			Measurement ownMeas =  FFT.calcNaturalFreq(magnitude, frequency, ourAddress);
-			System.out.println("magnitude: " + ownMeas.magnitude + " frequency: " + ownMeas.frequency);
+			System.out.println("magnitude: " + ownMeas.magnitude + 
+											" frequency: " + ownMeas.frequency);
 			
 			// exchange measurement with measurement of other sensors
 			communication.ExchangeData(ownMeas, otherAddress, HOST_PORT);
-			
 
 			Utils.sleep(2000);
 		}   

### 2015-02-19 SHM_JahrSchlich/sensornode/src/sensornode/Measurement.java

@@ -1,3 +1,8 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornode;
 public class Measurement{
 	
@@ -8,8 +13,7 @@ public class Measurement{
 	public double prediction = 0;
 	
 	// initiates a measurement
-	Measurement(String a, double m, float f, int e){
-		
+	Measurement(String a, double m, float f, int e){		
 		address = a;
 		magnitude = m;
 		frequency = f;
@@ -17,8 +21,7 @@ public class Measurement{
 	}
 	
 	// initiates a measurement
-	Measurement(){
-		
+	Measurement(){		
 		address = "a";
 		magnitude = 0;
 		frequency = 0;

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/AccelerationSampler.java

@@ -1,18 +1,18 @@
-package sensornodeNN;
+/**
+ * @author Jahr&Schlich
+ * 
+ */
 
+package sensornodeNN;
 import com.sun.spot.resources.Resources;
 import com.sun.spot.resources.transducers.IAccelerometer3D;
 import com.sun.spot.util.Utils;
 import java.lang.Math;
 import com.sun.spot.resources.transducers.ITriColorLED;
 
-
-
-public class AccelerationSampler 
-{
+public class AccelerationSampler {
 	IAccelerometer3D accelerometer = 
 			(IAccelerometer3D) Resources.lookup(IAccelerometer3D.class);
-	
 	
     ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED7");
     ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED6");
@@ -20,18 +20,12 @@ public class AccelerationSampler
 	public double[] getaccelerationArray (int samplePeriodIfListening, 
 										  int samplePeriodIfMeasuring, 
 										  int arrayLength, 
-										  double threshold) 
-	{
-		
+										  double threshold) 	{ 
 		double[] accelerationArray = new double[arrayLength];
-		
-        try {
-        	
+		try {
         	double accely = accelerometer.getAccelY();
-        	
-            // Flash an LED to indicate a sampling event
-            led.setRGB(255, 255, 0);
-            led.setOn();
+        	// Flash an LED to indicate a sampling event
+            led.setRGB(255, 255, 0); led.setOn();
         	
         	// measure acceleration with low samplerate, until acceleration 
             // exceeds threshold
@@ -40,10 +34,8 @@ public class AccelerationSampler
         		accely = accelerometer.getAccelY();
 			}
             led.setOff();
-            
             // light an LED to indicate a sampling event
-            led2.setRGB(0, 255, 255);
-            led2.setOn();
+            led2.setRGB(0, 255, 255); led2.setOn();
         	
         	// measure acceleration with high samplerate, store values in array
 			for (int i = 0; i < arrayLength; i++) {
@@ -52,7 +44,6 @@ public class AccelerationSampler
 				//System.out.println("accY=  " + accelerationArray[i]);
 				Utils.sleep(samplePeriodIfMeasuring - 
 										(System.currentTimeMillis() - start));
-				
 			}
 			
 			// turn off LED
@@ -61,8 +52,6 @@ public class AccelerationSampler
[…]

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/CommunicationNN.java

@@ -1,3 +1,8 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornodeNN;
 
 import com.sun.spot.io.j2me.radiogram.*;
@@ -13,8 +18,7 @@ import com.sun.spot.util.Utils;
 import javax.microedition.io.*;
 
 public class CommunicationNN {
-		
-	    String ourAddress = System.getProperty("IEEE_ADDRESS");
+		String ourAddress = System.getProperty("IEEE_ADDRESS");
         ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED2");
         ITriColorLED led3 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED3");
         ITriColorLED led5 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED5");
@@ -24,96 +28,68 @@ public class CommunicationNN {
         DataInputStream dis;
         DataOutputStream dos; 
         
+        // receive measurements from other sensors
         public Measurement ReceiveData(String otherAddress, int hostPort) {
-            //object for other measurement
+            //instantiate object for received measurement of other sensor
             Measurement othMeas = new Measurement();
         	try {     
-        		
-        		//System.out.println("starting ExchangeData. ");
-        		conn = (RadiostreamConnection)Connector.open("radiostream://" + otherAddress + ":" + hostPort);
-        		// set timeout for connection
-        		// conn.setTimeout(10000); 
+
+        		conn = (RadiostreamConnection)Connector.open("radiostream://" 
+        								+ otherAddress + ":" + hostPort);
         		DataInputStream dis = conn.openDataInputStream();
         		DataOutputStream dos = conn.openDataOutputStream();
                  
-        		//System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
-
-				//System.out.println("try to recieve." );	
 				// receive data from stream
-        		
 	            othMeas.address = otherAddress;
 	            othMeas.frequency = dis.readFloat();
 	    		othMeas.magnitude = dis.readDouble();
 	    		othMeas.error = dis.readInt();
-	    		
-				//System.out.println("datastream received." );
-	    		
-		        // blink third LED
-		        led3.setRGB(0, 255, 255);
-	            led3.setOn();
+
+		        // blink third LED to indicate received measurement
+		        led3.setRGB(0, 255, 255); led3.setOn();
 	            
-	            // send "okay"
+	            // send "okay", so other sensor node stops to try to send
[…]

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/ErrorMeasurement.java

@@ -1,8 +1,5 @@
 package sensornodeNN;
 
-
-
-
 /**
  * 
  * This class contains several methods of error measurement for use in a static

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/FFT.java

@@ -1,9 +1,10 @@
-package sensornodeNN;
 /**
  * @author Orlando Selenu
- * modified by Jahr&Schlich
+ * additional methods by Jahr&Schlich
  * 
  */
+package sensornodeNN;
+
 
 import java.lang.Math;
 
@@ -145,7 +146,7 @@ public class FFT {
     }
     
     
-	// calculate magnitude as norm of imaginary and real part 
+	// calculate magnitudes as norm of imaginary and real part 
     public static double[] calcMagnitude(double[] transform)
     {
     	int nummeas = transform.length / 2;
@@ -158,7 +159,7 @@ public class FFT {
 		return magnitude;
     }
     
-    // calculate frequencies from samplerate and sampled values
+    // calculate frequencies from samplerate and number of sampled values
     public static double[] calcFreq(double[] transform, double samplerate)
     {
     	int nummeas = transform.length / 2;
@@ -170,14 +171,14 @@ public class FFT {
 		return frequency;
     }
     
-	//get maximal magnitude -> natural frequency
+	//get maximal magnitude and of natural frequency
     public static Measurement calcNaturalFreq (double[] magnitude, double[] frequency, String ourAddress)
     {
     	// split array in half to get rid of double frequencies
     	double[] newMagnitude = new double[magnitude.length/2];
     	for(int i =0;i < newMagnitude.length;i++)
     		newMagnitude[i] = magnitude[i];
-    	
+    	    	
     	// determine maximal magnitude
 		double maxMagnitude = newMagnitude[0];
 		int indexMax = 0;
@@ -187,12 +188,9 @@ public class FFT {
 		    	indexMax = i;
 		    }
 		}
-		
 		//get frequency for maximal  magnitude
 		float maxFrequency = (float) frequency[indexMax];
-		//System.out.println("maxMagnitude= "+ maxMagnitude+
-		//									", maxFrequency= " + maxFrequency);
-		
+
 		Measurement measurement = new Measurement(ourAddress, maxMagnitude, maxFrequency,0);
         return measurement;
     }

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/MainSpot.java

@@ -1,3 +1,8 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornodeNN;
 
 import com.sun.spot.io.j2me.radiogram.*;
@@ -12,68 +17,63 @@ import javax.microedition.midlet.MIDletStateChangeException;
 import com.sun.spot.resources.transducers.IAccelerometer3D;
 import com.sun.spot.service.BootloaderListenerService;
 
-
-
 public class MainSpot extends MIDlet {
 	
-	// measurement
+	// constants for measurement
     private int SAMPLE_PERIOD_LISTENING = 1 * 100;  // in milliseconds
     private int SAMPLE_PERIOD_MEASURING = 20;  // in milliseconds
     private int ARRAY_LENGTH = 256;
     private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
     private double THRESHOLD = 0.2;
 
-    // communication
+    // constants for communication
     private String BASE_NAME = "0014.4F01.0000.77BA";
     private String[] SENSOR_NAMES = {"0014.4F01.0000.792D",
-    								"0014.4F01.0000.7840"};
-//    								"0014.4F01.0000.7AFA"}
+    								"0014.4F01.0000.7840",
+    								"0014.4F01.0000.7AFA"}
     private String ourAddress = System.getProperty("IEEE_ADDRESS");
     private int HOST_PORT = 67;
     private int HOST_PORT_BASE = 68;
     
-    
-    // neural network
+    //  constants for neural network
     private int HIDDEN_UNITS = 3;
-    
     private int trainingEvents = 6;
 
     protected void startApp() throws MIDletStateChangeException {
-
+		ITriColorLED led4 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED4");
+		int numberOfSensors = SENSOR_NAMES.length;	
+		
         // Listen for downloads/commands over USB connection
     	new BootloaderListenerService().getInstance().start();
     	// initalize communication between spots
 		CommunicationNN communication = new CommunicationNN();
-		
-		//Radiogram for communication to basestation
-		//communication.EstablishConnection(HOST_PORT);
-		
-        //initialize AccelerationSampler
+		//initialize AccelerationSampler
         AccelerationSampler sensorSampler = new AccelerationSampler();
-		
[…]

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/Measurement.java

@@ -1,3 +1,9 @@
+/**
+ * 
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornodeNN;
 
 public class Measurement{

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/NetworkTraining.java

@@ -1,47 +1,38 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
 package sensornodeNN;
 
 public class NetworkTraining {
 	  
-
 	public void Train(NeuralNetwork net, TrainingSampleLesson lesson) {
 
-		 
 		int runs = 100000; // number of random patterns to train
 		double eta = 0.5; //learning rate
 		double threshold = 1e-5;
 	 
-		
 		// Train neural network until error falls below limit
 		double error;
 		double newerror;
 		int numOfPhases = 0;
-	
 		do{
 			numOfPhases++;
-			
 			error = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
-			
-			System.out.println("Error 1: "+ error);	
-			
+			System.out.println("Error 1: "+ error);
+			// lower training rate with every cycle
 			eta = eta/2;
-			
+			//train the neural network
 			net.trainBackpropagationOfError(lesson, runs, eta);
-			
-			//net.trainResilientBackpropagation(lesson, runs, false);
-			
-			
-			
+			//calculate new error
 			newerror = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
-			
 			System.out.println("Error after training phase "+ numOfPhases + " = " + newerror );
-			//System.out.println("Abbruchbedingung "+ Math.abs(newerror-error)*100 / newerror);
-			
+			// try to lower error 10 times
 			if (numOfPhases > 10) break;
 			
 		}while(newerror > threshold);
 		
-		System.out.println(numOfPhases + " training phases. ");
-
+		System.out.println(numOfPhases + " trainings. ");
 		}
-
 	}

### 2015-02-19 SHM_JahrSchlich/sensornodeNN/src/sensornodeNN/TrainingSampleLesson.java

@@ -1,9 +1,7 @@
 package sensornodeNN;
 
-
 //import java.io.Serializable;
 import java.util.Random;
-
 
 
 /**

### 2015-02-19 Code WA/sensornode/src/sensornode/AccelerationSampler.java
Code Basestation aufgerumt, ordner Code WA neu erstellt für den reinen code

@@ -1,0 +1,58 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
+package sensornode;
+
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.IAccelerometer3D;
+import com.sun.spot.util.Utils;
+import java.lang.Math;
+import com.sun.spot.resources.transducers.ITriColorLED;
+
+public class AccelerationSampler {
+	IAccelerometer3D accelerometer = 
+			(IAccelerometer3D) Resources.lookup(IAccelerometer3D.class);
+	
+    ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED7");
+    ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED6");
+	
+	public double[] getaccelerationArray (int samplePeriodIfListening, 
+										  int samplePeriodIfMeasuring, 
+										  int arrayLength, 
+										  double threshold) 	{ 
+		double[] accelerationArray = new double[arrayLength];
+		try {
+        	double accely = accelerometer.getAccelY();
+        	// Flash an LED to indicate a sampling event
+            led.setRGB(255, 255, 0); led.setOn();
+        	
+        	// measure acceleration with low samplerate, until acceleration 
+            // exceeds threshold
+        	while (Math.abs(accely) < threshold) {
+        		Utils.sleep(samplePeriodIfListening);
+        		accely = accelerometer.getAccelY();
+			}
+            led.setOff();
+            // light an LED to indicate a sampling event
+            led2.setRGB(0, 255, 255); led2.setOn();
+        	
+        	// measure acceleration with high samplerate, store values in array
+			for (int i = 0; i < arrayLength; i++) {
+				long start = System.currentTimeMillis();
+				accelerationArray[i] = accelerometer.getAccelY();
+				//System.out.println("accY=  " + accelerationArray[i]);
+				Utils.sleep(samplePeriodIfMeasuring - 
+										(System.currentTimeMillis() - start));
+			}
+			
+			// turn off LED
+            led2.setOff();
+			
+        } catch (Exception e) {
+            System.err.println("Caught " + e + " during sensor sampling.");
+        }
+        return accelerationArray;
+	}
+}

### 2015-02-19 Code WA/sensornode/src/sensornode/CommunicationNormal.java

@@ -1,0 +1,68 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
+package sensornode;
+
+import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.io.j2me.radiostream.*;
+import java.io.DataInputStream;
+import java.io.DataOutputStream;
+
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.ITriColorLED;
+import com.sun.spot.util.Utils;
+
+import javax.microedition.io.*;
+
+public class CommunicationNormal {
+	    String ourAddress = System.getProperty("IEEE_ADDRESS");
+        ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED2");
+        ITriColorLED led3 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED3");
+        
+        RadiostreamConnection conn;
+        DataInputStream dis;
+        DataOutputStream dos; 
+        
+        public void ExchangeData (Measurement ourMeas, String otherAddress, int hostPort) {
+
+        	try {     
+        		System.out.println("starting ExchangeData. ");
+        		conn = (RadiostreamConnection)Connector.open("radiostream://" + otherAddress + ":" + hostPort);
+        		// set timeout for connection
+        		conn.setTimeout(5000); 
+        		// open streams
+        		DataInputStream dis = conn.openDataInputStream();
+        		DataOutputStream dos = conn.openDataOutputStream();
+        		System.out.println("Starting communication on" + ourAddress + " with " + otherAddress);
+        		boolean okay = false;
+        		while(okay == false){
+        			try{
+	    				dos.writeFloat(ourMeas.frequency);
+	    				dos.writeDouble(ourMeas.magnitude);
+	    				dos.writeInt(ourMeas.error);
+	    				dos.flush();
+	    				System.out.println("datastream flushed." );
+	    				
+	    		        // blink second LED
+	    		        led2.setRGB(0, 255, 0); led2.setOn();
+	    	            
+	    				okay = dis.readBoolean();
+	    				System.out.println("okay received." );
+	    				 // blink third LED
+	    		        led3.setRGB(0, 255, 255); led3.setOn();
+	    		        
+    				} catch(Exception e){
+    					continue;
+    				}
+        		}
+        		//close radiostreams
+        		dis.close(); dos.close(); conn.close();
+        		led2.setOff(); led3.setOff();
+        		
[…]

### 2015-02-19 Code WA/sensornode/src/sensornode/FFT.java

@@ -1,0 +1,195 @@
+/**
+ * @author Orlando Selenu
+ * additional methods by Jahr&Schlich
+ * 
+ */
+package sensornode;
+
+import java.lang.Math;
+import com.sun.squawk.util.MathUtils;
+
+public class FFT {
+    /**
+     * The Fast Fourier Transform (generic version, with NO optimizations).
+     * 
+     * @param inputReal
+     *            an array of length n, the real part
+     * @param inputImag
+     *            an array of length n, the imaginary part
+     * @param DIRECT
+     *            TRUE = direct transform, FALSE = inverse transform
+     * @return a new array of length 2n
+     */
+    public static double[] performFFT(final double[] inputReal, double[] inputImag,
+	    boolean DIRECT) 
+    {
+		// - n is the dimension of the problem
+		// - nu is its logarithm in base e
+		int n = inputReal.length;
+	 
+		// If n is a power of 2, then ld is an integer (_without_ decimals)
+		double ld = MathUtils.log(n) / MathUtils.log(2.0);
+	 
+		// Here I check if n is a power of 2. If exist decimals in ld, I quit
+		// from the function returning null.
+		if (((int) ld) - ld != 0) 
+		{
+		    System.out.println("The number of elements is not a power of 2.");
+		    return null;
+		}
+	 
+		// Declaration and initialization of the variables
+		// ld should be an integer, actually, so I don't lose any information in
+		// the cast
+		int nu = (int) ld;
+		int n2 = n / 2;
+		int nu1 = nu - 1;
+		double[] xReal = new double[n];
+		double[] xImag = new double[n];
+		double tReal, tImag, p, arg, c, s;
+			
+		// Here I check if I'm going to do the direct transform or the inverse
+		// transform.
+		double constant;
+		if (DIRECT)
+		    constant = -2 * Math.PI;
+		else
+		    constant = 2 * Math.PI;
+	 
+		// I don't want to overwrite the input arrays, so here I copy them. This
+		// choice adds \Theta(2n) to the complexity.
+		for (int i = 0; i < n; i++) 
+		{
+		    xReal[i] = inputReal[i];
+		    xImag[i] = inputImag[i];
+		}
+		 
+		// First phase - calculation
+		int k = 0;
+		for (int l = 1; l <= nu; l++) 
+		{
[…]

### 2015-02-19 Code WA/sensornode/src/sensornode/MainSpot.java

@@ -1,0 +1,74 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
+package sensornode;
+
+import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.ITriColorLED;
+import com.sun.spot.util.Utils;
+
+import javax.microedition.io.*;
+import javax.microedition.midlet.MIDlet;
+import javax.microedition.midlet.MIDletStateChangeException;
+
+import com.sun.spot.resources.transducers.IAccelerometer3D;
+import com.sun.spot.service.BootloaderListenerService;
+//import java.io.Writer;
+
+public class MainSpot extends MIDlet {
+
+    private int HOST_PORT = 67;
+    private int SAMPLE_PERIOD_LISTENING = 1 * 500;  // in milliseconds
+    private int SAMPLE_PERIOD_MEASURING = 20;  // in milliseconds
+    private int ARRAY_LENGTH = 256;
+    private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
+    private double THRESHOLD = 0.15;
+    private int NUMBER_OF_OTHER_SENSORS = 1;
+    
+    private String ourAddress = System.getProperty("IEEE_ADDRESS");
+    private String otherAddress = "0014.4F01.0000.763F";
+    
+    protected void startApp() throws MIDletStateChangeException {
+
+        // Listen for downloads/commands over USB connection
+    	new BootloaderListenerService().getInstance().start();
+    	// initalize communication between spots
+		CommunicationNormal communication = new CommunicationNormal();
+	    //initialize AccelerationSampler
+        AccelerationSampler sensorSampler = new AccelerationSampler();
+	   
+        while (true){
+	        // get acceleration for this sensornode
+	        double[] accelerationArray = sensorSampler.getaccelerationArray
+					        (SAMPLE_PERIOD_LISTENING, SAMPLE_PERIOD_MEASURING, 
+					         ARRAY_LENGTH, THRESHOLD);	        
+	        // perform fft
+			double[] transform = FFT.performFFT(accelerationArray, 
+									new double[accelerationArray.length], true);
+			// calculate magnitude, frequencies and natural frequency
+			double[] magnitude = FFT.calcMagnitude(transform);
+			double[] frequency = FFT.calcFreq(transform, SAMPLERATE);
+			// instanziate und initialize measurement object
[…]

### 2015-02-19 Code WA/sensornode/src/sensornode/Measurement.java

@@ -1,9 +1,9 @@
 /**
  * @author Jahr&Schlich
- *
+ * 
  */
 
-package basestation;
+package sensornode;
 public class Measurement{
 	
 	public String address;
@@ -13,7 +13,7 @@ public class Measurement{
 	public double prediction = 0;
 	
 	// initiates a measurement
-	Measurement(String a, double m, float f, int e){
+	Measurement(String a, double m, float f, int e){		
 		address = a;
 		magnitude = m;
 		frequency = f;
@@ -21,7 +21,7 @@ public class Measurement{
 	}
 	
 	// initiates a measurement
-	Measurement(){
+	Measurement(){		
 		address = "a";
 		magnitude = 0;
 		frequency = 0;

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/AccelerationSampler.java

@@ -1,0 +1,57 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
+package sensornodeNN;
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.IAccelerometer3D;
+import com.sun.spot.util.Utils;
+import java.lang.Math;
+import com.sun.spot.resources.transducers.ITriColorLED;
+
+public class AccelerationSampler {
+	IAccelerometer3D accelerometer = 
+			(IAccelerometer3D) Resources.lookup(IAccelerometer3D.class);
+	
+    ITriColorLED led = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED7");
+    ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED6");
+	
+	public double[] getaccelerationArray (int samplePeriodIfListening, 
+										  int samplePeriodIfMeasuring, 
+										  int arrayLength, 
+										  double threshold) 	{ 
+		double[] accelerationArray = new double[arrayLength];
+		try {
+        	double accely = accelerometer.getAccelY();
+        	// Flash an LED to indicate a sampling event
+            led.setRGB(255, 255, 0); led.setOn();
+        	
+        	// measure acceleration with low samplerate, until acceleration 
+            // exceeds threshold
+        	while (Math.abs(accely) < threshold) {
+        		Utils.sleep(samplePeriodIfListening);
+        		accely = accelerometer.getAccelY();
+			}
+            led.setOff();
+            // light an LED to indicate a sampling event
+            led2.setRGB(0, 255, 255); led2.setOn();
+        	
+        	// measure acceleration with high samplerate, store values in array
+			for (int i = 0; i < arrayLength; i++) {
+				long start = System.currentTimeMillis();
+				accelerationArray[i] = accelerometer.getAccelY();
+				//System.out.println("accY=  " + accelerationArray[i]);
+				Utils.sleep(samplePeriodIfMeasuring - 
+										(System.currentTimeMillis() - start));
+			}
+			
+			// turn off LED
+            led2.setOff();
+			
+        } catch (Exception e) {
+            System.err.println("Caught " + e + " during sensor sampling.");
+        }
+        return accelerationArray;
+	}
+}

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/CommunicationNN.java

@@ -1,0 +1,118 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
+package sensornodeNN;
+
+import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.io.j2me.radiostream.*;
+
+import java.io.DataInputStream;
+import java.io.DataOutputStream;
+
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.ITriColorLED;
+import com.sun.spot.util.Utils;
+
+import javax.microedition.io.*;
+
+public class CommunicationNN {
+		String ourAddress = System.getProperty("IEEE_ADDRESS");
+        ITriColorLED led2 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED2");
+        ITriColorLED led3 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED3");
+        ITriColorLED led5 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED5");
+        ITriColorLED led6 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED6");
+        
+        RadiostreamConnection conn;
+        DataInputStream dis;
+        DataOutputStream dos; 
+        
+        // receive measurements from other sensors
+        public Measurement ReceiveData(String otherAddress, int hostPort) {
+            //instantiate object for received measurement of other sensor
+            Measurement othMeas = new Measurement();
+        	try {     
+
+        		conn = (RadiostreamConnection)Connector.open("radiostream://" 
+        								+ otherAddress + ":" + hostPort);
+        		DataInputStream dis = conn.openDataInputStream();
+        		DataOutputStream dos = conn.openDataOutputStream();
+                 
+				// receive data from stream
+	            othMeas.address = otherAddress;
+	            othMeas.frequency = dis.readFloat();
+	    		othMeas.magnitude = dis.readDouble();
+	    		othMeas.error = dis.readInt();
+
+		        // blink third LED to indicate received measurement
+		        led3.setRGB(0, 255, 255); led3.setOn();
+	            
+	            // send "okay", so other sensor node stops to try to send
+        		dos.writeBoolean(true);
+				dos.flush();
+
+				// blink second LED
+		        led2.setRGB(0, 255, 0); led2.setOn();
+	            
+	            // close streams and connection
+        		dis.close(); dos.close(); conn.close();
+        		
+        		// turn off the light
+        		led2.setOff(); led3.setOff();
+	    		        		
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/ErrorMeasurement.java

@@ -1,0 +1,207 @@
+package sensornodeNN;
+
+/**
+ * 
+ * This class contains several methods of error measurement for use in a static
+ * way.
+ * 
+ * @author David Kriesel / dkriesel.com
+ * 
+ */
+public class ErrorMeasurement {
+
+	/**
+	 * Calculates the sum of absolute errors of the given network over all
+	 * samples in the training lesson.
+	 * 
+	 * @param net
+	 * @param lesson
+	 * 
+	 * @return calculated error
+	 */
+	public static double getErrorAbsoluteSum(NeuralNetwork net,
+			TrainingSampleLesson lesson) {
+
+		double[][] inputs = lesson.getInputs();
+		double[][] desiredOutputs = lesson.getDesiredOutputs();
+
+		double error = 0.0;
+
+		for (int i = 0; i < inputs.length; i++) {
+
+			double[] outputs = net.propagate(inputs[i]);
+
+			for (int j = 0; j < outputs.length; j++) {
+				double temp = 0.0;
+				temp = desiredOutputs[i][j] - outputs[j];
+				temp = Math.abs(temp);
+				error += temp;
+			}
+		}
+		return error;
+
+	}
+
+	/**
+	 * Calculates the sum of euclidean errors of the given network over all
+	 * samples in the training lesson.
+	 * 
+	 * @param net
+	 * @param lesson
+	 * 
+	 * @return calculated error
+	 */
+	public static double getErrorEuclideanSum(NeuralNetwork net,
+			TrainingSampleLesson lesson) {
+
+		double[][] inputs = lesson.getInputs();
+		double[][] desiredOutputs = lesson.getDesiredOutputs();
+
+		double error = 0.0;
+
+		for (int i = 0; i < inputs.length; i++) {
+
+			double[] outputs = net.propagate(inputs[i]);
+			
+			System.out.println("outputs: " + outputs[0]);
+
+			for (int j = 0; j < outputs.length; j++) {
+				double temp = 0.0;
+				
+				temp = desiredOutputs[i][j] - outputs[j];
+
+				System.out.println( i + " temp: " +temp);
+				
+				
+				temp = Math.abs(temp);
+				
+				//temp = temp * temp;
+				//temp = Math.sqrt(temp);
+				
+				error += temp;
+			}
+		}
+		return error;
+	}
+
+	/**
+	 * Calculates the Squared percentage error of the given network over all
+	 * samples in the training lesson. This method is proposed by Prechelt
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/FFT.java

@@ -1,0 +1,197 @@
+/**
+ * @author Orlando Selenu
+ * additional methods by Jahr&Schlich
+ * 
+ */
+package sensornodeNN;
+
+
+import java.lang.Math;
+
+import com.sun.squawk.util.MathUtils;
+
+public class FFT {
+    /**
+     * The Fast Fourier Transform (generic version, with NO optimizations).
+     * 
+     * @param inputReal
+     *            an array of length n, the real part
+     * @param inputImag
+     *            an array of length n, the imaginary part
+     * @param DIRECT
+     *            TRUE = direct transform, FALSE = inverse transform
+     * @return a new array of length 2n
+     */
+    public static double[] performFFT(final double[] inputReal, double[] inputImag,
+	    boolean DIRECT) 
+    {
+		// - n is the dimension of the problem
+		// - nu is its logarithm in base e
+		int n = inputReal.length;
+	 
+		// If n is a power of 2, then ld is an integer (_without_ decimals)
+		double ld = MathUtils.log(n) / MathUtils.log(2.0);
+	 
+		// Here I check if n is a power of 2. If exist decimals in ld, I quit
+		// from the function returning null.
+		if (((int) ld) - ld != 0) 
+		{
+		    System.out.println("The number of elements is not a power of 2.");
+		    return null;
+		}
+	 
+		// Declaration and initialization of the variables
+		// ld should be an integer, actually, so I don't lose any information in
+		// the cast
+		int nu = (int) ld;
+		int n2 = n / 2;
+		int nu1 = nu - 1;
+		double[] xReal = new double[n];
+		double[] xImag = new double[n];
+		double tReal, tImag, p, arg, c, s;
+			
+		// Here I check if I'm going to do the direct transform or the inverse
+		// transform.
+		double constant;
+		if (DIRECT)
+		    constant = -2 * Math.PI;
+		else
+		    constant = 2 * Math.PI;
+	 
+		// I don't want to overwrite the input arrays, so here I copy them. This
+		// choice adds \Theta(2n) to the complexity.
+		for (int i = 0; i < n; i++) 
+		{
+		    xReal[i] = inputReal[i];
+		    xImag[i] = inputImag[i];
+		}
+		 
+		// First phase - calculation
+		int k = 0;
+		for (int l = 1; l <= nu; l++) 
+		{
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/MainSpot.java

@@ -1,0 +1,184 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
+package sensornodeNN;
+
+import com.sun.spot.io.j2me.radiogram.*;
+import com.sun.spot.resources.Resources;
+import com.sun.spot.resources.transducers.ITriColorLED;
+import com.sun.spot.util.Utils;
+
+import javax.microedition.io.*;
+import javax.microedition.midlet.MIDlet;
+import javax.microedition.midlet.MIDletStateChangeException;
+
+import com.sun.spot.resources.transducers.IAccelerometer3D;
+import com.sun.spot.service.BootloaderListenerService;
+
+public class MainSpot extends MIDlet {
+	
+	// constants for measurement
+    private int SAMPLE_PERIOD_LISTENING = 1 * 100;  // in milliseconds
+    private int SAMPLE_PERIOD_MEASURING = 20;  // in milliseconds
+    private int ARRAY_LENGTH = 256;
+    private double SAMPLERATE = 1000. / (double)SAMPLE_PERIOD_MEASURING;
+    private double THRESHOLD = 0.2;
+
+    // constants for communication
+    private String BASE_NAME = "0014.4F01.0000.77BA";
+    private String[] SENSOR_NAMES = {"0014.4F01.0000.792D",
+    								"0014.4F01.0000.7840",
+    								"0014.4F01.0000.7AFA"}
+    private String ourAddress = System.getProperty("IEEE_ADDRESS");
+    private int HOST_PORT = 67;
+    private int HOST_PORT_BASE = 68;
+    
+    //  constants for neural network
+    private int HIDDEN_UNITS = 3;
+    private int trainingEvents = 6;
+
+    protected void startApp() throws MIDletStateChangeException {
+		ITriColorLED led4 = (ITriColorLED)Resources.lookup(ITriColorLED.class, "LED4");
+		int numberOfSensors = SENSOR_NAMES.length;	
+		
+        // Listen for downloads/commands over USB connection
+    	new BootloaderListenerService().getInstance().start();
+    	// initalize communication between spots
+		CommunicationNN communication = new CommunicationNN();
+		//initialize AccelerationSampler
+        AccelerationSampler sensorSampler = new AccelerationSampler();
+        
+		// NEURAL NETWORK CREATION
+		// create descriptor
+		NeuralNetworkDescriptor descriptorNN = 
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/Measurement.java

@@ -1,9 +1,11 @@
 /**
+ * 
  * @author Jahr&Schlich
- *
+ * 
  */
 
-package basestation;
+package sensornodeNN;
+
 public class Measurement{
 	
 	public String address;
@@ -14,6 +16,7 @@ public class Measurement{
 	
 	// initiates a measurement
 	Measurement(String a, double m, float f, int e){
+		
 		address = a;
 		magnitude = m;
 		frequency = f;
@@ -22,6 +25,7 @@ public class Measurement{
 	
 	// initiates a measurement
 	Measurement(){
+		
 		address = "a";
 		magnitude = 0;
 		frequency = 0;

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/MersenneTwisterFast.java

@@ -1,0 +1,1344 @@
+package sensornodeNN;
+
+
+
+import java.io.DataInputStream;
+import java.io.DataOutputStream;
+import java.io.IOException;
+import java.util.Random;
+
+import com.sun.squawk.util.MathUtils;
+
+
+
+/**
+ * <h3>MersenneTwister and MersenneTwisterFast</h3>
+ * <p>
+ * <b>Version 16</b>, based on version MT199937(99/10/29) of the Mersenne
+ * Twister algorithm found at <a
+ * href="http://www.math.keio.ac.jp/matumoto/emt.html"> The Mersenne Twister
+ * Home Page</a>, with the initialization improved using the new 2002/1/26
+ * initialization algorithm By Sean Luke, October 2004.
+ * 
+ * <p>
+ * <b>MersenneTwister</b> is a drop-in subclass replacement for
+ * java.util.Random. It is properly synchronized and can be used in a
+ * multithreaded environment. On modern VMs such as HotSpot, it is approximately
+ * 1/3 slower than java.util.Random.
+ * 
+ * <p>
+ * <b>MersenneTwisterFast</b> is not a subclass of java.util.Random. It has the
+ * same public methods as Random does, however, and it is algorithmically
+ * identical to MersenneTwister. MersenneTwisterFast has hard-code inlined all
+ * of its methods directly, and made all of them final (well, the ones of
+ * consequence anyway). Further, these methods are <i>not</i> synchronized, so
+ * the same MersenneTwisterFast instance cannot be shared by multiple threads.
+ * But all this helps MersenneTwisterFast achieve well over twice the speed of
+ * MersenneTwister. java.util.Random is about 1/3 slower than
+ * MersenneTwisterFast.
+ * 
+ * <h3>About the Mersenne Twister</h3>
+ * <p>
+ * This is a Java version of the C-program for MT19937: Integer version. The
+ * MT19937 algorithm was created by Makoto Matsumoto and Takuji Nishimura, who
+ * ask: "When you use this, send an email to: matumoto@math.keio.ac.jp with an
+ * appropriate reference to your work". Indicate that this is a translation of
+ * their algorithm into Java.
+ * 
+ * <p>
+ * <b>Reference. </b> Makato Matsumoto and Takuji Nishimura, "Mersenne Twister:
+ * A 623-Dimensionally Equidistributed Uniform Pseudo-Random Number Generator",
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/NetworkTraining.java

@@ -1,0 +1,38 @@
+/**
+ * @author Jahr&Schlich
+ * 
+ */
+
+package sensornodeNN;
+
+public class NetworkTraining {
+	  
+	public void Train(NeuralNetwork net, TrainingSampleLesson lesson) {
+
+		int runs = 100000; // number of random patterns to train
+		double eta = 0.5; //learning rate
+		double threshold = 1e-5;
+	 
+		// Train neural network until error falls below limit
+		double error;
+		double newerror;
+		int numOfPhases = 0;
+		do{
+			numOfPhases++;
+			error = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
+			System.out.println("Error 1: "+ error);
+			// lower training rate with every cycle
+			eta = eta/2;
+			//train the neural network
+			net.trainBackpropagationOfError(lesson, runs, eta);
+			//calculate new error
+			newerror = ErrorMeasurement.getErrorRootMeanSquareSum(net, lesson);
+			System.out.println("Error after training phase "+ numOfPhases + " = " + newerror );
+			// try to lower error 10 times
+			if (numOfPhases > 10) break;
+			
+		}while(newerror > threshold);
+		
+		System.out.println(numOfPhases + " trainings. ");
+		}
+	}

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/NeuralNetwork.java

@@ -1,0 +1,3920 @@
+package sensornodeNN;
+
+//import java.io.Serializable;
+import java.util.Hashtable;
+import java.util.Random;
+
+import com.sun.squawk.util.Arrays;
+import com.sun.squawk.util.MathUtils;
+
+/**
+ * <b>Instantiated second using a NeuralNetworkDescriptor instance,</b> this
+ * class represents the core functionality of SNIPE: A neural network of
+ * arbitrary topology with a lightweight and calculation efficient data
+ * structure, the possibility to change the entire network topology as well as
+ * the synaptic weights and many other features; In the documentation of this
+ * class, detail information about the efficient data structure of the neural
+ * network is given, as well as information about its features. Goals of the
+ * design include:
+ * 
+ * <ol>
+ * <li><b>Generalized data structure for arbitrary network topologies</b>, so
+ * that virtually all network structures can be realized or even easily
+ * hand-crafted.
+ * <li><b>Built-In, fast and easy-to-use learning operators</b> for gradient
+ * descent or evolutionary learning, as well as mechanisms for efficent control
+ * of even large network populations.
+ * <li><b>Mechanisms for design and control of even large populations of neural
+ * networks</b>
+ * <li><b>Optimal speed neural network data propagation</b> in contrast to naive
+ * data structures, even in special cases like multi layer perceptrons and
+ * sparse networks, as well as low computational topology editing effort.
+ * <li><b>Low memory consumption</b> - grows only with the number of existing
+ * synapses, not quadratically with the number of neurons
+ * <li><b>In-situ processing</b> - no extra memory or preprocessing of the data
+ * structure is necessary in order to use the network after editing
+ * <li><b>Usage of only low level data structures (arrays)</b> for easy
+ * portability. It is not the goal to quench the last tiniest bit of
+ * asymptotical complexity out of the structure, but to make it usable, light
+ * weight and fast in praxis.
+ * <li><b>No object-oriented overload</b>, like objects for every neuron or even
+ * synapses, etc.
+ * </ol>
+ * 
+ * 
+ * 
+ * <h1>Fundamentals</h1>
+ * 
+ * 
+ * 
+ * <p>
+ * The Neuron enumeration structure is defined as follows: Neuron 0 is the bias
+ * neuron, at least neuron 1 is an input neuron which may be followed by other
+ * input neurons, depending on the input dimensionality. The most important
+ * information considering the neuron index set is, that it is partitioned in
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/NeuralNetworkDescriptor.java

@@ -1,0 +1,677 @@
+package sensornodeNN;
+
+//import java.io.Serializable;
+
+
+
+/**
+ * <b>Read and instantiate first</b> - A NeuralNetworkDescriptor object defines
+ * a set of high-level general parameters that are used to create a possibly
+ * large group of neural network instances; so first, you have to create a
+ * NeuralNetworkDescriptor, then use it to create as many neural network
+ * instances as you wish; moreover, some fundamental information about SNIPE
+ * usage is given in the documentation of this class.
+ * 
+ * <h1>Fundamentals and Usage</h1>
+ * 
+ * To create a neural network, you first define a descriptor, and then create a
+ * neural network using the descriptor. This layout is useful for the creation
+ * of even large populations of networks, for instance when using population
+ * based metaheuristics.
+ * 
+ * I suggest you to
+ * <ol>
+ * <li>define what layers your network(s) shall contain creating an int array as
+ * described in the constructor documentation (or use one of the quick
+ * constructor wrappers),
+ * <li>then create a descriptor instance using the layer array,
+ * <li>then use one of the quick functions starting with "setSettings" if you
+ * want to use common network topologies,
+ * <li>tweak some further descriptor settings (for example, neuron behaviors
+ * e.g. activity functions) and
+ * <li>create as many networks as you wish out of the descriptor (if you wish,
+ * use one of the "create" methods of the descriptor itself).
+ * </ol>
+ * Once networks are created, settings in the descriptor can still be changed
+ * via the setters (useful if you for example want to allow or deny synapse
+ * types later on for all networks created using one descriptor). However, the
+ * neurons per layer are directly defined using the constructor and do not have
+ * a setter. They can't be changed.
+ * 
+ * <h1>Descriptors Outline Networks, they do not define them</h1>
+ * 
+ * Descriptors <i>outline</i> networks, but don't define the exact topology or
+ * even the number of inner neurons or last, the synaptic weight values. The
+ * neuron per layer array is a good example for this. With this array, you can
+ * define neuron numbers per layer to be static or to be randomized every time a
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/NeuronBehavior.java

@@ -1,0 +1,74 @@
+package sensornodeNN;
+
+
+
+/**
+ * 
+ * Defines the interface for a neuron behavior, which denotes the generalization
+ * of activity functions.
+ * 
+ * <p>There are two kinds of assigning behaviors to neurons: Neurons can just be
+ * assigned a reference to a behavior, which saves memory and is okay for standard
+ * state-less activation functions. However, there may be other desired neuron
+ * behaviors that require neuron-dedicated internal data. In this case, every
+ * neuron needs a separate, dedicated instance of the neuron behavior class.
+ * Please implement the needsDedicatedInstancePerNeuron method in order to
+ * define your kind of neuron behavior when you inherit from this interface.
+ * 
+ * @author David Kriesel / dkriesel.com
+ * 
+ */
+public interface NeuronBehavior{
+	/**
+	 * @param x
+	 *            usually the net input
+	 * @return the activation value of the neuron behavior given x.
+	 */
+	public double computeActivation(double x);
+
+	/**
+	 * @param x
+	 *            usually the net input
+	 * @return the first derivative of the activation function at x or
+	 *         Double.NaN if not applicable. Note that if you return Double.NaN,
+	 *         gradient training will be faulty (return NaNs, too, possibly
+	 *         destroying network functionality).
+	 */
+	public double computeDerivative(double x);
+
+	
+	/**
+	 * @return a separate instance of this class, initialized with the same
+	 *         parameters as this one.
+	 */
+	public NeuronBehavior getDedicatedInstance();
+
+	/**
+	 * Returns the absolute location of the absolute maxima of the second
+	 * derivative. If for instance 1 is returned here, the maxima of the second
+	 * derivative are located at +1 and -1. Only positive values should be
+	 * returned. It is important to know the approximate location of those
+	 * maxima because they define the range of desired outputs a network learns
+	 * best with (see "Efficient Backprop", LeCun). Return Double.NaN if not
+	 * applicable. This function is also used as parameter for randomization of
+	 * weight initialization, however, if you return Double.NaN, it is supposed
+	 * to be 1.0.
+	 * 
+	 * @return the absolute location of the maxima of the second derivative or
+	 *         Double.NaN if not applicable.
+	 */
+	public double getAbsoluteMaximumLocationOfSecondDerivative();
+
+	/**
+	 * This method tells the neural net initialization mechanisms whether or not
+	 * every neuron assigned this behavior should get a dedicated instance of
[…]

### 2015-02-19 Code WA/sensornodeNN/src/sensornodeNN/TrainingSampleLesson.java

@@ -1,0 +1,532 @@
+package sensornodeNN;
+
+//import java.io.Serializable;
+import java.util.Random;
+
+
+/**
+ * This class represents a set of training samples in order to train a neural
+ * network. Once created, an instance of this class checks whether input or
+ * output are null, of inequal internal length, not given, etc. Inputs and
+ * desired output samples are given as 2D-double-arrays
+ * <ul>
+ * <li>inputs[n][i] and
+ * <li>desiredOutputs[n][o],
+ * </ul>
+ * where n is the number of samples, i is the input dimensionality and o is the
+ * output dimensionality.
+ * 
+ * <p>
+ * It is recommended to both optimize inputs and desired outputs using the
+ * methods starting with "optimize" (read their documentation).
+ * 
+ * <p>
+ * If teaching classification problems you should consider placing the binary
+ * values at -x and x, where x is the absolute location of the absolute maxima
+ * of the 2nd derivative of the activation function (LeCun - Efficient
+ * Backprop). Use the scaleDesiredOutputsForClassificationProblem method in
+ * order to do this.
+ * 
+ * @author David Kriesel / dkriesel.com
+ */
+public class TrainingSampleLesson{// implements Serializable {
+
+	private static final long serialVersionUID = 1L;
+	private double[][] inputs;
+	private double[][] desiredOutputs;
+
+	/**
+	 * Creates a TrainingSampleLesson Instance.
+	 * 
+	 * @param inputs
+	 *            The first array index represents the index of a training
+	 *            sample, the second represents the number of an input neuron.
+	 *            Attention: This array is cloned to avoid conflicts with
+	 *            changes in the original one.
+	 * @param desiredOutputs
+	 *            The first array index represents the index of a training
+	 *            sample, the second represents the number of an output neuron.
+	 *            Attention: This array is cloned to avoid conflicts with
+	 *            changes in the original one.
+	 */
+	public TrainingSampleLesson(double[][] inputs, double[][] desiredOutputs) {
+
+		if (inputs == null) {
+			throw new IllegalArgumentException("Inputs null.");
+		}
+		if (desiredOutputs == null) {
[…]

### 2015-02-19 Paper/PaperJahrSchlich.tex
Bilder für die Präsentation als .svg geadded, abgegebene Version des papers geadded

@@
 \setlength{\emergencystretch}{3pt}
-\hyphenation{sample-rate auto-nomous moni-toring methods}
+\hyphenation{sample-rate auto-nomous moni-toring methods during}
 
@@
 The obtained sensor data is collected by sensor nodes, and then analyzed and stored on a computer system over long periods of time. 
-The analysis of the sensor data can reveal abnormal changes in material and geometric behaviour at an early stage.
+The analysis of the sensor data can reveal abnormal changes in material and geometric behavior at an early stage.
 
@@
 Changes in the natural frequency result from changes in the vibration behavior, which indicate structural damage \citep{salawu1997detection}.
-The natural frequency can be determined by analysing the frequency spectrum of a vibration of a structure.
+The natural frequency can be determined by analyzing the frequency spectrum of a vibration of a structure.
 The frequency spectrum shows the relationship between the frequencies and magnitudes---magnitudes represent the influence a certain frequency has in a vibration.
@@
 This paper is organized as follows:
-First, a dezentralized wireless structural health monitoring system is designed and implemented. 
+First, a decentralized wireless structural health monitoring system is designed and implemented. 
 Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
@@
 6. information retrieval, see \citet{BisbySHM}.
-The dataflow is illustrated in \autoref{fig:flow}.
+The data flow is illustrated in \autoref{fig:flow}.
 Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and
@@
     \centering
-    \includegraphics{figures/dataflow_tasks_nn.eps}
-    \caption{Dataflow and associated SHM tasks}
+    \includegraphics[scale=0.9]{figures/dataflow_tasks_nn.eps}
+    \caption{Data flow and associated SHM tasks}
     \label{fig:flow}
@@
 The SHM system is implemented in an object-oriented way using Java programming language. 
-Object orientation uses objects, which are instanciated using classes. 
+Object orientation uses objects, which are instantiated using classes. 
 A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
@@
 The measured values are stored in the sensors nodes in an array.
-The different phases are indicatad by lighting different LEDs.
+The different phases are indicated by lighting different LEDs.
 The \texttt{FFT} class performs a FFT on the measured accelerations to transform the measured values from time domain into frequency domain. 
@@
 The \texttt{Communi\-cation} class ensures the wireless communication by opening a radio connection between the sensor node and the basestation to transfer the \texttt{Measure\-ment} objects from the sensor node to the basestation.
-For starting the operation of the sensor node, the entry point of the programm is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
+For starting the operation of the sensor node, the entry point of the program is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
 Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the monitoring tasks.
@@
     \includegraphics[width = \textwidth]{figures/uml-sensornode.pdf}
-    \caption{Class diagram of the \texttt{sensornode} package ((Update auf NN folgt))}
+    \caption{Class diagram of the \texttt{sensornode} package}
     \label{fig:UML-sn}
@@
 In this section, laboratory experiments, devised to validate the proposed SHM system, are described.
-First, a description of the test setup is given; second, the data aquisition, data processing and sensor fault detection are illustrated; and, finally, the results are discussed.
+First, a description of the test setup is given; second, the data acquisition, data processing and sensor fault detection are illustrated; and, finally, the results are discussed.
 
@@
 The plates are mounted on threaded rods with a vertical clearance of 23\,cm.
-At the bottom of the structure, the rods are fixed into a solid block of 40\,cm\,$\times$\,60\,cm\,$\times$\,30\,cm.
+At the bottom of the structure, the rods are embedded into a solid block of 40\,cm\,$\times$\,60\,cm\,$\times$\,30\,cm.
 The SHM system is installed on the test structure by mounting one wireless sensor node to each of the three top stories.
@@
 
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[width=\textwidth]{figures/testsetup.png}
+	\begin{subfigure}{0.48\textwidth}
+		\includegraphics[width=\textwidth]{figures/testsetupcorrect.jpg}
     	\caption{Regular sensor arrangement}
@@
     \end{subfigure}
-  	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[width=\textwidth]{figures/testsetup.png}
+    \quad
+  	\begin{subfigure}{0.48\textwidth}
+		\includegraphics[width=\textwidth]{figures/testsetuperror.jpg}
     	\caption{Simulated sensor fault}
@@
 If the deviation exceeds a threshold, a sensor fault is assumed.
-After the sensor fault detection, the following are stored into the database: the time of the measurement, the natural frequency, the measured magnitude, the expected magnitude, the deviation,  and a boolean value that indicates if a sensor fault was detected.
+After the sensor fault detection, the following variables are stored into the database: the time of the measurement, the natural frequency, the measured magnitude, the expected magnitude, the deviation,  and a boolean value that indicates if a sensor fault was detected.
 
@@
     \centering
-    %\includegraphics{figures/traininputs.pdf}
-    \caption{correlation between number of traing measurements and mean error during collection phase}
+    \includegraphics{figures/traininputs.pdf}
+    \caption{Correlation of number of training measurements and mean deviation \quad(between expected magnitude and measured magnitude)}
     \label{fig:traininputs}
@@
 The expected magnitudes deviated from the measured magnitudes (\autoref{tab:sensorfault}).
-The SHM system recognized the deviation and the boolean was set to "TRUE" accordingly.
+The SHM system recognized the deviation and the "fault detected" boolean was set to "TRUE" accordingly, showing that the implemented sensor fault detection increases the dependability of the SHM system.
 
@@
 
-Through the embedded FFT analysis of sampling events, the data traffic is decreased from 512 values per measurement and sensor node to 4 values per sampling event.
+Through the embedded FFT analysis of measurements, the data traffic is decreased from 512 values per measurement and sensor node to 4 values.
 A number stored in the used data format double occupies 8\,bytes of space.
@@
 Autonomous sensor fault detection has been realized by implementing a neural network into each sensor node.
-The neural networks have been trained to predict expected sensor measurements to be compared to actual measurments, in order to detect sensor faults.
-The structural integrity of the monitored structure are detected by observing the natural frequency.
+The neural networks have been trained to predict expected sensor measurements to be compared to actual measurements, in order to detect sensor faults.
+The structural integrity of the observed structure is monitored by calculating the natural frequency.
 To verify the proposed approach, the SHM system has been installed on a test structure to be tested in laboratory experiments.
-During several testruns, the prediction accuracy could be reduced up to a deviation of average 0.47\%.
+During several test runs, the prediction accuracy could be reduced up to a deviation of average 0.47\%.
 Simulated sensor faults have been detected reliably.

### 2015-02-24 Paper/finalBLAU.tex
Einpflegen der final remarks

--- 2015-02-19 Paper/PaperJahrSchlich.tex
@@
 
-%\makeatletter
-%\renewcommand\section{\@startsection{section}{1}{\z@}%
-%                     {-7.5ex \@plus -1ex \@minus -10.2ex}%
-%                     {2.3ex \@plus.2ex}%
-%                     {\sffamily\large\bfseries}}
-%\makeatother
-
 \usepackage{geometry}
@@
 
-\title{\textbf{ \Large {Decentralized autonomous sensor fault detection using neural networks}}}
+\title{\textbf{ \Large {Decentralized autonomous sensor fault detection in wireless sensor networks using neural networks}}}
 \author
@@
 
+\frenchspacing
+
 % Make the title.
@@
 \sloppy
-\setlength{\emergencystretch}{3pt}
+\setlength{\emergencystretch}{6pt}
 \hyphenation{sample-rate auto-nomous moni-toring methods during}
@@
 \citet{BisbySHM} defines SHM as "a non-destructive \textit{in-situ} structural evaluation method that uses any of several sensors which are attached to, or embedded in, a structure".
-The obtained sensor data is collected by sensor nodes, and then analyzed and stored on a computer system over long periods of time. 
-The analysis of the sensor data can reveal abnormal changes in material and geometric behavior at an early stage.
+\colorbox{cyan}{The} sensor data is collected by sensor nodes, and then analyzed and stored on a computer system. 
+Through analyses of the sensor data, abnormal changes in material and geometric behavior can be \colorbox{cyan}{discovered} at an early stage.
 
 Traditionally, the sensor nodes are connected to computer systems with cables.
-Using wired SHM systems has several disadvantages, including expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
-In wireless SHM systems, the sensor nodes communicate---through a basestation with each other and with computer systems--- via wireless transceivers, eradicating wiring-specific problems.
+The use of wired SHM systems has several disadvantages, for example expensive wiring, high installation and labor costs as well as inaccessibility of optimal sensor location with wires.
+
+Eradicating wiring-specific problems, the sensor nodes in wireless SHM systems communicate via wireless transceivers \colorbox{cyan}{[...]}.
 However, when using wireless SHM systems, special attention has to be paid to data transfer. 
@@
 
-As data reduction strategy, the natural frequency of a structure can be calculated on the sensor nodes instead of transmitting the raw measurements.
+As a data reduction strategy, the natural frequency of a structure can be calculated on the sensor nodes instead of transmitting all raw measurements.
 Changes in the natural frequency result from changes in the vibration behavior, which indicate structural damage \citep{salawu1997detection}.
@@
 The natural frequency is defined as the frequency with the highest magnitude.
-While the frequencies are specific for a structure, the scale of the magnitudes are dependent on the structure, the degree of the vibration and the observed location on the structure.
-During a vibration, the magnitudes on different locations are correlated with each other.
-\citep{rao2011fast}.
+While the frequencies are unique for a structure, the scale of the magnitudes depend not only on the structure, but also on the degree of the vibration and the sensor location at the structure.
+During a vibration, the magnitudes on different locations are correlated with each other \citep{rao2011fast}.
 
 Over their lifetime, the wireless sensor nodes can become inaccurate, faulty, or may even break.
-A fault can be defined as a defect of a sensor that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
+A fault can be defined as a defect of a sensor that leads to an error. 
+\colorbox{cyan}{An} error is the manifestation of a fault---an incorrect system state---that may result in a failure.
 To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
@@
 Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors \citep{Smarsly2014}.
-Virtual sensor measurements are computed for each sensor and then compared to the actual measurements. 
-If the setup of a structure is known, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor.
+Specifically, virtual sensor measurements are computed for each sensor and then compared to the actual measurements. 
+If the properties of a structure are known, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor, i.e.
+a priori knowledge about the structure is required.
 
@@
 
-In this study, a wireless SHM system capable of decentralized autonomous fault detection, is presented.
+In this study, a wireless SHM system capable of decentralized autonomous fault detection is presented.
 The sensor nodes calculate the natural frequency of the structure to monitor the structural integrity.
 The corresponding magnitudes are used for sensor fault detection:
-A neural network is implemented into each sensor node and trained---with magnitudes of past vibrations---to map the correlation of the magnitudes calculated by different sensor nodes.
-During runtime, the neural networks compute expected magnitudes for the embedding sensor node.
-A deviation between the expected and the measured magnitude indicates a sensor fault.
+A neural network is implemented into each sensor node. 
+Using the magnitudes of past vibrations, each neural network is trained to map the correlation of the magnitudes calculated by different sensor nodes.
+
+During runtime of the SHM system, the neural networks compute expected magnitudes for the respective sensor node.
+Deviations between the expected and the measured magnitude indicates a sensor faults.
 
@@
 Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
-Then, the system is tested in laboratory experiments. 
-Finally, the experimental results are discussed and future research directions are proposed.
+Finally, the system is tested in laboratory experiments \colorbox{cyan}{with simulated sensor faults} and the experimental results are discussed.
 
@@
 The sensor nodes and the basestation are of type "Oracle Sun SPOT". 
-The Sun SPOTS are equipped with several components, e.g. the sensor nodes include an internal accelerometer and eight independent RGB-LEDs.
-The "3-axis digital output accelerometer with three sensitivity ranges" \citep{eDemo2010} between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz.
+The Sun SPOTs are equipped with several components, e.g. the sensor nodes include an internal accelerometer and eight independent RGB-LEDs.
+The 3-axis digital output accelerometer ranges between $\pm$\,2\,g and $\pm$\,8\,g has a maximum sampling rate of 125\,Hz \citep{eDemo2010}.
 
-The SHM system performs the following monitoring tasks:
+The SHM system performs the following monitoring tasks, \colorbox{cyan}{as definded by \citet{BisbySHM}}:
 1. data acquisition,
@@
 5. diagnostics and 
-6. information retrieval, see \citet{BisbySHM}.
+6. information retrieval.
 The data flow is illustrated in \autoref{fig:flow}.
-Tasks 1 to 3 are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and
-determine the natural frequency and the corresponding magnitude of the structure using a fast Fourier transform (FFT).
-The magnitudes of the natural frequencies are fed into a neural network to detect sensor faults.
-The processed data is then transmitted via radio to the basestation and, finally, to the host computer.
+Tasks 1 to 3, \colorbox{cyan}{representing a sampling event}, are executed by the sensor nodes: During system operation, the sensor nodes acquire acceleration measurements and determine the natural frequency as well as the corresponding magnitude of the structure using a fast Fourier transform (FFT) and a peak picking algorithm.
+In every sensor node, the magnitudes of the natural frequencies are fed into a neural network to detect sensor faults.
+The processed data is then transmitted wirelessly to the basestation and, finally, to the host computer.
 On the host computer, task 4 is conducted by storing the data in a MySQL database.
-Task 5 and 6 are conducted on the host computer in further steps.
+Task 5 and task 6 are conducted on the host computer in further steps.
 
@@
 
-The neural network represents the structure of the SHM system.
-Each sensor node of the SHM system embeds a distinct neural network; a sensor node is represented by either an input neuron or an output neuron, as shown in \autoref{fig:neuralnetwork}.
-Measurements of other sensor nodes are fed into the input neurons of the neural network; the output neuron returns the predicted measurement of the embedding sensor node.
-Hidden neurons are used for computation and do not have a physical counterpart.
+The neural networks represent the structure of the SHM system.
+Each sensor node of the SHM system embeds a distinct neural network. As shown in \autoref{fig:neuralnetwork}, a sensor node is represented by either an input neuron or an output neuron.
+The measured magnitudes of other sensor nodes are fed into the input neurons of the neural network; the output neuron returns the predicted magnitudes of the \colorbox{cyan}{respective sensor node}.
+Hidden neurons are used for computation and \colorbox{cyan}{are not represented physically}.
 
@@
 Object orientation uses objects, which are instantiated using classes. 
-A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
-The SHM system is organized in two packages---\texttt{sensornode} and \texttt{basestation} that organize several Java classes belonging to a program.
+\colorbox{cyan}{Restlicher Absatz verändert: }
+A class includes methods and attributes.
+Methods allow the objects to perform actions; attributes store object-specific data.
+The SHM system consists of two packages, \texttt{sensornode} and \texttt{basestation}, that organize several Java classes.
 
-\autoref{fig:UML-sn} illustrates the classes of the \texttt{sensor\-node} package.
-The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Measure\-ment}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the Java library "SNIPE" \citep{SNIPE} that implement a framework for neural networks.
-The classes are embedded directly into the sensor nodes.
+\autoref{fig:UML-sn} illustrates the classes of the \texttt{sensor\-node} package in detail.
+The package \texttt{sensor\-node} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Sample}, \texttt{Network\-Training}, \texttt{Communi\-cation} and \texttt{Main\-Spot}, forming the SHM system, and several classes obtained from the external Java library "SNIPE" \citep{SNIPE} which is used to implement the neural networks.
+The classes \colorbox{cyan}{[...]} are embedded directly into the sensor nodes.
 
+For starting the operation of the sensor node, the entry point of the program is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
+Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to autonomously perform the monitoring tasks.
 The \texttt{Acceleration\-Sampler} class is responsible for measuring the acceleration.
-There are two phases: first, the acceleration is measured with a low sampling rate.
-Once the acceleration exceeds a threshold, the second phase is entered by increasing the sampling rate. 
-The measured values are stored in the sensors nodes in an array.
-The different phases are indicated by lighting different LEDs.
+There are two phases: First, \colorbox{cyan}{before the excitation,} the acceleration is measured with a low sampling rate.
+Once the acceleration exceeds a pre-defined threshold, the second phase is entered by increasing the sampling rate. 
+The measured values are stored in the sensor nodes in an array.
+The different phases are indicated by lighting different LEDs of the sensor nodes.
 The \texttt{FFT} class performs a FFT on the measured accelerations to transform the measured values from time domain into frequency domain. 
-With the transformed data, the magnitudes and the correlating frequencies of the measured vibrations are calculated.
-Finally, the natural frequency is determined by extracting the maximal magnitude.
-The \texttt{Measure\-ment} class defines analyzed measurements by storing the following parameters: the address of the sensor node, the maximal magnitude, the natural frequency and a boolean value that indicates if a sensor fault was detected during runtime.
-Using the SNIPE library, a neural network is described and created for fault detection.
-The \texttt{Network\-Training} class defines a routine to train the neural network with previously measured \texttt{Measure\-ment} objects.
-The \texttt{Communi\-cation} class ensures the wireless communication by opening a radio connection between the sensor node and the basestation to transfer the \texttt{Measure\-ment} objects from the sensor node to the basestation.
-For starting the operation of the sensor node, the entry point of the program is the \texttt{start\-App()} method in the \texttt{Main\-Spot} class. 
-Within the \texttt{Main\-Spot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class, the SNIPE neural network classes, the \texttt{Network\-Training} class and the \texttt{Communi\-cation} class are created to perform the monitoring tasks.
+With the transformed data, the frequencies and the magnitudes of the measured vibrations are calculated.
+Finally, the natural frequency is determined by extracting the maximum magnitude.
+The \texttt{Sample} class defines objects representing analyzed measurements.
+\colorbox{cyan}{The \texttt{Sample} objects store} the following parameters: The address of the sensor node, the maximum magnitude, the natural frequency, and a boolean value that indicates if a sensor fault was detected during runtime.
+Using the SNIPE library, a neural network is implemented for fault detection.
+The \texttt{Network\-Training} class defines a routine to train the neural network with \texttt{Sample} objects of previous sampling events.
+The \texttt{Communi\-cation} class ensures the wireless communication by opening a radio connection between the sensor node and the basestation to communicate the \texttt{Sample} objects from the sensor node to the basestation.
 
@@
 The package \texttt{base\-station} runs on the host computer and operates the basestation.
-It includes the classes \texttt{Database\-Handler} and \texttt{Main\-Base}.
-The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
+It includes the classes \texttt{Database\-Handler}, \texttt{Main\-Base}, and \colorbox{cyan}{\texttt{Sample}, as described in the package \texttt{sensornode}}.
+
 The entry point of the program is the \texttt{run()} method in the \texttt{Main\-Base} class. 
 The \texttt{Main\-Base} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
+The \texttt{Database\-Handler} class establishes a connection to a MySQL database installed on the host computer, creates a database table, unless the specified table already exists, and inserts the data into the database table.
 
@@
 	\end{subfigure}
-  	\caption{Laboratory setup (Source: own photographs)}
+  	\caption{Laboratory setup}
 \end{figure}
 
-Several test runs were conducted.
+Several test runs have been conducted.
 During each test run, all sensor nodes run simultaneously.
-A test run includes a training phase and a data collection phase.
-The training phase consists of the description, the creation and the training of a neural network.
-During the training of the neural network, several measurements are performed and used as training input.
-A measurement includes the excitation of the structure, acceleration sampling, calculation of the natural frequency and corresponding magnitude, and data exchange with the other sensor nodes.
+A test run includes 1. a training phase and 2. a data collection phase \colorbox{cyan}{on every sensor node}.
+The training phase consists of the implementation and the training of a neural network.
+During the training of the neural network, several sampling events are performed and used as training input.
+A sampling event includes the excitation of the structure, sampling of 512 acceleration measurements, calculation of the natural frequency and corresponding magnitude, and data exchange with the other sensor nodes.
 
-The data collection phase consists of any desired number of measurements, sensor fault detection, and data storage.
-After every measurement, the measured magnitudes of the other sensor nodes are used as input of the neural network to predict the expected magnitude of the embedding sensor nodes for sensor fault detection.
-The deviation of the expected magnitude and the measured magnitude is calculated.
+The data collection phase \colorbox{cyan}{consists of any desired number} of sampling events, sensor fault detection, and data storage.
+After every sampling event, the measured magnitudes of the other sensor nodes are used as input of the neural network to predict the expected magnitude of the embedding sensor nodes for sensor fault detection.
+The deviation of the expected magnitude and the measured magnitude is calculated by the sensor node.
 If the deviation exceeds a threshold, a sensor fault is assumed.
-After the sensor fault detection, the following variables are stored into the database: the time of the measurement, the natural frequency, the measured magnitude, the expected magnitude, the deviation,  and a boolean value that indicates if a sensor fault was detected.
+Upon sensor fault detection, the following variables are stored into the database: the time of the sampling event, the natural frequency, the measured magnitude, the expected magnitude, the deviation,  and a boolean value that indicates if a sensor fault has been detected.
 
-The neural network has been tested with varying numbers of training measurements (\autoref{fig:traininputs}).
-With increasing numbers of training measurements, the prediction accuracy of the expected magnitude increases.
-With 24 training measurements, a minimum deviation---between the expected magnitude and the measured magnitude---of average 0.47\% has been reached.
+The neural networks have been tested with varying numbers of sampling events \colorbox{cyan}{as training input} (training events) (\autoref{fig:traininputs}).
+With increasing numbers of (training events), the prediction accuracy of the expected magnitude increases.
+With 24 sampling events, a minimum deviation between the expected magnitude and the measured magnitude of 0.47\% in average has been reached.
 
@@
     \includegraphics{figures/traininputs.pdf}
-    \caption{Correlation of number of training measurements and mean deviation \quad(between expected magnitude and measured magnitude)}
+    \caption{Correlation of number of training events and mean deviation between expected magnitudes and measured magnitudes}
     \label{fig:traininputs}
@@
 
-To validate the autonomous sensor fault detection, a sensor fault was simulated by shifting and rotating one of the sensor nodes after completion of the training phase.
-The expected magnitudes deviated from the measured magnitudes (\autoref{tab:sensorfault}).
-The SHM system recognized the deviation and the "fault detected" boolean was set to "TRUE" accordingly, showing that the implemented sensor fault detection increases the dependability of the SHM system.
+To validate the autonomous sensor fault detection, a sensor fault has been simulated by shifting and rotating one of the sensor nodes after completing the training phase.
+As a result, it has been observed that the expected magnitudes significantly deviate from the measured magnitudes (\autoref{tab:sensorfault}).
+The SHM system has recognized the deviation, corroborating that the implemented sensor fault detection increases the dependability of the SHM system.
 
@@
 	\centering
-	\caption{Deviation between measured magnitude and expected magnitude with and without simulated sensor faults}
+	\caption{Deviation between measured magnitudes and expected magnitudes with and without simulated sensor faults}
 	\begin{tabular}{c c c c}
@@
 
-Through the embedded FFT analysis of measurements, the data traffic is decreased from 512 values per measurement and sensor node to 4 values.
-A number stored in the used data format double occupies 8\,bytes of space.
-The data traffic is decreased by (512$-$4)$\times$8\,B\,=\,4064\,B\,=\,3.97\,MB per sampling event, decreasing data traffic by 99.3\,\%.
+In total, through the embedded FFT analysis of acceleration measurements, the data traffic is decreased from 512 values to 4 values per sampling event and sensor node.
+A value stored in the double data format requires 8\,bytes of space.
+Thus, the data traffic is decreased by (512$-$4)$\times$8\,B\,=\,4,064\,B\,=\,3.97\,MB per sampling event, decreasing data traffic by 99.3\,\%.
 
@@
 
-This paper has shown a decentralized autonomous sensor fault detection strategy for structural health monitoring systems based on neural networks. 
-Autonomous sensor fault detection has been realized by implementing a neural network into each sensor node.
-The neural networks have been trained to predict expected sensor measurements to be compared to actual measurements, in order to detect sensor faults.
+This paper has presented a decentralized autonomous sensor fault detection strategy for structural health monitoring systems based on neural networks. 
+Autonomous sensor fault detection has been realized by implementing neural networks into the sensor nodes.
+The neural networks have been trained to predict expected sensor data to be compared to measured data, \colorbox{cyan}{in order to} detect sensor faults.
+To verify the proposed approach, the SHM system has been installed on a test structure to be tested in laboratory experiments.
 The structural integrity of the observed structure is monitored by calculating the natural frequency.
-To verify the proposed approach, the SHM system has been installed on a test structure to be tested in laboratory experiments.
-During several test runs, the prediction accuracy could be reduced up to a deviation of average 0.47\%.
 Simulated sensor faults have been detected reliably.
-Through decentralized data processing, embedded in each sensor node, the data traffic has been reduced by 99.3\,\%.
-It has been proven, that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.
+Through decentralized data processing, implemented into each sensor node, the data traffic has been reduced by 99.3\,\%.
+In summary, it has been proven, that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.
 
